### 2023-08-25 Manuscript/proof_outline.tex
ported the manuscript file to lean4/mathlib4

% please run using
% lualatex --shell-escape proof_outline.tex

\documentclass[sigplan,screen]{acmart}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
%\usepackage{amsmath}

\pdfoutput=1

\usepackage{minted}
\setminted{encoding=utf-8}
\usepackage{fontspec}
%\setmainfont{dejavu-serif}
\setmonofont[Scale=0.833]{dejavu-sans}
%\setmainfont{FreeSerif}
%\setmonofont{FreeMono}
\usepackage{xcolor, graphics}
\definecolor{mygray}{rgb}{0.97,0.97,0.97}
\newcommand{\leanline}{\mintinline[breaklines, breakafter=., breaklines]{Lean}}
%\usepackage{refcheck}

% Definitions and packages for Lean code
\usepackage{color}
\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green

\usepackage{enumitem}
\setlist[enumerate]{%
align=left, itemsep=0pt, leftmargin=0pt, labelindent=0pt, listparindent=0pt, labelwidth=0pt, itemindent=!
}
%\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
%\newtheorem{theorem}[proposition]{Theorem}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[proposition]{Lemma}
\newtheorem{corollary}[proposition]{Korollar}
\newtheorem{definition}[proposition]{Definition}
\newtheorem{remark}[proposition]{Remark}

\author{Rémy Degenne}
\affiliation{%
   \institution{Univ. Lille, Inria, CNRS, Centrale Lille, UMR 9189-CRIStAL}
   \city{Lille}
   \country{France}}
\email{remy.degenne@inria.fr}

\author{Peter Pfaffelhuber}
\affiliation{%
   \institution{University of Freiburg}
   \city{Freiburg}
   \country{Germany}}
\email{p.p@stochastik.uni-freiburg.de}

\title{Formalizing the Kolmogorov extension Theorem in mathlib4}

\keywords{probability, measure theory, Lean4, formal mathematics, proof assistant, mathlib4}

% Tool to generate below available at http://dl.acm.org/ccs.cfm.
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10002950.10003648</concept_id>
%<concept_desc>Mathematics of computing~Probability and statistics</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10003752.10003790.10002990</concept_id>
%<concept_desc>Theory of computation~Logic and verification</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Mathematics of computing~Probability and statistics}
%\ccsdesc[500]{Theory of computation~Logic and verification}

%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
%%% Then please reload our instructions page and copy-and-paste the NEW commands into your article.
%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
%%%
%%% The following is specific to CPP '23 and the paper
%%% 'A Formalization of Doob's Martingale Convergence Theorems in Mathlib'
%%% by Kexing Ying and Rémy Degenne.
%%%
%\setcopyright{licensedothergov}
%\acmPrice{15.00}
%\acmDOI{10.1145/3573105.3575675}
%\acmYear{2023}
%\copyrightyear{2023}
%\acmSubmissionID{poplws23cppmain-p22-p}
%\acmISBN{979-8-4007-0026-2/23/01}
%\acmConference[CPP '23]{Proceedings of the 13th ACM SIGPLAN International Conference on Certified Programs and Proofs}{January 13--14, 2024}{London, UK}
%\acmBooktitle{Proceedings of the 13th ACM SIGPLAN International Conference on Certified Programs and Proofs (CPP '24), January 13--14, 2024, London, UK}
%\received{2023-09-18}
%\received[accepted]{??}

\usepackage{environ}
\NewEnviron{killcontents}{}
\let\proof\killcontents
\let\endproof\endkillcontents

\begin{document}

\begin{abstract}
    xxx
\end{abstract}

\maketitle

{\color{blue}
  \noindent
  TODO
  \begin{itemize}
  \item Make sure that the line of argument for code and mauscript are in line.
  \end{itemize}

}

\section{Introduction}
One of the main building blocks of modern probability theory are stochastic processes, which are usually defined as any collection of random variables -- $(X_t)_{t\in \iota}$ with $X_t$ taking values in some $\alpha_t$, say -- defined on some joint probability space $(\Omega, \mathcal A, \mathbf P)$. In order to study such processes, it is fundamental to talk about their distribution, i.e.\ a probability distribution on the product set $\prod_{t\in\iota} \alpha_t$. The usual approach to construct (the distribution of) a stochastic process works as follows: describe properties of the distribution of the stochastic process $P_J$ at some finite number of times $J = \{t_1,...,t_n\} \subseteq \iota$. The resulting family of probability measures $(P_J)_{J \subseteq \iota \text{ finite}}$ has to be {\em projective} in the sense that the projection of $P_J$ to $H\subseteq J$ has to be equal to $P_H$. In other words, when describing the distribution of the stochastic process at all times in $J$, and then forgetting (i.e.\ projecting) all properties for times in $J\setminus H$, results in the description of properties at times in $H$. One may then ask if this already gives a complete description of the process for all times. For uncountable $\iota$, e.g. $\iota = [0,\infty)$, one is tempted to be pessimistic at first sight since measures -- which describe the distribution of the stochastic process -- usually only deal well with a countable number of measurable events. However, it is the achievement of Kolmogorov that the finite-dimensional distributions in fact provide a unique description of the distribution of a stochastic process, as long as the underlying family of state spaces $(\alpha_t)_{t\in\iota}$ is nice enough. This means that these spaces are usually required to be complete and separable metric (in particular topological) spaces. This distrubution is a measure on the product-$\sigma$-field $\mathcal F :=\bigotimes_{t\in\iota} \mathcal B(\alpha_t)$ (assuming that $\mathcal B(\alpha_t)$ is the Borel $\sigma$-algebra on $\alpha_t$). Here, $\mathcal F$ is generated by finite projections and hence any element of $\mathcal F$ may only depend on at most countably many $t\in \iota$, making this a rather coarse $\sigma$-algebra. (In particular, note that this is not the Borel $\sigma$-algebra of the product topology.) The result responsible for this insight is usually denoted the {\em Kolmogorov extension theorem}, formulated in \cite{kolmogoroff1933grundbegriffe}. We note that a version of the extension result was proved by Daniell in the 1930s, but this paper was not acknowledged by the probabilists of that time \cite{aldrich2007but}. Due to his contribution, the theorem is often called the {\em Daniell-Kolmogorov extension}.

The goal of our contribution is a formalization of the proof of that theorem in {\tt mathlib4}, the mathematical library of {\tt Lean3} \cite{mathlib2020lean}. The Kolmogorov extension theorem is on the interface between measure theory and probability theory. Here, we rely on a decent amount of formalized mathematics in the measure-theory part of {\tt mathlib4} (outer measures, above all), while not requiring any specific previous formalization of probability theory. (In fact, most of our results are formulated in terms of finite rather than probability measures.)

At first sight, it might be surprising that martingales, a certain class of stochastic process, have already been formalized in Lean \cite{ying2023formalization}, although the Kolmogorov extension theorem (on the existence of stochastic processes) is only available by our contribution. Note, however, that a martingale $(X_t)_{t\in \iota}$ is defined as a family of random variables (satisfying some properties) on a fixed probability space $(\Omega, \mathcal F, \mathbf P)$, while the Kolmogorov extension theorem is on the construction of a probability (or finite) measure $\mathbf P$, on which we can define random variables (with certain properties). So, our work complements \cite{ying2023formalization} in the sense that probability spaces $(\Omega, \mathcal F, \mathbf P)$ on which you can define a martingale exist. However, we do not give any specific example for the construction of a specific instance of such a space.

The main steps in our construction, which were previously missing in {\tt mathlib4}, are (more mathematical details are below):
\begin{enumerate}
\item a formalization of (the set-system of) semi-rings (see Definition~\ref{def:semi} below); 
\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
\item a proof that single probability measures on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg} for a definition; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below); 
\item the classical Carathéodory extension theorem, providing us a candidate for the measure which we want to construct (see Theorem~\ref{T:masseind}); 
\item the Kolmogorov extension theorem, as based on the previous steps.
\end{enumerate}
The Kolmogorov extension theorem has been previously formalized in Isabelle/HOL \cite{Immler2012}. However, this formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical. However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations in these two interactive theorem provers.

\subsubsection*{Possible future work}
Let us describe some future projects extending {\tt mathlib4} which become possible by our contribution.

\noindent
{\em Instances of stochastic processes:} An obvious application of Kolmogorov's extension theorem is the construction of basic stochastic processes like Markov chains \cite[Section~11]{Kallenberg2020}, the Poisson process \cite[Section~13]{Kallenberg2020} and Brownian motion \cite[Section~14]{Kallenberg2020}. (Note that other Gaussian processes -- indexed by $[0, \infty)$ or $\mathbb R$ might as well be constructed the same way. In addition, fields like the Gaussien free field indexed by $\mathbb R^d$ \cite{sheffield2007gaussian} can also be given.) For these tasks, we would have to define the finite-dimensional distributions (Poisson and normally distributed, respectively), and apply the extension theorem. This task requires the formalization of multi-dimensional Poisson and normal distributions, which -- in textbooks -- is usually done using characteristic functions. Since these are not yet part of {\tt mathlib4}, we postpone this task to the future.

\noindent
{\em Sample-path properties:} The Kolmogorov extension gives the existence of a distribution of a stochastic process $(X_t)_{t\in\iota}$ with certain properties. Extra work is needed in order to show that -- on the same probability space -- we can as well define a version (i.e.\ another process $(Y_t)_{t\in\iota}$ with $\mathbf P(X_t = Y_t) = 1$ for all $t\in\iota$) which is right-continuous with left limits (for the Poisson process) or continuous (for Brownian motion), respectively. For the former, this follows at least from some general principles of Markov processes (e.g.\ \cite[Theorem 4.3.6]{EthierKurtz1986}). For the latter, this requires formalization of the Kolmogorov-Chentsov criterion 
\cite[Theorem~4.23]{Kallenberg2020}.

\section{Formalization of the Daniell-Kolmogorov extension}
We start by stating the exact result. We are going to formulate the main result in a modern fashion, as e.g.\ found in Theorem 2.2 of \cite{rao1971projective}, Theorem 7.7.1 of Volume~2 of \cite{bogachev2007measure}, Theorem 15.26 of \cite{guide2006infinite}, or \cite{border1998expository}. Note that these formulations split general assumptions on the underlying space(s) (e.g.\ a metric property) from the property which is needed in the proof (inner regularity with respect to compact sets). Other -- highly readable -- references such as \cite{Billingsley1995} state the extension theorem only in special cases such as $\alpha_t = \mathbb R$ for all $t$. 
%In addition, there are still new proofs found for the general extension theorem \cite{chin2019new}, but they require some results not yet formalized in mathlib4 (Borel isomorphism of $\mathbb R$ to $\{0,1\}^{\mathbb N}$), so we do not follow this route.  
Whenever we feel necessary, we rely on a standard textbooks on Probability Theory \cite{Klenke2013, Kallenberg2020} for general purpose.

Before we can state the result, we begin with some mathematical notions. We start off slowly from metric spaces, but quickly introduce projective families of measures. In all definitions, $\alpha$ will be some set (or type, since {\tt Lean4} is a dependently typed language, \leanline{α : Type _}).

\begin{definition}[Metric and topological spaces]
  \begin{enumerate}
  \item \sloppy A pseudo-metric on $\alpha$ is a symmetric map $r : \alpha
    \times \alpha \to [0,\infty)$ satisfying the triangle inequality,
      i.e.\ $r(x,z) \leq r(x,y) + r(y,z)$ for all $x,y,z\in\alpha$.
      \\ If $r$ also satisfies $r(x,y) = 0$ iff $x=y$, we call it a
      metric.  \\ If $r : \alpha \times \alpha \to [0,\infty]$
      (i.e.\ $r(x,y) = \infty$ is allowed), we call $r$ an extended
      (pseudo-)metric.
    \item Let $r$ be an (extended pseudo-)metric. A sequence $x_1,
      x_2,... \in \alpha$ is called Cauchy if for all $\varepsilon>0$
      there is $N\in\mathbb N$ such that $r(x_n, x_m) < \varepsilon$
      for all $m,n > N$. The (pseudo-extended) metric is called
      complete if every Cauchy-sequency has a limit in
      $\alpha$.
    \item Some\footnote{We denote by $2^\alpha$ the power set of
      $\alpha$, i.e.\ the set of all subsets of $\alpha$.} $\mathcal O
      \subseteq 2^\alpha$ is called a topology, if it satisfies (i)
      $\emptyset, \alpha \in \mathcal O$, (ii) $\mathcal O$ is stable
      under arbitrary unions, i.e.\ if $A_i \in \mathcal O$ for all
      $i\in\iota$ and $\iota$ is arbitrary, then $\bigcup_{i\in \iota}
      A_i \in \mathcal O$, (iii) $\mathcal O$ is stable under finite
      intersections, i.e.\ if $A_1,...,A_n \in \mathcal O$, then
      $\bigcap_{i\in \iota} A_i \in \mathcal O$.
  \end{enumerate}
\end{definition}

\begin{remark}[Metric and topological spaces in {\tt mathlib4}] \mbox{}
  \begin{enumerate}
    \item \sloppy In \leanline{Lean4}, we have the \leanline{class
      PseudoEMetricSpace (α : Type u) extends EDist}, which comes with
      the components, \leanline{edist_self}, \leanline{edist_comm},
      \leanline{edist_triangle} providing the properties of the
      pseudo-metric, and \leanline{toUniformSpace},
      \leanline{uniformity_edist} defining a uniform space from the
      extended pseudo-metric. Such a space does not come with a
      metric, but with a filter on $\alpha \times \alpha$, which
      describes which points in $\alpha$ are near. For example, the
      diagonal of $\alpha \times \alpha$ is a subset of all sets in
      the uniformity; see \cite{james2013topologies} for details. We
      note that uniform spaces with a countably generated uniformity
      filter is pseudometrizable, i.e.\ there exists a
      pseudo-metric-space structure that generates the same
      uniformity; see \leanline{UniformSpace.pseudoMetrizableSpace},
      which formalizes a result stated in
      \cite{melikhov2011metrizable}.
    \item \sloppy As for topological spaces, we have \leanline{class
      TopologicalSpace(α : Type u) : Type u}, which comes with
      \leanline{isOpen}, \leanline{isOpen_univ},
      \leanline{isOpen_inter} and \leanline{isOpen_sUnion}, which are
      exactly the properties of a topological space described above.
    \item Any (extended pseudo-)metric on $\alpha$ defines a topology,
      namely the topology generated by $\mathcal H := \{\{y:
      r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
      (0,\infty)\}$. Actually, in \leanline{mathlib4}, as we saw
      above, every extended pseudo-metric defines a uniform space, and
      a uniform space is a \leanline{class UniformSpace(α : Type u)
        extends TopologicalSpace , UniformSpace.Core} with xxx
      \leanline{to_topological_space} giving a topological space,
      \leanline{to_core} which stores the uniformity filter, and
      \leanline{is_open_uniformity}, which gives all open sets. In
      particular, a uniform space defines a topological space, which
      can be used by the typeclass system.
  \end{enumerate}
\end{remark}

\begin{remark}[Generated topology]
  \begin{enumerate}
  \item The intersection of any number of topologies is again a
    topology. For this reason, if $\mathcal H \subseteq 2^\alpha$, we
    define the topology $\mathcal O := \bigcap_{\mathcal F \supseteq
      \mathcal H \text{ topology}} \mathcal F$; see
    \leanline{TopologicalSpace.generateFrom}. This is called the
    topology generated by $\mathcal H$. If $\mathcal H$ is closed
    under finite intersections, we call $\mathcal H$ a basis for
    $\mathcal O$; see \leanline{TopologicalSpace.IsTopologicalBasis}.
  \item \sloppy We call the topology (generated from an extended
    pseudo-metric) separable (see
    \leanline{TopologicalSpace.SeparableSpace}) if there is a
    countable $s \subseteq \iota$ such that $\inf\{r(x,y) : y \in s\}
    = 0$ for all $x\in\alpha$. \\ If there is a countable basis of the
    topology, it is separable; see
    \leanline{TopologicalSpace.SecondCountableTopology.to_separableSpace}.
  \end{enumerate}
\end{remark}

\noindent
Finally, we can introduce measures. Still, we need even more
set-systems, i.e.\ $\sigma$-algebras.

\begin{definition}[$\sigma$-algebras and measures]
  \begin{enumerate}
  \item 
    We call $\mathcal F \subseteq 2^\alpha$ a $\sigma$-algebra (on
    $\alpha$) if (i) $\alpha \in \mathcal F$, (ii) $\mathcal F$ is
    stable under complements, i.e.\ $A \in \mathcal F \Rightarrow A^c
    \in \mathcal F$, (iii) $\mathcal F$ is stable under countable
    unions, i.e.\ $A_1, A_2,... \in \mathcal F \Rightarrow
    \bigcup_{n=1}^\infty A_n \in \mathcal F$. We call $(\alpha,
    \mathcal F)$ a measurable space.
  \item For some $\sigma$-algebra $\mathcal F$ on $\alpha$, a function
    $\mu : \mathcal F \to [0, \infty]$ is called a measure,
    if\footnote{Since $\alpha \in \mathcal F$ and $\mathcal F$ is
      stable under complements, we also have that $\emptyset \in
      \mathcal F$} (i) $\mu(\emptyset) = 0$, (ii) $\mu$ is countably
    additive, i.e.\ for $A_1, A_2,... \in \mathcal F$ pairwise
    disjoint, we have $\mu\Big( \bigcup_{n=1}^\infty A_n\Big) =
    \sum_{n=1}^\infty \mu(A_n)$. We call $(\alpha, \mathcal F, \mu)$ a
    measure space. \\ In addition, $\mu$ is called finite if
    $\mu(\alpha) < \infty$ and a probability measure if
    $\mu(\alpha)=1$.
  \end{enumerate}
\end{definition}

\begin{remark}[Measur(abl)e spaces in \leanline{mathlib4}]
  \begin{enumerate}
  \item \sloppy The \leanline{class MeasurableSpace(α : Type u_7) :
    Type u_7} is very similar to a topological space in
    \leanline{mathlib4}, since it comes with
    \leanline{MeasurableSet'}, \leanline{measurableSet_empty},
    \leanline{measurableSet_compl}, and
    \leanline{measurableSet_iUnion}.
  \item For a measure space, we require a measurable space as a
    variable, and add the set function: \leanline{structure
      MeasureTheory.Measure(α : Type u_6) [inst : MeasurableSpace α]
      extends MeasureTheory.OuterMeasure}. Here $\sigma$-additivity of
    a \leanline{Measure α} reads xxx note on outer measures
\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  m_iUnion : ∀ ⦃f : ℕ → Set α⦄,
    (∀ (i : ℕ), MeasurableSet (f i)) →
    Pairwise (Disjoint on f) →
    ↑toOuterMeasure (⋃ (i : ℕ), f i) =
    ∑' (i : ℕ), ↑toOuterMeasure (f i)
\end{minted}    
  \end{enumerate}
\end{remark}

\noindent

\begin{remark}[Generated $\sigma$-algebra, image measure]
  We will frequently need two basic results: 
  \begin{enumerate}
  \item The intersection of any number of $\sigma$-algebras is again a
    $\sigma$-algebra. For this reason, if $\mathcal H \subseteq
    2^\alpha$, we define the $\sigma$-algebra $\sigma(\mathcal H) :=
    \bigcap_{\mathcal F \supseteq \mathcal H \text{ $\sigma$-algebra}}
    \mathcal F$. This is called the $\sigma$-algebra generated from
    $\mathcal H$; see \leanline{MeasurableSpace.generateFrom}.  \\ In
    particular, if $\mathcal O$ is a topology on $\alpha$, we call
    $\mathcal B := \sigma(\mathcal O)$ the Borel $\sigma$-algebra on
    $\alpha$. In \leanline{mathlib4}, this is
    \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
      def borel (α : Type u) [TopologicalSpace α] :
        MeasurableSpace α :=
        generateFrom { s : Set α | IsOpen s }
    \end{minted}
    In addition, \leanline{mathlib4} provides a similar notion, which is
    \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
      class OpensMeasurableSpace (α : Type*)
        [TopologicalSpace α] [h : MeasurableSpace α] :
        Prop where borel_le : borel α ≤ h
    \end{minted}
    Here, all open sets are measurable, so the $\sigma$-algebra defining
    \leanline{h} might be larger than the Borel $\sigma$-algebra. We will
    need both notions in our proofs.
  \item Let $(\alpha, \mathcal F)$ and $(\beta, \mathcal G)$ be
    measurable spaces. Some $f : \alpha \to \beta$ is called
    measurable (with respect to $\mathcal F$ and $\mathcal G$) if
    $f^{-1} \mathcal G \subseteq \mathcal F$.  \\ In this case, if
    $\mu$ is a measure on $\mathcal F$, the measure $\nu: \mathcal G
    \to [0, \infty], \nu(B) := \mu (f^{-1}(B))$ is called the image
    (or push-forward) measure of $\mu$ under $f$. We write $\nu :=
    f_\ast \mu$; see \leanline{MeasureTheory.Measure.map}.
  \end{enumerate}
\end{remark}

The extension theorem is on finite measures on product spaces. In
particular, the product is taken over a set of arbitrary
cardinality. The next definition covers the important concept of a
projective family of measures. In short term, we define measures on
any finite subset of indices in a consistent way. Recall that finite
sets (of some type \leanline{ι} are formalized with \leanline{Finset
  ι}. Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
quotient type of the \leanline{List}-type with respect to permutations
of the list) together with a proof that it contains no duplicates.

\begin{definition}[Projective family and projective limit] \mbox{}
  \begin{enumerate}
  \item For some set $\iota$, we will write $J\subseteq_f \iota$ if
    $J\subseteq \iota$ and $J$ is finite.
    %We will write $J\subseteq_c \iota$ if $J\subseteq \iota$ and $J$
    %is countable.
  \item Let $\iota$ be some (index) set and $(\alpha_i)_{i\in\iota}$ a
    family of sets. For $J\subseteq \iota$, we denote $\alpha_J :=
    \prod_{j \in J} \alpha_j$ and $\pi_J : \alpha_\iota \to \alpha_J$
    the projection. For $H\subseteq J \subseteq \iota$, we write
    $\pi_H^J$ for the projection $\alpha_J \to \alpha_H$.
  \item Let $\mathcal F_i$ be a $\sigma$-algebra in $\alpha_i$,
    $i\in\iota$. For $J\subseteq_f \iota$, let $\mathcal F_J$ be the
    product-$\sigma$-algebra on $\alpha_J$, and $\mathcal F_\iota$ be
    the $\sigma$-algebra generated by cylinder sets
    $\{\pi_J^{-1}\prod_{j\in J} A_j: J \subseteq_f \iota, A_j \in
    \sigma(E_j), j\in J\}$.
  \item A family $(P_J)_{J\subseteq_f I}$, where $P_J$ is a finite
    measure on $\mathcal F^J$, is called projective if
    $$P_H = (\pi_H^{J})_\ast P_J$$ for all
    $H\subseteq J \subseteq_f I$. (Recall that $A \mapsto
    (\pi_H^{J})_\ast P_J(A) := P_J((\pi_H^{J})^{-1}A)$
    is called the image measure of $P_J$ under $\pi_H^{J}$.)
  \item If, for some projective family $(P_J)_{J\subseteq_f \iota}$,
    there is a finite measure $P_\iota$ on $\mathcal F_\iota$ with
    $P_J = (\pi_J)_\ast P_\iota$ for all $J\subseteq_f \iota$, then we
    call $P_\iota$ projective limit of $(P_J)_{J\subseteq_f \iota}$.
  \end{enumerate}
\end{definition}

In the formalization, we use \leanline{variable {ι : Type _} {α : ι →
    Type _} }, which fixes the index set $\iota$ and all spaces
$\alpha_t, t\in\iota$. In addition, note that the projective property
we use here works as long as we have a preorder (which is the subset
relation on \leanline{Finset ι} below).

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def IsProjective [Preorder ι]
    (P : ∀ j : ι, α j) (π : ∀ {i j : ι}, j ≤ i → α i → α j) : Prop :=
    ∀ (i j) (hji : j ≤ i), P j = π hji (P i)
\end{minted}

\noindent
With this, we can define the projetive family as follows. Note that
\leanline{Lean4} automatically uses the subset relation as
\leanline{[Preorder (Finset ι)]} when \leanline{isProjective} is
called.

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def IsProjectiveMeasureFamily
    [∀ i, MeasurableSpace (α i)]
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j)) :
    Prop :=
    IsProjective P (fun I _ hJI μ => μ.map
    fun x : ∀ i : I, α i => fun j => x ⟨j, hJI j.2⟩ :
    ∀ (I J : Finset ι) (_ : J ⊆ I), Measure (∀ i : I, α i)
    → Measure (∀ j : J, α j))
\end{minted}

\noindent
It is worth understanding the precise connection ot
\leanline{isProjective} and \leanline{isProjectiveMeasureFamily}. In
the latter, the first variable of \leanline{isProjective} is the
family \leanline{P} of finite measures for all finite subsets of
\leanline{ι}. The second variable is the functions which maps two sets
\leanline{I J : Finset ι} and a proof \leanline{hJI} of \leanline{J ⊆
  I} together with \leanline{P I} to the image measure on
\leanline{J}, which is \leanline{μ.map (fun x : ∀ i : I, α i => fun j
  => x ⟨j, hJI j.2⟩)}. The map defined by the \leanline{=>}-notation
maps every $(x_i)_{i\in I}$ to a function of \leanline{j}, whose type
is the subtype of \leanline{I}, consisting of a value and a proof of
\leanline{J ⊆ I}. In other words, this is $(x_j)_{j\in J}$.

\noindent
Now we are ready to formulate the Kolmogorov extension theorem:

\begin{theorem}[Kolmogorov extension]\label{T1}
  For all $t\in\iota$, let $\alpha_t$ be a separable, complete
  pseudo-extended-metric space and $\mathcal F_t$ the Borel
  $\sigma$-algebra generated by its topology. Let $(P_J)_{J\subseteq_f
    \iota}$ be a projective family of finite measures and $P$ be
  defined on $\mathcal A := \bigcup_{J \subseteq_f \iota} \mathcal
  F_J$ given by $P(A) = P_J(A)$ for $A\in\mathcal F_J$. Then, there is
  a unique extension of $P$ to $\sigma(\mathcal A)$.
\end{theorem}

Rather than giving the formalization of this theorem, we give the
definition of the resulting measure (which is the projective
limit). We give the formalized proof at the end of this section, since
we first have to provide a formalization of all tools needed in the
proof.

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def projectiveLimitWithWeakestHypotheses
    [∀ i, PseudoEMetricSpace (α i)]
    [∀ i, BorelSpace (α i)]
    [∀ i, TopologicalSpace.SecondCountableTopology (α i)]
    [∀ i, CompleteSpace (α i)] [Nonempty (∀ i, α i)]
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    [∀ i, IsFiniteMeasure (P i)]
    (hP : IsProjectiveMeasureFamily P) :
    Measure (∀ i, α i)
\end{minted}

We note that we extend the standard assumption that all $\alpha_t$ are
separable, complete metric spaces (or Polish, i.e.\ separable and
metrizable through a complete metric) to cover the case of extended
pseude-metric spaces. Such spaces do not satisfy the frequently used
Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
open balls around $x$ and $y$ overlap.

\subsection{Extending a set function}
In the formulation of Theorem~\ref{T1}, we extend $P$, which is
defined on a union of $\sigma$-algebras. However, unions of
$\sigma$-algebras in general are not $\sigma$-algebras, but they can
be used to define the $\sigma$-algebra generated by the union. So, we
need to extend $P$ to the $\sigma$-algebra generated by $\mathcal
A$. This is exactly what Carathéodory's extension theorem is made
for. In fact, we implemented this result in greater generality than
needed for the proof of the extension theorem. Note that $\mathcal A$
in Theorem~\ref{T1} is a ring of sets (see the next definition)
containing the whole set. We will work with the weaker semi-ring as
introduced next in the formulation of Carathéodory's extension
theorem.

\begin{definition}[Semi-ring, ring]\label{def:semi}
  Let $\alpha$ be some set. We call $\mathcal H \subseteq 2^\alpha$ a
  \emph{semi-ring}, if it is (i) a $\pi$-system (i.e.\ closed under
  $\cap$) and (ii) for all $A, B \in\mathcal H$ there is $\mathcal K
  \subseteq_f \mathcal H$ with\footnote{We write $A\uplus B$ for
    $A\cup B$ if $A\cap B=\emptyset$.}  $B\setminus A = \biguplus_{K
    \in \mathcal K} K$.  \\ We call $\mathcal H \subseteq 2^\alpha$ a
  \emph{ring}, if it is closed under $\cup$ and under set-differences.
\end{definition}

Using \leanline{variable {α : Type _} {C : Set (Set α)} {s t : Set α}
}, we define a semi-ring as follows:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  structure SetSemiring (C : Set (Set α)) : Prop where
    empty_mem : ∅ ∈ C
    inter_mem : ∀ (s) (_ : s ∈ C) (t) (_ : t ∈ C), s ∩ t ∈ C
    diff_eq_Union' :
      ∀ (s) (_ : s ∈ C) (t) (_ : t ∈ C),
      ∃ (I : Finset (Set α)) (_h_ss : ↑I ⊆ C)
      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id),
        t \ s = ⋃₀ I
\end{minted}

Recall that \leanline{Lean4} indicates a coercion by \leanline{↑},
which in this case is from \leanline{Finset} to \leanline{Set}.

\noindent
We do not show the formalization of rings here. Note, however, that
any ring is a semi-ring since $A\cap B = A \setminus (A \setminus B)$,
i.e.\ every ring is a $\pi$-system. In our formlaization, we also use
the notion of a field $\mathcal H$, which is a ring with $\alpha \in
\mathcal H$. Since we a re dealing with product spaces, we will use
cylinders, which are defined as follows:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def cylinder (s : Finset ι) (S : Set (∀ i : s, α i)) :
    Set ((i : ι) → α i) :=
    (fun f : (i : ι) → α i => fun i : s => f i) ⁻¹' S
\end{minted}

In this definition, \leanline{s} is a finite subset of the index set
\leanline{ι}, and for some $S$ in the finite product $\prod_{i \in s}
\alpha_i$, consider the projection $\pi_s : \prod_{i \in \iota}
\alpha_i \to \prod_{i \in s} \alpha_i$, and consider the preimage of
$S$. (This is what the last line in previous definition gives.) From
this, we can defined the set of all measurable cylinders.

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def cylinders : Set (Set ((i : ι) → α i)) :=
    ⋃ (s) (S) (_ : MeasurableSet S), {cylinder s S}
\end{minted}

The set-system of cylinders is in fact a field, hence a ring and hence
a semi-ring:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem setField_cylinders : SetField (cylinders α) :=
    { empty_mem := empty_mem_cylinders α
      univ_mem := univ_mem_cylinders α
      union_mem := union_mem_cylinders
      diff_mem := diff_mem_cylinders }
\end{minted}

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem setRing_cylinders :
    MeasureTheory.SetRing (cylinders α) :=
    setField_cylinders.toSetRing
\end{minted}

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem setSemiringCylinders :
    MeasureTheory.SetSemiring (cylinders α) :=
    setField_cylinders.setSemiring
\end{minted}

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem generateFrom_cylinders :
    MeasurableSpace.generateFrom (cylinders α)
    = MeasurableSpace.pi 
\end{minted}

measurable product space

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  instance MeasurableSpace.pi
    [m : ∀ a, MeasurableSpace (π a)] :
    MeasurableSpace (∀ a, π a) :=
    ⨆ a, (m a).comap fun b => b a
\end{minted}

Let us state an important lemma on semi-rings:

\begin{lemma}\label{l1}
  Let $\mathcal H$ be a semiring, $\mathcal I \subseteq_f \mathcal H$,
  $A \in \mathcal H$. Then, there is $\mathcal K \subseteq_f \mathcal
  H$ such that all sets in $\mathcal K$ are pairwise disjoint and $A
  \setminus \bigcup_{I \in \mathcal I} I = \biguplus_{K\in \mathcal K}
  K$.
\end{lemma}

The formalization of the proof is somewhat straight-forward, but
requires induction over finite sets.

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem exists_disjoint_finset_diff_eq
    (hC : SetSemiring C) (hs : s ∈ C)
    (I : Finset (Set α)) (hI : ↑I ⊆ C) :
    ∃ (J : Finset (Set α)) (_h_ss : ↑J ⊆ C)
    (_h_dis : PairwiseDisjoint (J : Set (Set α)) id),
    s \ ⋃₀ I = ⋃₀ J
\end{minted}

\noindent
Thes existance-statement of the above lemma actually gives rise to a
definition:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def diff₀ (hC : SetSemiring C)
    (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
    [DecidableEq (Set α)] : Finset (Set α) :=
    (hC.exists_disjoint_finset_diff_eq hs I hI).choose \ {∅}
\end{minted}

Complementing Lemma~\ref{l1}, we actually need another result for some
constructions needed in the proofs: For some $A_1,...,A_m$ and $n\neq
m$, we need finite sets $\mathcal K_1,...,\mathcal K_m$ such that $A_n
\setminus \bigcup_{i=1}^{n-1} A_i = \biguplus_{K \in \mathcal K_n} K$
for $n=1,...,m$. Using this, we can write
$$ \bigcup_{n=1}^m A_n = A_1 \uplus (A_2 \setminus A_1) \uplus \cdots
\uplus \Big(A_m \setminus \bigcup_{i=1}^{m-1}A_i\Big) =
\bigcup_{n=1}^m \biguplus_{K \in \mathcal K_n} K,$$ i.e.\ we can write
a union of sets in the semi-ring as a disjoint union of sets in the
semi-ring. The formalized version is as follows:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def indexedDiff₀ (hC : SetSemiring C)
    (J : Finset (Set α)) (hJ : ↑J ⊆ C)
    (n : Fin J.card) : Finset (Set α) :=
    hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
    (finsetLT_subset' J hJ n)
\end{minted}

\begin{proof}
  We proceed by induction on $\mathcal I$. If $\mathcal I$ is a
  singleton, the assertion is true by the definition of a semiring. If
  it holds for some $\mathcal I$ (i.e.\ there is $\mathcal K
  \subseteq_f \mathcal H$ with $A \setminus \bigcup_{I \in \mathcal I}
  I = \biguplus_{K\in \mathcal K} K$), let us consider $\mathcal I' =
  \{J\} \cup \mathcal I$ for some $J \notin \mathcal I$. For each $K
  \in \mathcal K$, Write $K \setminus J = \biguplus_{J_K \in \mathcal
    J_K} J_K$ for some $\mathcal J_K \subseteq_f \mathcal H$ (which
  exists by the definition of a semi-ring). Then, write
  \begin{align*}
    A \setminus \bigcup_{I' \in \mathcal I'} I & = \Big(A \setminus
    \bigcup_{I \in \mathcal I} I\Big) \setminus J = \biguplus_{K\in
      \mathcal K} K \setminus J = \biguplus_{K\in \mathcal K}
    \biguplus_{J_K \in \mathcal J_K} J_K.
  \end{align*}
  This concludes the proof, since the latter disjoint union is over a
  finite set.
\end{proof}

\noindent
\subsection{Carathéodory's extension theorem}
Measures are set functions defined on a $\sigma$-algebra (i.e.\ an
algebra stable under countable unions), satisfying some properties
which we recall below. Mostly, defining such a measure on the full
$\sigma$-algebra is not possible, but defining a set function on a
semi-ring is possible. As an example, consider Lebesgue-measure on
$\mathbb R$, with the Borel $\sigma$-algebra on $\mathbb R$, which is
$\sigma(\mathcal O)$, where $\mathcal O$ is the set of open sets, and
$\sigma(\mathcal O)$ is the smallest $\sigma$-algebra containing all
open sets. Since this set system is defined only abstractly, it is
hard to know which volume each of these sets should be assigned to at
first sight. However, the volume of an interval is easy, since it may
be defined by the length of the interval. So, in order to construct
measures from a set function $m$ on a (semi-)ring $\mathcal H$
(e.g.\ the set of all semi-open intervals), it has been a fundamental
insight of Carathéodory that one may start by defining a set function
(outer measure) $\mu$ on $2^\alpha$, and then show $\mu$ extends $m$
and defines a measure on $\sigma(\mathcal H)$. (Note that the set of
semi-open intervals generates the Borel $\sigma$-algebra on $\mathbb
R$.) We will follow this construction, and start by stating some basic
concepts.

\begin{definition}
  For some set $\alpha$, let $\mathcal H \subseteq 2^\alpha$ and $m :
  \mathcal H \to [0,\infty]$.
  \begin{enumerate}
  \item $m$ is called additive if for $\mathcal K \subseteq_f \mathcal
    H$ pariwise disjoint and $\bigcup_{K \in \mathcal K} K \in
    \mathcal H$, we have $m \Big(\bigcup_{K \in \mathcal K} K \Big) =
    \sum_{K \in \mathcal K} m(K)$. If the same holds for\footnote {We
      write $A \subseteq_c B$ if $A$ is a countable subset of
      $B$.}$\mathcal K \subseteq_c \mathcal H$ pariwise disjoint, we
    say that $m$ is $\sigma$-additive.
  \item The set-function $m$ is called sub-additive if for $\mathcal K
    \subseteq_f \mathcal H$ and $\bigcup_{K \in \mathcal K} K \in
    \mathcal H$, we have $m \Big(\bigcup_{K \in \mathcal K} K \Big)
    \leq \sum_{K \in \mathcal K} m(K)$. (Note that elements of
    $\mathcal K$ need not be disjoint.) Here, $\sigma$-sub-additivity
    is defined in the obvious way using $\mathcal K\subseteq_c
    \mathcal H$.
  \item If $m(A) \leq m(B)$ for $A\subseteq B$ and $A,B\in\mathcal H$,
    we say that $m$ is monotone.
  \item If $\mathcal H$ is a $\sigma$-algebra and $m$ is
    $\sigma$-additive, we call $m$ a measure.
  \item If $\mathcal H = 2^\alpha$ and $m$ is monotone and
    $\sigma$-sub-additive with $m(\emptyset)=0$, we call $m$ an outer
    measure.
  \end{enumerate}
\end{definition}

\noindent
An additive $m : \mathcal H \to [0,\infty]$ is formalized as follows:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  structure AddContent (C : Set (Set α)) where
    toFun : Set α → ℝ≥0∞
    empty' : toFun ∅ = 0
    add' : ∀ (I : Finset (Set α))
      (_h_ss : ↑I ⊆ C) (_h_dis : PairwiseDisjoint
        (I : Set (Set α)) id)
      (_h_mem : ⋃₀ ↑I ∈ C),
      toFun (⋃₀ I) = ∑ u in I, toFun u
\end{minted}

\noindent
In this formalization, we introduced $m(\emptyset)=0$ as a separate
hypothesis with \leanline{empty'}. Note that this also follows
formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
m(\emptyset)$. However, sicne \leanline{Finset}s are assumed to have
distinct elements, we cannot have $\emptyset$ twice in $I$.

% extend_content put $\infty$ outside the set system

For the concrete application we have in mind, we need to be more
specific which semiring to use, based on a projective family of
probability measures. We do this using \leanline{kolContent}, which
is based on the following function, we takes a family of measures and
a cylinder set as input:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def kolmogorovFun
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    (s : Set (∀ i, α i))
    (hs : s ∈ cylinders α) : ℝ≥0∞ :=
    P (cylinders.finset hs) (cylinders.set hs)
\end{minted}

\noindent
With this, we define an \leanline{add_content} on cylinders. (Note
that \leanline{extend_content} in fact defines an additive content on
a semiring.)

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def kolContent
    (hP : IsProjectiveMeasureFamily P) :
    AddContent (cylinders α) :=
    extendContent setSemiringCylinders
    (kolmogorovFun P) (kolmogorovFun_empty hP)
    (kolmogorovFun_additive hP)
\end{minted}

\noindent
From the next lemma, we will need monotonicity and sub-additivity, as
well as $\sigma$-additive $\Rightarrow$ $\sigma$-sub-additive in the
proof of the Carathéodory extension, Theorem~\ref{T:masseind}.

\begin{lemma}[Set-functions on semi-rings]\label{l:halbextRing}
  Let $\mathcal H$ be a semi-ring and $m: \mathcal H\to [0,\infty]$
  additive. Then, $m$ is monotone and sub-additive. In addition, $m$
  is $\sigma$-additive iff it is $\sigma$-sub-additive.
\end{lemma}

\begin{proof}
  We start by monotonicity. Let $A, B\in\mathcal H$ with $A\subseteq
  B$ and $\mathcal K \subseteq_f \mathcal H$ with $B \setminus A =
  \biguplus_{K\in\mathcal K} K$. Therefore, we can write $m(A) \leq
  m(A) + \sum_{K\in\mathcal K} m(K) = m(B)$.

  Next, we claim that for $\biguplus_{I\in\mathcal I} I \subseteq A$
  (and all sets belong to $\mathcal H$, we have $\sum_{I \in \mathcal
    I} m(I) \leq m(A)$. For this, write $A \setminus \biguplus_{I\in
    \mathcal I} I = \biguplus_{K\in\mathcal K} K$ as in
  Lemma~\ref{l1}. Then,
  $$ m(A) = m\Big(\biguplus_{I\in\mathcal I} I \uplus
  \biguplus_{K\in\mathcal K} K\Big) = \sum_{I\in\mathcal I} m(I) +
  \sum_{K\in\mathcal K} m(K) \geq \sum_{I\in\mathcal I} m(I). $$

  For sub-additivity, let $\mathcal I \subseteq_f \mathcal H$ with
  $\bigcup_{I \in \mathcal I} I \in \mathcal H$. Without loss of
  generality, we write $\mathcal I = \{I_1,...,I_n\}$ for some $n$. We
  need to show $m\Big(\bigcup_{k=1}^n I_i\Big) \leq \sum_{k=1}^n
  m(I_k)$. For $k=2,...,n$, we write
  $$ \bigcup_{k=1}^n I_k = \biguplus_{k=1}^n \Big(I_k \setminus
  \bigcup_{j=1}^{k-1} I_j\Big) = \biguplus_{k=1}^n \biguplus_{K_k \in
    \mathcal K_k} K_k$$ with $\mathcal K_k$ as in Lemma~\ref{l1}. So,
  since $\biguplus_{K_k \in \mathcal K_k} K_k \subseteq I_k \in
  \mathcal H$,
  $$ m\Big(\bigcup_{k=1}^n I_i\Big) = \sum_{k=1}^n \sum_{K_k \in
    \mathcal K_k} m(K_k) \leq \sum_{k=1}^n m(I_k).$$

  \noindent
  Now, we show that $m$ is $\sigma$-additive $\iff$ it is
  $\sigma$-sub-additive.

  \noindent
  '$\Rightarrow$': Here, just copy the proof of sub-additivity, but
  using countable $\mathcal I$, i.e.\ $n=\infty$. For '$\Leftarrow$',
  let $\mathcal I \subseteq_c \mathcal H$ and consist of
  disjoint sets with $A = \biguplus_{I \in \mathcal I} I \in\mathcal
  H$. Since $m$ is monotone and for any $\mathcal I' \subseteq_f
  \mathcal I$, we have $\biguplus_{I\in\mathcal I'} I \subseteq A$
  (hence $\sum_{I \in \mathcal I'} m(I') \leq m(A)$),
  $$ \sum_{I \in \mathcal I} m(I) = \sup_{\mathcal I' \subseteq_f
    \mathcal I} \sum_{I \in \mathcal I'} m(I) \leq m(A) \leq \sum_{I
    \in \mathcal I} m(I)$$ by $\sigma$-sub-additivity. So,
  $\sigma$-additivity follows.
\end{proof}

\noindent
Although some material on outer measures was covered in {\tt mathlib4}
already, the classical extension theorem (extending a set function $m$
on a semiring $\mathcal H$)) was not provided yet. In particular, this
result states that the outer measure coincides with $m$ on $\mathcal
H$. All statements on equality of $\mu$ and $m$ present in {\tt
  mathlib4} at the time of writing have too many requirements.

The next result talks about an \leanline{induced_outer_measure} in
terms of {\tt mathlib4}.

Theorem 2.5

\begin{proposition}[Outer measure induced by a set function on a semi-ring] %\mbox{}
  Let \label{P:auss} $\mathcal H$ be a semi-ring and $m: \mathcal
  H\to\mathbb R_+$ additiv. For $A\subseteq E$ let
  $$ \mu(A) := \inf_{\mathcal G \in \mathcal U(A)}
    \sum_{G\in\mathcal G} m(G)$$ where
  $$ \mathcal U(A) := \big\{\mathcal G \subseteq_c \mathcal H,
    A\subseteq \bigcup_{G\in\mathcal G} G\big\}$$ is the set of
    countable coverings of $A$. Then, $\mu$ is an outer measure.
\end{proposition}

\sloppy Let us now formulate the classical results by
Carathéodory. The first leads to the definition of the measurable
space \leanline{OuterMeasure.caratheodory}, covered in {\tt
  mathlib4}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}

\begin{theorem}[\boldmath $\mu$-measurable sets are a
    $\sigma$-algebra]\label{T:cara} Let $\mu$ be an outer measure on
  $E$ and $\mathcal F$ the set of $\mu$-measurable sets,
  i.e.\ the set of sets $A$ satisfying
  \begin{align}
    \label{eq:mumess}
    \mu(B) = \mu(B\cap A) + \mu(B\cap A^c), \qquad B \subseteq E.
  \end{align}
  Then, $\mathcal F$ is a $\sigma$-Algebra and $\mu|_{\mathcal F}$ is
  a measure. In addition, $\mathcal N:=\{N\subseteq \Omega:
  \mu(N)=0\}\subseteq \mathcal F$, i.e.\ $\mathcal F$ is complete.
\end{theorem}

\noindent
\sloppy The second result, is not yet covered in {\tt
  mathlib4}. However, for product spaces, in
\leanline{measureTheory.Constructions.Pi}, there is
\leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
\subseteq \mathcal F$ in a special case. In addition, there is
\leanline{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
H$ in the same special case. See also \cite[Theorem
  2.5]{Kallenberg2020}.

\begin{theorem}[Carathéodory extension]\label{T:masseind}
  Let $\mathcal H$ be a semiring and $m: \mathcal H\to\mathbb R_+$
  $\sigma$-finite and $\sigma$-additive. Furthermore, let $\mu$ be the
  induced outer measure from Proposition~\ref{P:auss} and $\mathcal F$
  the $\sigma$-algebra from Theorem~\ref{T:cara}. Then,
  $\sigma(\mathcal H)\subseteq\mathcal F$ and $\mu$ coincides with $m$
  on $\mathcal H$.
\end{theorem}

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def Measure.ofAddSubaddCaratheodory
    (hC : SetSemiring C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (m_empty : m ∅ hC.empty_mem = 0)
    (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
      (h_mem : ⋃₀ ↑I ∈ C), m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
      @Measure α (inducedOuterMeasure m
        hC.empty_mem m_empty).caratheodory
\end{minted}

\noindent
Here, \leanline{MeasureTheory.OuterMeasure.caratheodory} is already
implemented and gives the assertion that $\sigma(\mathcal
H)\subseteq\mathcal F$.

\noindent
In the proof of Kolmogorov's extension theorem, we use the following
definition, which is based on
\leanline{Measure.ofAddSubaddCaratheodory}.

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def Measure.ofAddContent
    [mα : MeasurableSpace α] (hC : SetSemiring C)
    (hC_gen : MeasurableSpace.generateFrom C = mα)
    (m : AddContent C)
    (m_sigma_subadd :
      ∀ ⦃f : ℕ → Set α⦄ (_hf : ∀ i, f i ∈ C) (_hf_Union : (⋃ i, f i) ∈ C),
        m (⋃ i, f i) ≤ ∑' i, m (f i)) :
        Measure α
\end{minted}

\begin{proof}
  First, note that $m$ is $\sigma$-sub-additive by
  Lemma~\ref{l:halbextRing}.

  ~

  \noindent\emph{Step 1: $\mu|_{\mathcal H} = m$:} Let $A\in\mathcal
  H$. Choose $\mathcal K \subseteq_c \mathcal H$ with
  $A\subseteq \bigcup_{K\in\mathcal K} K$ and
  $$\mu(A) \geq \sum_{K\in\mathcal K} m(K) - \varepsilon.$$ By $ A =
  \bigcup_{K\in\mathcal K} K\cap A$,
  \begin{align*}
    \mu(A) \leq m(A) \leq \sum_{K\in\mathcal K} m(K\cap A) \leq
    \sum_{K\in\mathcal K} m(K) \leq \mu(A) + \varepsilon,
  \end{align*}
  where we have used $\sigma$-sub-additivity of $m$ in the second and
  monotonicity of $m$ in the third $\leq$ (see
  Lemma~\ref{l:halbextRing}).  With $\varepsilon\to 0$ we find that
  $\mu(A) = m(A)$.

  ~

  \noindent\emph{Step 2: $\sigma(\mathcal H)\subseteq \mathcal F$:}
  Let $B \subseteq E$, $A\in\mathcal H$ and $\varepsilon>0$. Choose
  $\mathcal K \subseteq_c \mathcal H$ such that $B \subseteq
  \bigcup_{K \in \mathcal K} K$ and $\mu(B) \geq \sum_{K\in\mathcal K}
  m(K) - \varepsilon.$ Then, by additivity of $m$,
  \begin{align*}
    \mu(B)+\varepsilon & \geq \sum_{K\in\mathcal K} \mu(K) =
    \sum_{K\in\mathcal K} \mu(K\cap A) + \sum_{K\in\mathcal K}
    \mu(K\cap A^c) \\ & \geq \mu(B\cap A) + \mu(B\cap A^c).
  \end{align*}
  By sub-additivity of $\mu$, we find that $\mu(B) \leq \mu(B\cap A) +
  \mu(B \cap A)$, so letting $\varepsilon\to 0$ leads to $\mu(A) =
  \mu(E\cap A) + \mu(E\cap A^c)$. This implies that $A$ is $\mathcal
  F$-measurable, and we have shown $\sigma(\mathcal H) \subseteq
  \mathcal F$, since $\mathcal F$ is a $\sigma$-algebra.

\end{proof}

\subsection{$\sigma$-additivity of set functions}
For the proof of Theorem~\ref{T1}, note that $P$ as given in the
result is a finite additive set function on the ring $\mathcal A$. In
order to use Theorem~\ref{T:masseind}, we therefore have to show
$\sigma$-additivity. For this, we will use Lemma~\ref{l:stetigcompact}
below, i.e.\ inner regularity of $P$ with respect to compact
sets. Before, we will show useful alternative conditions for
$\sigma$-additivity, which do not make use of any topological
structure of the underlying space.

\begin{lemma}[$\sigma$-additivity and continuity at $\emptyset$]\label{P:stetigmass}
  Let $\mathcal R$ be a ring and $m:\mathcal R\to \mathbb R_+$
  additive. Then, the following are equivalent:
  \begin{enumerate}
    \item $m$ is $\sigma$-additive;
    \item $m$ is $\sigma$-sub-additive;
    \item $m$ is continuous from below, i.e.\ for $A, A_1, A_2,... \in
      \mathcal R$ with $A_1 \subseteq A_2 \subseteq \cdots$, we have
      $m(A) = \lim_{n\to\infty} m(A_n)$.
    \item $m$ is continuous from above in $\emptyset$, i.e.\ for $A_1,
      A_2,...\in\mathcal R$ with $A_1 \supseteq A_2 \supseteq ...$ and
      $\bigcap_{n=1}^\infty A_n = \emptyset$, we have
      $\lim_{n\to\infty} m(A_n)=0$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  1.$\Leftrightarrow$2.\ was already shown in
  Lemma~\ref{l:halbextRing}, since $\mathcal R$ is a semiring.

  ~

  \noindent 1.$\Rightarrow$3.: For $A_1, A_2,...$ as in 3., we write
  (with $A_0 := \emptyset$)
  $$ m(A) = m\Big( \biguplus_{n=1}^\infty A_n \setminus
  \Big(\bigcup_{l=1}^{k-1} A_l\Big)\Big) = \lim_{n\to\infty} m\Big(
  \biguplus_{k=1}^n A_k \setminus \Big(\bigcup_{l=1}^{k-1}
  A_l\Big)\Big) = \lim_{n\to\infty} m(A_n).$$

  ~

  \noindent 3.$\Rightarrow$1.: Let $A_1, A_2,... \in \mathcal R$ be
  disjoint and $A = \biguplus_{n=1}^\infty A_n$. Set $B_n :=
  \bigcup_{k=1}^n A_k$, such that $B_1, B_2,...$ satisfy 3. Hence,
  $$ m\Big( \biguplus_{n=1}^\infty A_n\Big) = m\Big(
  \bigcup_{n=1}^\infty B_n\Big) = \lim_{n\to\infty} m\Big(
  \bigcup_{k=1}^n B_n\Big) = \lim_{n\to\infty} m\Big(
  \biguplus_{k=1}^n A_n\Big) = \sum_{n=1}^\infty m(A_n).$$

  \noindent 3.$\Rightarrow$4.: Let $A_1, A_2,\dots \in\mathcal R$ as
  in 4. Set $B_n := A_1\setminus A_n$. Then, $B=A_1,B_1,B_2,\dots
  \in\mathcal R$ satisfy 3., and therefore
  $$m(A_1) = \lim_{n\to\infty} m(B_n) = m(A_1) -
  \lim_{n\to\infty}m(A_n),$$ and 4.\ follows.

  ~

  \noindent 4.$\Rightarrow$3.: Let $A,A_1,A_2,\dots \in\mathcal R$ as
  in 3. Let $B_n := A\setminus A_n \in \mathcal R, n\in\mathbb
  N$. Then, $\bigcap_{n=1}^\infty B_n = \emptyset$, so $$0 =
  \lim_{n\to\infty} \mu(B_n) = \mu(A) - \lim_{n\to\infty} \mu(A_n),$$
  and 3.\ follows.
\end{proof}

\noindent
As an example, we give the formalization of $4.\Rightarrow 1.$:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem countably_additive_of_todo (hC : SetRing C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (hm_ne_top : ∀ {s} (hs : s ∈ C), m s hs ≠ ∞)
    (hm_add : ∀ {s t : Set α} (hs : s ∈ C) (ht : t ∈ C),
      Disjoint s t →
      m (s ∪ t) (hC.union_mem hs ht) = m s hs + m t ht)
    (hm : ∀ ⦃s : ℕ → Set α⦄ (hs : ∀ n, s n ∈ C),
      Antitone s → (⋂ n, s n) = ∅ →
      Tendsto (fun n => m (s n) (hs n)) atTop (𝓝 0))
      ⦃f : ℕ → Set α⦄ (h : ∀ i, f i ∈ C) (hUf : (⋃ i, f i) ∈ C)
      (h_disj : Pairwise (Disjoint on f)) :
    m (⋃ i, f i) hUf = ∑' i, m (f i) (h i)
\end{minted}

\noindent
Next, we will be extending our analysis to the case of a topological
space. We define inner (and outer) regularity of set functions.

\begin{definition}[Inner regularity] \label{def:innerreg}
  Let $\alpha$ be some set, equipped with a topology, and $m$ be a
  set-function on some $\mathcal H \subseteq 2^\alpha$.
  \begin{enumerate}
  \item Let $p, q : 2^\alpha \to \{\text{true, false}\}$. Then, $m$ is
    called inner regular with respect to $p$ and $q$, if 
    $$ m(A) = \sup\{m(F) : p(F) = \text{true}, F \subseteq A\}$$ for
    all $A \in \mathcal H$ with $q(A) = \text{true}$.
  \item If $q(A) = \text{true}$ iff $A$ is measurable, we neglect the
    {\em and $q$}. If $p(A) = \text{true}$ iff $A$ is closed (ompact,
    closed and compact), we say that $m$ is inner regular with respect
    to closed (compact, compact and closed) sets.
  \end{enumerate}
\end{definition}

The above definition closely resembles its formalization in
\leanline{mathlib4}:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def MeasureTheory.Measure.InnerRegular
    {α : Type u_1}  {_ : MeasurableSpace α}
    (μ : Measure α) (p q : Set α → Prop) :=
    ∀ ⦃U⦄, q U → ∀ r < μ U, ∃ K, K ⊆ U ∧ p K ∧ r < μ K
\end{minted}

For the next result, recall that for compact sets $C_1, C_2,...$ with
$\bigcap_{n=1}^\infty C_n = \emptyset$, there is some $N$ with
$\bigcap_{n=1}^N C_n = \emptyset$.

\begin{lemma}\label{l:stetigcompact}
  Let $\alpha$ be a topological space and $\mu$ be an additive set
  function on a ring $\mathcal R$ contained by the Borel
  $\sigma$-algebra, and which is inner regular with respect to compact
  sets. Then, $\mu$ is $\sigma$-additive.
\end{lemma}

\begin{proof}
  According to Lemma~\ref{P:stetigmass}, we need to show continuity of
  $\mu$ in $\emptyset$, so let $A_1 \supseteq A_2 \supseteq... \in
  \mathcal R$ satisfy $\bigcap_{n=1}^\infty = \emptyset$ and
  $\varepsilon>0$. Let $\delta_1, \delta_2,..>0$ with
  $\sum_{n=1}^\infty \delta_n < \varepsilon$. For each $n$, let $C_n
  \subseteq A_n \in \mathcal R$ be compact with $\mu(A_n) \leq
  \mu(C_n) + \delta_n$. We have that $\bigcap_{n=1}^\infty C_n
  \subseteq \bigcap_{n=1}^\infty A_n = \emptyset$, so there is $N \in
  \mathbb N$ with $\bigcap_{n=1}^N C_n = \emptyset$. So, for any
  $m>N$, we have that
  $$ \mu(A_m) = \mu\Big( \Big(\bigcap_{n=1}^m A_n\Big) \setminus
  \Big(\bigcap_{n=1}^m C_n\Big)\Big) \leq \sum_{n=1}^m \mu(A_n
  \setminus C_n) \leq \sum_{n=1}^m \delta_n < \varepsilon.$$ This
  concludes the proof.
\end{proof}

Here is the formalization using a \leanline{kol_content}:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem kolContent_sigma_additive_of_innerRegular
    (hP : IsProjectiveMeasureFamily P)
    (hP_inner : ∀ J,
      (P J).InnerRegular (fun s => IsCompact s ∧ IsClosed s)
      MeasurableSet)
    ⦃f : ℕ → Set (∀ i, α i)⦄ (hf : ∀ i, f i ∈ cylinders α)
    (hf_Union : (⋃ i, f i) ∈ cylinders α)
    (h_disj : Pairwise (Disjoint on f)) :
    kolContent hP (⋃ i, f i) = ∑' i, kolContent hP (f i)
\end{minted}

\sloppy Using this, we also have
\leanline{kolContent_countably_subadditive_of_innerRegular} by using
\leanline{countably_subadditive_of_countably_additive}.

Recall that a finite measure on the Borel $\sigma$-algebra is inner
regular with respect to closed sets, but also outer regular with
respect to open sets.

\begin{lemma}\label{l:pseude1}
  Let $r$ be an extended pseudo-metric on $\alpha$, and the topology
  on $\alpha$ be given by $r$. If $\mu$ is a finite measure on the
  Borel $\sigma$-algebra $\mathcal B$ on $\alpha$, it is inner regular
  with respect to closed sets. In fact, we have also outer regularity
  with respect to open sets, i.e.
  $$ \mu(A) = \inf \{ \mu(O) : A \subseteq O \text{ open}\}.$$
\end{lemma}

\begin{proof}
  It suffices to prove that
  $$ \mathcal A := \Big\{A \in \mathcal B : \mu(B) = \sup_{F\subseteq
    B \text{ closed}} \mu(F) = \sup_{B\subseteq O \text{ open}}
  \mu(O)\Big\} $$ is a $\sigma$-algebra containing all closed
  sets. So, we proceed in two steps.  \\ {\em Step 1: $\mathcal A$ is
    a $\sigma$-algebra:} Since $\mu$ is finite, we have that $\mu(A^c)
  = \mu(\alpha) - \mu(A)$ for all $A$. Using this, we already have
  that $\mathcal A$ is closed under complements. So, we are left with
  showing that $\mathcal A$ is closed under countable unions. For
  this, let $A_1, A_2,... \in \mathcal A$ and $A :=
  \bigcup_{n=1}^\infty A_n$. Fix $\varepsilon > 0$ and a sequence
  $\delta_1, \delta_2,... > 0$ with $\sum_{n=1}^\infty \delta_n <
  \varepsilon$. For each $n$, let $F_n \subseteq A_n \subseteq O_n$
  with $F_n$ closed, $O_n$ open and $\mu(O_n \setminus F_n) \leq
  \delta_n$. Then, $O := \bigcup_{n=1}^\infty O_n$ is open and
  $$ 0 \leq \mu(O \setminus A) = \mu(O) - \mu(A) \leq
  \sum_{n=1}^\infty \mu(O_n \setminus A) \leq \sum_{n=1}^\infty
  \mu(O_n \setminus A_n) < \varepsilon.$$ This shows outer regularity
  of $\mu$ with respect to open sets at $A$. It remains to show inner
  regularity with respect to closed sets. For this, note that
  $\bigcup_{n=1}^N F_n$ is closed for all $N$ and recall from
  Proposition~\ref{P:stetigmass} that $\mu$ is continuous from
  below. Therefore, setting $F := \bigcup_{n=1}^\infty F_n$,
  $$ 0 \leq \mu(A) - \lim_{N\to\infty} \mu\Big(\bigcup_{n=1}^N
  F_n\Big) = \mu(A \setminus F) \leq \sum_{n=1}^\infty \mu(A_n
  \setminus F) \leq \sum_{n=1}^\infty \mu(A_n \setminus F_n) <
  \varepsilon.$$ This shows inner regularity of $\mu$ with respect to
  closed sets at $A$.

  \noindent
      {\em Step 2: $\mathcal A$ contains all closed sets:} Let
  $\varepsilon_ n\downarrow 0$ and, for $\varepsilon>0$ and some $A
  \subseteq \alpha$, $A^\varepsilon := \{x : \exists y \in A,
  r(x,y<\varepsilon\}$. Then, for $A$ closed, we have that $A =
  \bigcap_{n=1}^\infty A^{\varepsilon_n}$ and $A^{\varepsilon_n}$ is
  open for all $n$. Clearly, $\mu(A) = \sup\{\mu(F): F\text{ closed},
  F \subseteq A\}$. By continuity of $\mu$, we find that
  $\mu(A^{\varepsilon_n}) \xrightarrow{n\to\infty} \mu(A)$, therefore
  $\mu(A) = \inf\{\mu(O): F\text{ open}, A \subseteq O\}$, and we have
  shown that $A \in \mathcal A$.
\end{proof}

For the formalization, we use a new result: \leanline{P.InnerRegular
  (fun s => IsCompact s ∧ IsClosed s) IsClosed}, which holds for every
second countable extended pseudo metric space, which is
\leanline{OpensMeasurable}. Then, from \leanline{mathlib4}, we use
\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
that every open set is inner regular with respect to closed sets, and
\leanline{Measure.InnerRegular.measurableSet_of_open} to show the
above lemma.

%xxx From Remy: For \leanline{inner_regular_etc} we already do it in
%three parts: our result shows \leanline{P.inner_regular (λ s,
%is_compact s ∧ is_closed s) is_closed}, then we go from closed to
%open sets thanks to
%\leanline{docs3#measure_theory.measure.inner_regular.of_pseudo_emetric_space},
%and finally to measurable sets with
%\leanline{docs3#measure_theory.measure.inner_regular.measurable_set_of_open}. That
%last step uses the borel space assumption in the typeclass inference
%process to get its \leanline{outer_regular} argument.

In the next lemma, we will need that a closed subset of a compact set
is compact. (See \leanline{isCompact_of_isClosed_subset}.)

\begin{lemma}\label{l:tight}
  Let $r$ be an extended pseudo-metric on $\alpha$, and the topology
  on $\alpha$ be given by $r$. If $\mu$ is a finite measure on the
  Borel $\sigma$-algebra $\mathcal B$ on $\alpha$, the following are
  equivalent:
  \begin{enumerate}
    \item For all $\varepsilon>0$, there is some closed and compact
      $K$ with $\mu(K^c) < \varepsilon$.
    \item For all $\varepsilon>0$ and $A \in \mathcal B$, there is
      some closed and compact $K\subseteq A$ with $\mu(A \setminus K)
      < \varepsilon$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  2. $\Rightarrow$ 1.: This is clear since $\alpha \in \mathcal B$.
  1. $\Rightarrow$ 2.: Fix $\varepsilon>0$ and $A \in \mathcal B$. Let
  $K$ be closed and compact with $\mu(K^c) < \varepsilon/2$. By
  Lemma~\ref{l:pseude1}, there is some closed $F \subseteq A$
  with $\mu(A) < \mu(F) + \varepsilon/2$. Now, we have that $F \cap K
  \subseteq A$ is closed and compact, and with $A \setminus (F\cap K)
  = (A \setminus F) \cup (A \setminus K)$,
  $$\mu(A \setminus (F\cap K)) \leq \mu(A \setminus F) + \mu(A
  \setminus K) < \varepsilon/2 + \mu(K^c) < \varepsilon. $$
\end{proof}

We apply this result by using that 1.\ is satisfied for complete and
separable extended pseudo-metric spaces:

\begin{lemma}\label{L:relcoPol}
  Let $\alpha$ be a complete and separable, extended pseudo-metric
  space, and $\mu$ a finite measure on its Borel $\sigma$-algebra
  $\mathcal B$. Then, for any $\varepsilon>0$, there is some
  $K\subseteq \alpha$ with compact closure and\footnote{We write $\bar
    K$ for the closure of $K$.}  $\mu((\bar K)^c)<\varepsilon$.
\end{lemma}

\begin{proof}
 Since $\alpha$ is separable, there is a subset $\{x_1, x_2,...\}
 \subseteq_c \alpha$ with closure $\alpha$. By separability, for each
 $n\in \mathbb N$, $\alpha = \bigcup_{k=1}^\infty
 B_{\varepsilon_n}(x_k)$. Since $\mu$ is continuous from above in
 $\emptyset$ (see Lemma~\ref{P:stetigmass})
  \[
  0 = \mu\Big(\Big(\bigcup_{k=1}^\infty B_{\varepsilon_n}(x_k)\Big)^c
  \Big) = \lim_{N\to\infty} \mu\Big( \Big(\bigcup_{k=1}^N
  B_{\varepsilon_n}(x_k)\Big)^c\Big).
  \]
  Let $\delta_n\downarrow 0$ be summable with $\sum_{n=1}^\infty
  \delta_n = 1$ (e.g.\ $\delta_n = 2^{-n}$). Then, there is some
  $N_n\in\mathbb N$ with $\mu\Big( E\setminus \bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big) < \varepsilon \delta_n$. Now take
  \[
    A := \bigcap_{n=1}^\infty \bigcup_{k=1}^{N_n}
    B_{\varepsilon_n}(x_k),
  \]
  which by construction is totally bounded (for $r>0$ take $n$ such
  that $\varepsilon_n < r$, such that $B_r(x_1),...,B_r(x_{N_n})$
  cover $A$), and therefore has a compact closure as noted before the
  lemma. In addition, by $\sigma$-sub-additivity of $\mu$,
  $$ \mu((\overline A)^c) \leq \mu(A^c) = \mu\Big(
  \bigcup_{n=1}^\infty \Big(\Big(\bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big)^c\Big)\Big) \leq \sum_{n=1}^\infty
  \mu\Big( \Big(\bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon,$$ and we are done.
\end{proof}

Let us combine the last two results in one combined formalized result:

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem innerRegular_isCompact_isClosed_
    measurableSet_of_complete_countable
    [PseudoEMetricSpace α] [CompleteSpace α]
    [TopologicalSpace.SecondCountableTopology α]
    [BorelSpace α] (P : Measure α) [IsFiniteMeasure P] :
    P.InnerRegular (fun s => IsCompact s ∧ IsClosed s)
      MeasurableSet 
\end{minted}

\noindent
With the above results, the strategy of the proof of Theorem~\ref{T1}
is as follows: We have to show that under the given conditions,
condition~1 of Lemma~\ref{l:tight} holds. By the equivalent~2., we
find that $P$ is inner regular with respect to closed compact sets. In
particular, the conditions of Lemma~\ref{l:stetigcompact} are
satisfied, showing that $P$ is $\sigma$-additive. Then, we proceed as
described right before Proposition~\ref{P:stetigmass}. So, it remains
to show condition~1 of Lemma~\ref{l:tight} for complete and separable
extended pseudo-metric spaces. Note that -- as a special case of
\leanline{isCompact_iff_totallyBounded_isComplete} from {\tt
  mathlib4}, any subset of such a space is compact iff it is complete
and totally bounded. Since closed subsets of a complete space are
complete, the closure of a totally bounded set is (still totally
bounded, hence) compact.

%\begin{lemma}\label{l:totalBoundcompact}
%  Let $E$ be a uniform topologocal space\footnote{This is more general
%    than a metric space, i.e.\ every metric space is uniform.} Then,
%  $A\subseteq E$ is compact iff it is totally bounded\footnote{In a
%    metric space, this means that forall $\varepsilon>0$ you find a
%    finite number of balls covering the set} and
%  complete\footnote{i.e.,\ Cauchy filters converge.}.
%\end{lemma}

%\noindent
%The usual way to apply this result is in a Polish space. Here, $A$ is
%relatively compact iff it is totally bounded.

%\noindent
%Although, regularity of measures is covered in {\tt mathlib}, the next
%statement is not yet part of the library. Similar constructions will
%be important for Prohorov's theorem in the theory of weak convergence.
%In Lean, we can use \verb+ measure_theory.measure.inner_regular+ for
%the statement of this result. In the proof of the lemma, we will be
%using the last fact

\subsection{Proof of Kolmogorov's extension theorem}
Now we can put everything together: According to
Lemma~\ref{L:relcoPol}, condition~1 of Lemma~\ref{l:tight} is
satisfied for the set function $P$ on the ring $\mathcal A$ from
Theorem~\ref{T1}. The same lemma shows that the conditions for
Lemma~\ref{l:stetigcompact} are satisfied, so $P$ is $\sigma$-additive
on $\mathcal A$. As a consequence, we can use the Carathéodory
extension from Theorem~\ref{T:masseind}. This shows all other
assertions. The formalization of this proof resembles these
arguments. We leave out all instances in the reformulation of the
result and its proof (see below Theorem~\ref{T1} for a full
formulation):

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  noncomputable def projectiveLimitWithWeakestHypotheses
    [∀ i, PseudoEMetricSpace (α i)]
    [∀ i, BorelSpace (α i)]
    [∀ i, TopologicalSpace.SecondCountableTopology (α i)]
    [∀ i, CompleteSpace (α i)] [Nonempty (∀ i, α i)]
    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
    [∀ i, IsFiniteMeasure (P i)]
    (hP : IsProjectiveMeasureFamily P) :
    Measure (∀ i, α i) :=
      Measure.ofAddContent setSemiringCylinders
      generateFrom_cylinders (kolContent hP)
      (kolContent_countably_subadditive_of_innerRegular hP
        fun J => innerRegular_isCompact_isClosed_
        measurableSet_of_complete_countable (P J))    
\end{minted}

\section{Concluding thoughts}
{\em Using semi-rings in measure theory:} Frequently, the construction
of measures on some $\sigma$-algebra $\mathcal F$ (on some space
$\alpha$) uses outer measures, which can be defined on $2^\alpha$ (the
set of all subsets of $\alpha$). We provide a general framework using
Carathéodory's extension theorem, which states an extension of a
set-function on a semi-ring by an outer measure, not previously
implemented in mathlib4. This could also be used to re-define
Stieltjes measures (in particular Lebesgue measure) on $\mathbb R$, as
well as product measures (on finite products). More precisely,
e.g.\ \leanline{StieltjesFunction.outer_Ioc} and
\leanline{MeasureTheory.constructions.pi.pi_pi_aux} are specific
examples of the more general statement, that the outer measure extends
the set function a the semi-ring. (In the first case, it would be the
semi-ring of half-open intervals, in the second case, it would be the
semi-ring of cylinder sets on a product space.)

\noindent
{\em Prohorovs theorem:} The proof that single finite measures on a Polish space is tight (i.e.\ inner regular with respect to compact sets) is a special case of Prohorov's theorem, which states that a set $\mathcal M$ of finite measures is relatively compact (in the weak topology on the set of finite measures, i.e.\ every sequence of measures in the set has a weakly convergent sub-sequence) if and only if $\mathcal M$ is tight (i.e.\ for all $\varepsilon>0$ there is some compact $K$ such that $\sup_{m \in \mathcal M} m(K^c) < \varepsilon$). We treat here the special case of $\mathcal M$ being a singleton (hence is a compact set).

xxx design decisions?

\section*{Acknowledgements}

\bibliographystyle{plain}
\bibliography{main}

\end{document}

Our formalization broadly follows the procedure described in \cite{hytonen2016analysis} and leads to a definition with the following type.
\begin{lstlisting}
def condexp (ℬ : measurable_space Ω)
  { : measurable_space Ω}
  (μ : measure Ω) (f : Ω → E) : Ω → E :=
if hm : ℬ ≤ 
  then if h : sigma_finite (μ.trim hm) ∧ integrable f μ
    then if strongly_measurable[ℬ] f then f
      else (@ae_strongly_measurable'_condexp_L1 _ _ _ _ _ ℬ  μ hm h.1 _).mk (@condexp_L1 _ _ _ _ _ _ _ hm μ h.1 f)
    else 0
  else 0
\end{lstlisting}
\begin{definition}\label{def:condexp}
Let $(\Omega, \mathscr A, \mu)$ be a measure space and let $\mathscr B$ be a sub-$\sigma$-algebra of $\mathscr A$. Let $E$ be a Banach space (complete normed vector space) and $f : \Omega \to E$ be an i\
ntegrable function (with respect to $\mu$). Then a function $g : \Omega \to E$ is said to be a \emph{conditional expectation} of $f$ with respect to $\mathscr B$ if it is integrable, $\mathscr B$-strong\
ly measurable and for all $\mathscr B$-measurable sets $s$,
\begin{align*}
\int_{x \in s} g(x) d\mu(x) = \int_{x \in s} f(x) d\mu(x) \: .
\end{align*}
\end{definition}

%\documentclass[sigplan,10pt,anonymous,review]{acmart}
\documentclass[sigplan,10pt,screen]{acmart}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{enumitem}
\setlist[enumerate]{%
align=left, itemsep=0pt, leftmargin=0pt, labelindent=0pt, listparindent=0pt, labelwidth=0pt, itemindent=!
}

%\theoremstyle{plain}
%
\newtheorem{proposition}{Proposition}[section]
%\newtheorem{theorem}[proposition]{Theorem}
\newtheorem{lemma}[proposition]{Lemma}
\newtheorem{corollary}[proposition]{Korollar}
\newtheorem{definition}[proposition]{Definition}
\newtheorem{remark}[proposition]{Remark}

\pdfoutput=1

% Definitions and packages for Lean code
\usepackage{color}
\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green

\usepackage{listings}
\def\lstlanguagefiles{lstlean.tex}
\lstset{language=lean, basicstyle=\small}

% Other packages
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}

\input{quiver.tex}

\documentclass{article}
\usepackage[utf8x]{inputenc}

\usepackage{hyperref}
\usepackage{refcheck}
\usepackage[centertags]{amsmath} 
\usepackage{amssymb, amsthm}

\usepackage{enumitem}
\setlist[enumerate]{%
align=left, itemsep=0pt, leftmargin=0pt, labelindent=0pt, listparindent=0pt, labelwidth=0pt, itemindent=!
}

\theoremstyle{plain}
%
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}[proposition]{Theorem}
\newtheorem{lemma}[proposition]{Lemma}
\newtheorem{corollary}[proposition]{Korollar}
\newtheorem{definition}[proposition]{Definition}
\newtheorem{remark}[proposition]{Remark}

\title{The Kolmogorov Extension Theorem in Lean} \author{Rémy Degenne,
  Peter Pfaffelhuber}

\begin{document}

\maketitle

The goal of the project is to state and prove the Kolmogorov extension
theorem in {\tt Lean 4} and make a PR to {\tt mathlib4}.

\input{formalization}

\bibliographystyle{alpha}
\bibliography{main}

\end{document}

\begin{thebibliography}{EPW06b}
\bibitem[Bil95]{Billingsley1995} P.~Billingsley.  \newblock {\em
  {Probability and Measure. 3rd ed.}}  \newblock {Wiley Series in
  Probability and Statistics.}, 1995.
\bibitem[Imm12]{Immler2012} F.~Immler.  \newblock {\em {Generic
    Construction of Probability Spaces for Paths of Stochastic
    Processes in Isabelle/HOL.}} \newblock {Master’s Thesis in
  Informatik, TU Munich},
  \url{https://saloranta.de/immler/fabian/mastersthesis/thesis.pdf},
  2012.
\bibitem[Kal20]{Kallenberg2020} O.~Kallenberg.  \newblock {\em
  {Foundations of modern probability. 3rd ed.}}  \newblock
  {Probability and Its Applications. New York, NY: Springer.}, 2020.
\bibitem[Kle13]{Klenke2013} A.~Klenke.  \newblock {\em {Probability
    Theory. A Comprehensive Course. 2nd ed.}}  \newblock {Springer.},
  2013.
\end{thebibliography}

### 2023-09-08 Manuscript/proof_outline.tex
updated manuscript

@@
 
-\usepackage{environ}
-\NewEnviron{killcontents}{}
-\let\proof\killcontents
-\let\endproof\endkillcontents
+%\usepackage{environ}
+%\NewEnviron{killcontents}{}
+%\let\proof\killcontents
+%\let\endproof\endkillcontents
 
@@
 \begin{abstract}
-    xxx
+  We present a formalization of the Kolmogorov extension theorem,
+  which is a main building block in construction the (distribution of)
+  stochastic processes with arbitrary index sets in Lean4. Our
+  approach is based on mathlib, the mathematical library for Lean4. On
+  our way, we provide a formalization of the Caratheodory extension
+  theorem, which allows to extend a set function on a semi-ring to a
+  proper measure on the $\sigma$-algebra generated by the semi-ring. A
+  slight generalization of the classical Kolmogorov extension theorem
+  allows us to construct the (distribution of) a stochastic process on
+  a complete and separable pseudo-metric space (rather than a metric
+  space).
 \end{abstract}
@@
   \end{itemize}
-
 }
@@
 \section{Introduction}
-One of the main building blocks of modern probability theory are stochastic processes, which are usually defined as any collection of random variables -- $(X_t)_{t\in \iota}$ with $X_t$ taking values in some $\alpha_t$, say -- defined on some joint probability space $(\Omega, \mathcal A, \mathbf P)$.
-In order to study such processes, it is fundamental to talk about their distribution, i.e.\ a probability distribution on the product set $\prod_{t\in\iota} \alpha_t$.
-The usual approach to construct (the distribution of) a stochastic process works as follows: describe properties of the distribution of the stochastic process $P_J$ at some finite number of times $J = \{t_1,...,t_n\} \subseteq \iota$.
+One of the main building blocks of modern probability theory are stochastic processes, which are usually defined as any collection of random variables -- $(X_t)_{t\in \iota}$ with $X_t$ taking values in some $\alpha_t$ for all $t\in\iota$ -- defined on some joint probability space $(\Omega, \mathcal A, \mathbf P)$.
+(As usual, we will refer to $\iota$ as the index set of times.) In order to study such processes, it is fundamental to talk about their joint distribution, i.e.\ a probability distribution on the product set $\prod_{t\in\iota} \alpha_t$.
+The usual approach to construct (the distribution of) a stochastic process works as follows: describe properties of the distribution of the stochastic process $P_J$ at some arbitrary but finite number of times $J = \{t_1,...,t_n\} \subseteq \iota$.
 The resulting family of probability measures $(P_J)_{J \subseteq \iota \text{ finite}}$ has to be {\em projective} in the sense that the projection of $P_J$ to $H\subseteq J$ has to be equal to $P_H$.
@@
 $\iota = [0,\infty)$, one is tempted to be pessimistic at first sight since measures -- which describe the distribution of the stochastic process -- usually only deal well with a countable number of measurable events.
-However, it is the achievement of Kolmogorov that the finite-dimensional distributions in fact provide a unique description of the distribution of a stochastic process, as long as the underlying family of state spaces $(\alpha_t)_{t\in\iota}$ is nice enough.
-This means that these spaces are usually required to be complete and separable metric (in particular topological) spaces.
-This distrubution is a measure on the product-$\sigma$-field $\mathcal F :=\bigotimes_{t\in\iota} \mathcal B(\alpha_t)$ (assuming that $\mathcal B(\alpha_t)$ is the Borel $\sigma$-algebra on $\alpha_t$).
+However, it is the achievement of Kolmogorov that the finite-dimensional distributions in fact provide a unique description of the distribution of a stochastic process, as long as the underlying family of state spaces $(\alpha_t)_{t\in\iota}$ is nice enough (Polish, i.e.\ a separable topological space which can be metrized by a complete metric, for example).
+Assuming a family of topological space $(\alpha_t)_{t\in\iota}$, this distribution is a measure on the product-$\sigma$-field $\mathcal F :=\bigotimes_{t\in\iota} \mathcal B(\alpha_t)$ (where $\mathcal B(\alpha_t)$ is the Borel $\sigma$-algebra on $\alpha_t$).
 Here, $\mathcal F$ is generated by finite projections and hence any element of $\mathcal F$ may only depend on at most countably many $t\in \iota$, making this a rather coarse $\sigma$-algebra.
-(In particular, note that this is not the Borel $\sigma$-algebra of the product topology.) The result responsible for this insight is usually denoted the {\em Kolmogorov extension theorem}, formulated in \cite{kolmogoroff1933grundbegriffe}.
+(In particular, note that this is not the Borel $\sigma$-algebra of the product topology for infinite $\iota$.) The result responsible for this insight is usually denoted the {\em Kolmogorov extension theorem}, formulated in \cite{kolmogoroff1933grundbegriffe}.
 We note that a version of the extension result was proved by Daniell in the 1930s, but this paper was not acknowledged by the probabilists of that time \cite{aldrich2007but}.
@@
 
-The goal of our contribution is a formalization of the proof of that theorem in {\tt mathlib4}, the mathematical library of {\tt Lean3} \cite{mathlib2020lean}.
+The goal of our contribution is a formalization of the proof of that theorem in {\tt mathlib4}, the mathematical library of {\tt Lean4} \cite{mathlib2020lean}.
 The Kolmogorov extension theorem is on the interface between measure theory and probability theory.
@@
 
-The main steps in our construction, which were previously missing in {\tt mathlib4}, are (more mathematical details are below):
+The main steps in our construction, which were previously missing in {\tt mathlib4}, are (more mathematical details are below, and in standard textbooks on Probability Theory, e.g.\ \cite{Klenke2013, Kallenberg2020}):
 \begin{enumerate}
@@
 \item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
-\item a proof that single probability measures on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg} for a definition; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below);
+\item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below);
 \item the classical Carathéodory extension theorem, providing us a candidate for the measure which we want to construct (see Theorem~\ref{T:masseind}); 
-\item the Kolmogorov extension theorem, as based on the previous steps.
+\item the Kolmogorov extension theorem, as based on the previous steps (See Theorem~\ref{T1}).
 \end{enumerate}
 The Kolmogorov extension theorem has been previously formalized in Isabelle/HOL \cite{Immler2012}.
-However, this formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical.
+This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical.
 However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations in these two interactive theorem provers.
@@
 \subsubsection*{Possible future work}
-Let us describe some future projects extending {\tt mathlib4} which become possible by our contribution.
+\sloppy Let us describe some future projects extending {\tt mathlib4} which become possible by our contribution.
 
@@
 {\em Instances of stochastic processes:} An obvious application of Kolmogorov's extension theorem is the construction of basic stochastic processes like Markov chains \cite[Section~11]{Kallenberg2020}, the Poisson process \cite[Section~13]{Kallenberg2020} and Brownian motion \cite[Section~14]{Kallenberg2020}.
-(Note that other Gaussian processes -- indexed by $[0, \infty)$ or $\mathbb R$ might as well be constructed the same way.
+(Note that other Gaussian processes -- indexed by $[0, \infty)$ or $\mathbb R$ -- might as well be constructed the same way.
 In addition, fields like the Gaussien free field indexed by $\mathbb R^d$ \cite{sheffield2007gaussian} can also be given.) For these tasks, we would have to define the finite-dimensional distributions (Poisson and normally distributed, respectively), and apply the extension theorem.
@@
 Extra work is needed in order to show that -- on the same probability space -- we can as well define a version (i.e.\ another process $(Y_t)_{t\in\iota}$ with $\mathbf P(X_t = Y_t) = 1$ for all $t\in\iota$) which is right-continuous with left limits (for the Poisson process) or continuous (for Brownian motion), respectively.
-For the former, this follows at least from some general principles of Markov processes (e.g.\ \cite[Theorem 4.3.6]{EthierKurtz1986}).
+For the former, this follows from some general principles of Markov processes (e.g.\ \cite[Theorem 4.3.6]{EthierKurtz1986}).
 For the latter, this requires formalization of the Kolmogorov-Chentsov criterion
 \cite[Theorem~4.23]{Kallenberg2020}.
+
+\noindent
+{\em Theorem of Ionescu-Tulcea and related results:} While Kolmogorov's extension Theorem gives a result for arbitrary $\iota$, but requires some properties of the family $(\alpha_t)_{t\in\iota}$, the Theorem of Ionescu-Tulcea \cite[Theorem 8.24]{Kallenberg2020} can only deal with countable $\iota$, but has no restrictions on $(\alpha_t)_{t\in\iota}$.
+The proof of the latter, however, uses transition kernels (and an inductive construction of a content, which can be extended to a measure) rather than projective families.
+However, continuity of a measure at $\emptyset$ is used as well and a formalization would complement our work.
+In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} without requirements on $(\alpha_t)_{t\in\iota}$.
+All of these results can be used in order to construct countably many independent random variables, which are frequently used in concrete constructions in probability theory.
+Examples are (one direction of) the Borel-Cantelli-Lemma \cite[Theorem 4.28]{Kallenberg2020}, Kolmogorov's 0-1-law \cite[Theorem 4.13]{Kallenberg2020}, random walks \cite[Chapter 12]{Kallenberg2020}, branching processes  \cite[Chapter 13]{Kallenberg2020}, and percolation  \cite[Chapter 2.4]{Klenke2013}, to name just a few.
+While the first two are already formalized in \leanline{mathlib4}, it is only clear due to the Kolmorov extension theorem (or the Theorem of Ionesu-Tulcea) that infinitely many independent random variables can indeed be defined on a joint probability space.
+In this sense, our contribution is also important for giving more sense to already formalised parts of \leanline{mathlib}.
 
@@
 %In addition, there are still new proofs found for the general extension theorem \cite{chin2019new}, but they require some results not yet formalized in mathlib4 (Borel isomorphism of $\mathbb R$ to $\{0,1\}^{\mathbb N}$), so we do not follow this route.
-Whenever we feel necessary, we rely on a standard textbooks on Probability Theory \cite{Klenke2013, Kallenberg2020} for general purpose.
-
+
+\subsection{Formulating the result}
 Before we can state the result, we begin with some mathematical notions.
-We start off slowly from metric spaces, but quickly introduce projective families of measures.
+We start off from metric spaces, but quickly introduce projective families of measures.
 In all definitions, $\alpha$ will be some set (or type, since {\tt Lean4} is a dependently typed language, \leanline{α : Type _}).
@@
       there is $N\in\mathbb N$ such that $r(x_n, x_m) < \varepsilon$
-      for all $m,n > N$. The (pseudo-extended) metric is called
+      for all $m,n > N$. The (extended pseudo-) metric is called
       complete if every Cauchy-sequency has a limit in
@@
       \subseteq 2^\alpha$ is called a topology, if it satisfies (i)
-      $\emptyset, \alpha \in \mathcal O$, (ii) $\mathcal O$ is stable
-      under arbitrary unions, i.e.\ if $A_i \in \mathcal O$ for all
-      $i\in\iota$ and $\iota$ is arbitrary, then $\bigcup_{i\in \iota}
-      A_i \in \mathcal O$, (iii) $\mathcal O$ is stable under finite
-      intersections, i.e.\ if $A_1,...,A_n \in \mathcal O$, then
-      $\bigcap_{i\in \iota} A_i \in \mathcal O$.
+      $\emptyset, \alpha \in \mathcal O$, (ii) $\mathcal O$ is a
+      $\pi$-system, i.e.\ it is stable under finite intersections,
+      i.e.\ if $A, B \in \mathcal O$, then $A \cap B \in \mathcal O$,
+      and (iii) $\mathcal O$ is stable under arbitrary unions,
+      i.e.\ if $A_i \in \mathcal O$ for all $i\in\iota$ and $\iota$ is
+      arbitrary, then $\bigcup_{i\in \iota} A_i \in \mathcal O$.
   \end{enumerate}
@@
     \item \sloppy In \leanline{Lean4}, we have the \leanline{class
-      PseudoEMetricSpace (α : Type u) extends EDist}, which comes with
-      the components, \leanline{edist_self}, \leanline{edist_comm},
+      PseudoEMetricSpace (α : Type u) extends EDist}, where
+      \leanline{EDist} comes with an extended distance \leanline{edist
+        : α → α → ENNReal}, and the extension to
+      \leanline{PseudoEMetricSpace} consists of the components,
+      \leanline{edist_self}, \leanline{edist_comm},
       \leanline{edist_triangle} providing the properties of the
@@
       the uniformity; see \cite{james2013topologies} for details. We
-      note that uniform spaces with a countably generated uniformity
+      note that a uniform space with a countably generated uniformity
       filter is pseudometrizable, i.e.\ there exists a
@@
       \cite{melikhov2011metrizable}.
-    \item \sloppy As for topological spaces, we have \leanline{class
+    \item \sloppy For topological spaces, we have \leanline{class
       TopologicalSpace(α : Type u) : Type u}, which comes with
-      \leanline{isOpen}, \leanline{isOpen_univ},
-      \leanline{isOpen_inter} and \leanline{isOpen_sUnion}, which are
-      exactly the properties of a topological space described above.
+      \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen
+        Set.univ}, \leanline{isOpen_inter} and
+      \leanline{isOpen_sUnion}, which are exactly the properties of a
+      topological space described above.
     \item Any (extended pseudo-)metric on $\alpha$ defines a topology,
@@
       r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
-      (0,\infty)\}$. Actually, in \leanline{mathlib4}, as we saw
+      (0,\infty)\}$. Actually, in \leanline{mathlib4}, as we have seen
       above, every extended pseudo-metric defines a uniform space, and
       a uniform space is a \leanline{class UniformSpace(α : Type u)
-        extends TopologicalSpace , UniformSpace.Core} with xxx
-      \leanline{to_topological_space} giving a topological space,
-      \leanline{to_core} which stores the uniformity filter, and
-      \leanline{is_open_uniformity}, which gives all open sets. In
-      particular, a uniform space defines a topological space, which
-      can be used by the typeclass system.
+        extends TopologicalSpace , UniformSpace.Core} (where the
+      uniformity is defined in \leanline{UniformSpace.Core}), which
+      connects the \leanline{uniformity} with the topology using
+      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p |
+          p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
+      uniform space defines a topological space, which can be used by
+      the typeclass system.
   \end{enumerate}
@@
     countable $s \subseteq \iota$ such that $\inf\{r(x,y) : y \in s\}
-    = 0$ for all $x\in\alpha$. \\ If there is a countable basis of the
-    topology, it is separable; see
+    = 0$ for all $x\in\alpha$. (More generally, a
+    \leanline{TopologicalSpace α} is a \leanline{SeparableSpace} iff
+    \leanline{∃ s, Set.Countable s ∧ Dense s}, where the latter means
+    \leanline{∀ (x : α), x ∈ closure s}, i.e.\ the space equals its
+    own closure.) \\ If there is a countable basis of the topology, it
+    is separable; see
     \leanline{TopologicalSpace.SecondCountableTopology.to_separableSpace}.
@@
 \noindent
-Finally, we can introduce measures. Still, we need even more
-set-systems, i.e.\ $\sigma$-algebras.
-
-\begin{definition}[$\sigma$-algebras and measures]
+Finally, we can introduce measures. A measure is defined on a
+$\sigma$-algebra, which we introduce next.
+
+\begin{definition}[$\sigma$-algebras and measures]\label{def:sigma}
   \begin{enumerate}
@@
     We call $\mathcal F \subseteq 2^\alpha$ a $\sigma$-algebra (on
-    $\alpha$) if (i) $\alpha \in \mathcal F$, (ii) $\mathcal F$ is
+    $\alpha$) if (i) $\emptyset \in \mathcal F$, (ii) $\mathcal F$ is
     stable under complements, i.e.\ $A \in \mathcal F \Rightarrow A^c
@@
   \item For some $\sigma$-algebra $\mathcal F$ on $\alpha$, a function
-    $\mu : \mathcal F \to [0, \infty]$ is called a measure,
-    if\footnote{Since $\alpha \in \mathcal F$ and $\mathcal F$ is
-      stable under complements, we also have that $\emptyset \in
-      \mathcal F$} (i) $\mu(\emptyset) = 0$, (ii) $\mu$ is countably
-    additive, i.e.\ for $A_1, A_2,... \in \mathcal F$ pairwise
-    disjoint, we have $\mu\Big( \bigcup_{n=1}^\infty A_n\Big) =
-    \sum_{n=1}^\infty \mu(A_n)$. We call $(\alpha, \mathcal F, \mu)$ a
-    measure space. \\ In addition, $\mu$ is called finite if
-    $\mu(\alpha) < \infty$ and a probability measure if
-    $\mu(\alpha)=1$.
+    $\mu : \mathcal F \to [0, \infty]$ is called a measure, if (i)
+    $\mu(\emptyset) = 0$, (ii) $\mu$ is countably additive, i.e.\ for
+    $A_1, A_2,... \in \mathcal F$ pairwise disjoint, we have $\mu\Big(
+    \bigcup_{n=1}^\infty A_n\Big) = \sum_{n=1}^\infty \mu(A_n)$. We
+    call $(\alpha, \mathcal F, \mu)$ a measure space. \\ In addition,
+    $\mu$ is called finite if $\mu(\alpha) < \infty$ and a probability
+    measure if $\mu(\alpha)=1$.
   \end{enumerate}
@@
   \begin{enumerate}
-  \item \sloppy The \leanline{class MeasurableSpace(α : Type u_7) :
-    Type u_7} is very similar to a topological space in
-    \leanline{mathlib4}, since it comes with
-    \leanline{MeasurableSet'}, \leanline{measurableSet_empty},
-    \leanline{measurableSet_compl}, and
-    \leanline{measurableSet_iUnion}.
-  \item For a measure space, we require a measurable space as a
-    variable, and add the set function: \leanline{structure
-      MeasureTheory.Measure(α : Type u_6) [inst : MeasurableSpace α]
-      extends MeasureTheory.OuterMeasure}. Here $\sigma$-additivity of
-    a \leanline{Measure α} reads xxx note on outer measures
+  \item \sloppy The \leanline{class MeasurableSpace α} is very similar
+    to a topological space in \leanline{mathlib4}, since it comes with
+    \leanline{MeasurableSet' : Set α → Prop},
+    \leanline{measurableSet_empty : MeasurableSet' ∅},
+    \leanline{measurableSet_compl : (s : Set α) → MeasurableSet' s →
+      MeasurableSet' sᶜ}, and \leanline{measurableSet_iUnion},
+    i.e.\ properties (i)-(iii) from Definition~\ref{def:sigma}.1.
+  \item We are using outer measures in our construction. In
+    \leanline{mathlib4}, this is \leanline{structure
+      MeasureTheory.OuterMeasure α} which is a function
+    \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf
+      ∅ = 0} (i.e.\ the empty set has measure~$0$), \leanline{∀ {s₁ s₂
+        : Set α}, s₁ ⊆ s₂ → measureOf s₁ ≤ measureOf s₂}
+    (i.e.\ monotonicity) and \leanline{∀ (s : ℕ → Set α), measureOf (⋃
+      (i : ℕ), s i) ≤ ∑' (i : ℕ), measureOf (s i)} (which we call
+    $\sigma$-subadditivity). \\ Note that an outer measure is defined
+    on $2^\alpha$, whereas a measure is only defined on a subset (the
+    $\sigma$-algebra of measurable sets). More precisely: A
+    \leanline{MeasureSpace α} extends a \leanline{MeasurableSpace α}
+    by a \leanline{Measure α} (which is usually called
+    \leanline{volume}). For the measure, we have: \leanline{structure
+      MeasureTheory.Measure α [MeasurableSpace α] extends
+      MeasureTheory.OuterMeasure}. Here $\sigma$-additivity of a
+    \leanline{Measure α} reads
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
     measurable (with respect to $\mathcal F$ and $\mathcal G$) if
-    $f^{-1} \mathcal G \subseteq \mathcal F$.  \\ In this case, if
-    $\mu$ is a measure on $\mathcal F$, the measure $\nu: \mathcal G
-    \to [0, \infty], \nu(B) := \mu (f^{-1}(B))$ is called the image
-    (or push-forward) measure of $\mu$ under $f$. We write $\nu :=
-    f_\ast \mu$; see \leanline{MeasureTheory.Measure.map}.
+    $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib4},
+    see \leanline{Measurable} and note that this notion usually comes
+    with some \leanline{m : Measure α} and $f$ is almost everywhere
+    measurable (\leanline{AEMeasurable}) if it coincices with some
+    \leanline{Measurable g}, \leanline{m}-almost everywhere.\\ In this
+    case, if $\mu$ is a measure on $\mathcal F$, the measure $\nu:
+    \mathcal G \to [0, \infty], \nu(B) := \mu (f^{-1}(B))$ is called
+    the image (or push-forward) measure of $\mu$ under $f$. We write
+    $\nu := f_\ast \mu$; see \leanline{MeasureTheory.Measure.map}.
   \end{enumerate}
@@
 
-The extension theorem is on finite measures on product spaces. In
+We now come to our contribution, which we will PR to mathlib4: The
+extension theorem is on finite measures on product spaces. In
 particular, the product is taken over a set of arbitrary
@@
 any finite subset of indices in a consistent way. Recall that finite
-sets (of some type \leanline{ι} are formalized with \leanline{Finset
+sets (of some type \leanline{ι}) are formalized with \leanline{Finset
   ι}. Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
@@
   \item A family $(P_J)_{J\subseteq_f I}$, where $P_J$ is a finite
-    measure on $\mathcal F^J$, is called projective if
+    measure on $\mathcal F_J$, is called projective if
     $$P_H = (\pi_H^{J})_\ast P_J$$ for all
@@
     Type _} }, which fixes the index set $\iota$ and all spaces
-$\alpha_t, t\in\iota$. In addition, note that the projective property
-we use here works as long as we have a preorder (which is the subset
-relation on \leanline{Finset ι} below).
+$\alpha_t, t\in\iota$ as global variables. In addition, note that the
+projective property we use here works as long as we have a preorder
+(which is the subset relation on \leanline{Finset ι} below).
 
@@
 \noindent
-It is worth understanding the precise connection ot
+It is worth understanding the precise connection of
 \leanline{isProjective} and \leanline{isProjectiveMeasureFamily}. In
@@
 \leanline{ι}. The second variable is the functions which maps two sets
-\leanline{I J : Finset ι} and a proof \leanline{hJI} of \leanline{J ⊆
-  I} together with \leanline{P I} to the image measure on
-\leanline{J}, which is \leanline{μ.map (fun x : ∀ i : I, α i => fun j
-  => x ⟨j, hJI j.2⟩)}. The map defined by the \leanline{=>}-notation
-maps every $(x_i)_{i\in I}$ to a function of \leanline{j}, whose type
-is the subtype of \leanline{I}, consisting of a value and a proof of
+\leanline{I J : Finset ι} and a proof \leanline{hJI} of \leanline{J ⊆ I}
+together with \leanline{P I} (which is \leanline{μ} in the
+statement) to the image measure on \leanline{J}, which is
+\leanline{μ.map (fun x : ∀ i : I, α i => fun j => x ⟨j, hJI
+    j.2⟩)}. The map defined by the \leanline{=>}-notation maps every
+$(x_i)_{i\in I}$ to a function of \leanline{j}, whose type is the
+subtype of \leanline{I}, consisting of a value and a proof of
 \leanline{J ⊆ I}. In other words, this is $(x_j)_{j\in J}$.
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def projectiveLimitWithWeakestHypotheses
+  noncomputable def
+    projectiveLimitWithWeakestHypotheses
     [∀ i, PseudoEMetricSpace (α i)]
@@
 in Theorem~\ref{T1} is a ring of sets (see the next definition)
-containing the whole set. We will work with the weaker semi-ring as
-introduced next in the formulation of Carathéodory's extension
-theorem.
+containing the whole set. (This is sometimes called a field of sets.)
+We will work with the weaker semi-ring as introduced next in the
+formulation of Carathéodory's extension theorem; see
+e.g.\ \cite[Definition~1.9]{Klenke2013}.
 
@@
 
+\noindent
 Using \leanline{variable {α : Type _} {C : Set (Set α)} {s t : Set α}
@@
 
-Recall that \leanline{Lean4} indicates a coercion by \leanline{↑},
-which in this case is from \leanline{Finset} to \leanline{Set}.
+Let us remark that (i) \leanline{Lean4} indicates a coercion by
+\leanline{↑}, which in this case is from \leanline{Finset} to
+\leanline{Set} and (ii) we have \leanline{PairwiseDisjoint (s : Set ι)
+  (f : ι → α)} iff the images of any distinct elements of \leanline{ι}
+under \leanline{f} are different. (Hence, if \leanline{f = id}, the
+usual definition of pairwise disjoint sets unfolds.)
 
@@
 any ring is a semi-ring since $A\cap B = A \setminus (A \setminus B)$,
-i.e.\ every ring is a $\pi$-system. In our formlaization, we also use
-the notion of a field $\mathcal H$, which is a ring with $\alpha \in
-\mathcal H$. Since we a re dealing with product spaces, we will use
-cylinders, which are defined as follows:
+i.e.\ every ring is a $\pi$-system.
+%In our formalization, we also use the notion of a field $\mathcal H$,
+%which is a ring with $\alpha \in
+%\mathcal H$.
+Since we are dealing with product spaces, we will use cylinders, which
+are defined as follows:
 
@@
 $S$. (This is what the last line in previous definition gives.) From
-this, we can defined the set of all measurable cylinders.
+this, we can define the set of all measurable cylinders.
 
@@
 The set-system of cylinders is in fact a field, hence a ring and hence
-a semi-ring:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem setField_cylinders : SetField (cylinders α) :=
-    { empty_mem := empty_mem_cylinders α
-      univ_mem := univ_mem_cylinders α
-      union_mem := union_mem_cylinders
-      diff_mem := diff_mem_cylinders }
-\end{minted}
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem setRing_cylinders :
-    MeasureTheory.SetRing (cylinders α) :=
-    setField_cylinders.toSetRing
-\end{minted}
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem setSemiringCylinders :
-    MeasureTheory.SetSemiring (cylinders α) :=
-    setField_cylinders.setSemiring
-\end{minted}
-
+a semi-ring. This means we can prove the following:
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  theorem setField_cylinders : SetField (cylinders α) 
+  theorem setRing_cylinders :  SetRing (cylinders α)
+  theorem setSemiringCylinders : SetSemiring (cylinders α)   
+\end{minted}
+
+From the field/ring/semi-ring of cylinders, we have to define the
+generated $\sigma$-alebra. This uses the following:
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
     = MeasurableSpace.pi
-\end{minted}
-
-measurable product space
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+
   instance MeasurableSpace.pi
@@
     MeasurableSpace (∀ a, π a) :=
-    ⨆ a, (m a).comap fun b => b a
-\end{minted}
+    iSup a, (m a).comap fun b => b a
+\end{minted}
+Here, \leanline{comap {β : Type} (f : α → β) (m : MeasurableSpace β)}
+is the measurable space consisting of pre-images of measurable subsets
+of \leanline{β}, and the set of $\sigma$-algebras on the product
+$\prod_{i\in\iota} \alpha_i$ (or on any other space) is a complete
+lattice, i.e.\ the subset defined by the \leanline{comap}s has a
+supremum, which defines the $\sigma$-algebra generated by the
+cylinders.
 
@@
   $A \in \mathcal H$. Then, there is $\mathcal K \subseteq_f \mathcal
-  H$ such that all sets in $\mathcal K$ are pairwise disjoint and $A
+  H$ such that $\mathcal K$ contains pairwise disjoint sets and $A
   \setminus \bigcup_{I \in \mathcal I} I = \biguplus_{K\in \mathcal K}
@@
 \end{lemma}
-
-The formalization of the proof is somewhat straight-forward, but
-requires induction over finite sets.
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem exists_disjoint_finset_diff_eq
-    (hC : SetSemiring C) (hs : s ∈ C)
-    (I : Finset (Set α)) (hI : ↑I ⊆ C) :
-    ∃ (J : Finset (Set α)) (_h_ss : ↑J ⊆ C)
-    (_h_dis : PairwiseDisjoint (J : Set (Set α)) id),
-    s \ ⋃₀ I = ⋃₀ J
-\end{minted}
-
-\noindent
-Thes existance-statement of the above lemma actually gives rise to a
-definition:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def diff₀ (hC : SetSemiring C)
-    (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
-    [DecidableEq (Set α)] : Finset (Set α) :=
-    (hC.exists_disjoint_finset_diff_eq hs I hI).choose \ {∅}
-\end{minted}
-
-Complementing Lemma~\ref{l1}, we actually need another result for some
-constructions needed in the proofs: For some $A_1,...,A_m$ and $n\neq
-m$, we need finite sets $\mathcal K_1,...,\mathcal K_m$ such that $A_n
-\setminus \bigcup_{i=1}^{n-1} A_i = \biguplus_{K \in \mathcal K_n} K$
-for $n=1,...,m$. Using this, we can write
-$$ \bigcup_{n=1}^m A_n = A_1 \uplus (A_2 \setminus A_1) \uplus \cdots
-\uplus \Big(A_m \setminus \bigcup_{i=1}^{m-1}A_i\Big) =
-\bigcup_{n=1}^m \biguplus_{K \in \mathcal K_n} K,$$ i.e.\ we can write
-a union of sets in the semi-ring as a disjoint union of sets in the
-semi-ring. The formalized version is as follows:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def indexedDiff₀ (hC : SetSemiring C)
-    (J : Finset (Set α)) (hJ : ↑J ⊆ C)
-    (n : Fin J.card) : Finset (Set α) :=
-    hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
-    (finsetLT_subset' J hJ n)
-\end{minted}
 
@@
 
+The proof as well as its formalization is somewhat straight-forward,
+but requires induction over finite sets.
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  theorem exists_disjoint_finset_diff_eq
+    (hC : SetSemiring C) (hs : s ∈ C)
+    (I : Finset (Set α)) (hI : ↑I ⊆ C) :
+    ∃ (J : Finset (Set α)) (_h_ss : ↑J ⊆ C)
+    (_h_dis : PairwiseDisjoint (J : Set (Set α)) id),
+    s \ ⋃₀ I = ⋃₀ J
+\end{minted}
+
+\noindent
+The existence-statement of the above lemma actually gives rise to a
+definition. Here, we use \leanline{Exists.choose} in order to extract
+an element from an $\exists$-statement. In addition, we do not allow
+$\emptyset \in \mathcal K$ without loss of generality:
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  noncomputable def diff₀ (hC : SetSemiring C)
+    (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
+    [DecidableEq (Set α)] : Finset (Set α) :=
+    (hC.exists_disjoint_finset_diff_eq hs I hI).choose \ {∅}
+\end{minted}
+
+Extending Lemma~\ref{l1}, we would like to write a finite union of
+members of a semi-ring as a finite union of disjoint sets.
+
+\begin{lemma}\label{l2}
+  Let $\mathcal H$ be a semi-ring and $A_1,...,A_m \in \mathcal
+  H$. Then, there are $\mathcak K_1,...,\mathcal K_m \subseteq_f
+  \mathcal H$ disjoint such that $\bigcup_{n=1}^m \mathcal K_n$
+  contains disjoint sets and $\bigcup_{m=1}^n A_m = \biguplus_{m=1}^n
+  \biguplus_{K \in \mathcal K_n} K$.  
+\end{lemma}
+
+\begin{proof}
+  Indeed, we may write $ \bigcup_{m=1}^n A_m = \biguplus_{n=1}^m
+  \Big(A_n \setminus \bigcup_{i=1}^{n-1}A_i\Big).$ Then, the result
+  follows by applying Lemma~\ref{l1} to $A_n \setminus
+  \bigcup_{i=1}^{n-1}A_i$, $n=1,...,m$.
+\end{proof}
+
+\noindent
+The formalized version again gives rise to a definition:
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  noncomputable def indexedDiff₀ (hC : SetSemiring C)
+    (J : Finset (Set α)) (hJ : ↑J ⊆ C)
+    (n : Fin J.card) : Finset (Set α) :=
+    hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
+    (finsetLT_subset' J hJ n)
+\end{minted}
+
+Here, recall that \leanline{Fin n} is the subtype of \leanline{Nat}
+consisting of all numbers \leanline{<n}. In addition,
+\leanline{ordered_mem' (hJ : ↑J ⊆ C) (n : Fin (card J)) : ↑(ordered J)
+  n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
+is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin
+  (card J)) : Finset (Set α)} is the set consisting of the
+\leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
+\leanline{finsetLT_subset'} gives a proof that \leanline{↑(finsetLT J
+  n) ⊆ C}. So, in terms of Lemma~\ref{l2}, this gives, for
+$n=1,...,m$, a construction for $\mathcal K_n \subseteq_f \mathcal H$.
+
 \noindent
@@
 semi-open intervals generates the Borel $\sigma$-algebra on $\mathbb
-R$.) We will follow this construction, and start by stating some basic
-concepts.
+R$.) We will follow this abstract construction, and start by stating
+some basic concepts.
 
@@
 formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
-m(\emptyset)$. However, sicne \leanline{Finset}s are assumed to have
-distinct elements, we cannot have $\emptyset$ twice in $I$.
+m(\emptyset)$. However, since \leanline{Finset}s are assumed to have
+distinct elements, we cannot have \leanline{∅} twice in \leanline{I}.
 
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  noncomputable def extendContent
+    (hC : SetSemiring C)
+    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
+    (m_empty : m ∅ hC.empty_mem = 0)
+    (m_add :
+       ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
+       (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
+    (h_mem : ⋃₀ ↑I ∈ C),
+    m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop)) :
+    AddContent C
+
   noncomputable def kolContent

### 2023-09-08 Manuscript/proof_outline.tex
updated including 2.3

@@
 
-\title{Formalizing the Kolmogorov extension Theorem in mathlib4}
-
-\keywords{probability, measure theory, Lean4, formal mathematics, proof assistant, mathlib4}
+\title{Formalizing the Kolmogorov extension Theorem in Lean}
+
+\keywords{probability, measure theory, Lean, formal mathematics, proof assistant, mathlib}
 
@@
   which is a main building block in construction the (distribution of)
-  stochastic processes with arbitrary index sets in Lean4. Our
-  approach is based on mathlib, the mathematical library for Lean4. On
+  stochastic processes with arbitrary index sets in Lean. Our
+  approach is based on mathlib, the mathematical library for Lean. On
   our way, we provide a formalization of the Caratheodory extension
@@
 
-The goal of our contribution is a formalization of the proof of that theorem in {\tt mathlib4}, the mathematical library of {\tt Lean4} \cite{mathlib2020lean}.
+The goal of our contribution is a formalization of the proof of that theorem in {\tt Lean}.
+Here, we are going to use \leanline{Lean4}, since \leanline{mathlib} \cite{mathlib2020lean}, the mathematical library of \leanline{Lean} has recently been ported to \leanline{Lean4}.
+Subsequently, we aim to make a pull request for the publication of our code in \leanline{mathlib}.
 The Kolmogorov extension theorem is on the interface between measure theory and probability theory.
-Here, we rely on a decent amount of formalized mathematics in the measure-theory part of {\tt mathlib4} (outer measures, above all), while not requiring any specific previous formalization of probability theory.
+Here, we rely on a decent amount of formalized mathematics in the measure-theory part of {\tt mathlib} (outer measures, above all), while not requiring any specific previous formalization of probability theory.
 (In fact, most of our results are formulated in terms of finite rather than probability measures.)
@@
 
-The main steps in our construction, which were previously missing in {\tt mathlib4}, are (more mathematical details are below, and in standard textbooks on Probability Theory, e.g.\ \cite{Klenke2013, Kallenberg2020}):
+The main steps in our construction, which were previously missing in {\tt mathlib}, are (more mathematical details are below, and in standard textbooks on Probability Theory, e.g.\ \cite{Klenke2013, Kallenberg2020}):
 \begin{enumerate}
@@
 \subsubsection*{Possible future work}
-\sloppy Let us describe some future projects extending {\tt mathlib4} which become possible by our contribution.
+\sloppy Let us describe some future projects extending {\tt mathlib} which become possible by our contribution.
 
@@
 This task requires the formalization of multi-dimensional Poisson and normal distributions, which -- in textbooks -- is usually done using characteristic functions.
-Since these are not yet part of {\tt mathlib4}, we postpone this task to the future.
+Since these are not yet part of {\tt mathlib}, we postpone this task to the future.
 
@@
 Examples are (one direction of) the Borel-Cantelli-Lemma \cite[Theorem 4.28]{Kallenberg2020}, Kolmogorov's 0-1-law \cite[Theorem 4.13]{Kallenberg2020}, random walks \cite[Chapter 12]{Kallenberg2020}, branching processes  \cite[Chapter 13]{Kallenberg2020}, and percolation  \cite[Chapter 2.4]{Klenke2013}, to name just a few.
-While the first two are already formalized in \leanline{mathlib4}, it is only clear due to the Kolmorov extension theorem (or the Theorem of Ionesu-Tulcea) that infinitely many independent random variables can indeed be defined on a joint probability space.
+While the first two are already formalized in \leanline{mathlib}, it is only clear due to the Kolmorov extension theorem (or the Theorem of Ionesu-Tulcea) that infinitely many independent random variables can indeed be defined on a joint probability space.
 In this sense, our contribution is also important for giving more sense to already formalised parts of \leanline{mathlib}.
@@
 Other -- highly readable -- references such as \cite{Billingsley1995} state the extension theorem only in special cases such as $\alpha_t = \mathbb R$ for all $t$.
-%In addition, there are still new proofs found for the general extension theorem \cite{chin2019new}, but they require some results not yet formalized in mathlib4 (Borel isomorphism of $\mathbb R$ to $\{0,1\}^{\mathbb N}$), so we do not follow this route.
+%In addition, there are still new proofs found for the general extension theorem \cite{chin2019new}, but they require some results not yet formalized in mathlib (Borel isomorphism of $\mathbb R$ to $\{0,1\}^{\mathbb N}$), so we do not follow this route.
 
@@
 We start off from metric spaces, but quickly introduce projective families of measures.
-In all definitions, $\alpha$ will be some set (or type, since {\tt Lean4} is a dependently typed language, \leanline{α : Type _}).
+In all definitions, $\alpha$ will be some set (or type, since {\tt Lean} is a dependently typed language, \leanline{α : Type _}).
 
@@
 
-\begin{remark}[Metric and topological spaces in {\tt mathlib4}] \mbox{}
+\begin{remark}[Metric and topological spaces in {\tt mathlib}] \mbox{}
   \begin{enumerate}
-    \item \sloppy In \leanline{Lean4}, we have the \leanline{class
+    \item \sloppy In \leanline{Lean}, we have the \leanline{class
       PseudoEMetricSpace (α : Type u) extends EDist}, where
@@
       r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
-      (0,\infty)\}$. Actually, in \leanline{mathlib4}, as we have seen
+      (0,\infty)\}$. Actually, in \leanline{mathlib}, as we have seen
       above, every extended pseudo-metric defines a uniform space, and
@@
 
-\begin{remark}[Measur(abl)e spaces in \leanline{mathlib4}]
+\begin{remark}[Measur(abl)e spaces in \leanline{mathlib}]
   \begin{enumerate}
   \item \sloppy The \leanline{class MeasurableSpace α} is very similar
-    to a topological space in \leanline{mathlib4}, since it comes with
+    to a topological space in \leanline{mathlib}, since it comes with
     \leanline{MeasurableSet' : Set α → Prop},
@@
   \item We are using outer measures in our construction. In
-    \leanline{mathlib4}, this is \leanline{structure
+    \leanline{mathlib}, this is \leanline{structure
       MeasureTheory.OuterMeasure α} which is a function
@@
     $\mathcal B := \sigma(\mathcal O)$ the Borel $\sigma$-algebra on
-    $\alpha$. In \leanline{mathlib4}, this is
+    $\alpha$. In \leanline{mathlib}, this is
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
 \end{minted}
-    In addition, \leanline{mathlib4} provides a similar notion, which is
+    In addition, \leanline{mathlib} provides a similar notion, which is
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
     measurable (with respect to $\mathcal F$ and $\mathcal G$) if
-    $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib4},
+    $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib},
     see \leanline{Measurable} and note that this notion usually comes
@@
 
-We now come to our contribution, which we will PR to mathlib4: The
+We now come to our contribution, which we will PR to mathlib: The
 extension theorem is on finite measures on product spaces. In
@@
 With this, we can define the projetive family as follows. Note that
-\leanline{Lean4} automatically uses the subset relation as
+\leanline{Lean} automatically uses the subset relation as
 \leanline{[Preorder (Finset ι)]} when \leanline{isProjective} is
@@
 
-Let us remark that (i) \leanline{Lean4} indicates a coercion by
+Let us remark that (i) \leanline{Lean} indicates a coercion by
 \leanline{↑}, which in this case is from \leanline{Finset} to
@@
 \noindent
-An additive $m : \mathcal H \to [0,\infty]$ is formalized as follows:
+For additive contents, we need two definitions. \leanline{AddContent}
+defines a set-function \leanline{tuFun} on \leanline{Set α}, whereas
+\leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C
+  → ℝ≥0∞}.
 
@@
       toFun (⋃₀ I) = ∑ u in I, toFun u
-\end{minted}
-
-\noindent
-In this formalization, we introduced $m(\emptyset)=0$ as a separate
-hypothesis with \leanline{empty'}. Note that this also follows
-formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
-m(\emptyset)$. However, since \leanline{Finset}s are assumed to have
-distinct elements, we cannot have \leanline{∅} twice in \leanline{I}.
-
-% extend_content put $\infty$ outside the set system
-
-For the concrete application we have in mind, we need to be more
-specific which semiring to use, based on a projective family of
-probability measures. We do this using \leanline{kolContent}, which
-is based on the following function, we takes a family of measures and
-a cylinder set as input:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def kolmogorovFun
-    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
-    (s : Set (∀ i, α i))
-    (hs : s ∈ cylinders α) : ℝ≥0∞ :=
-    P (cylinders.finset hs) (cylinders.set hs)
-\end{minted}
-
-\noindent
-With this, we define an \leanline{add_content} on cylinders. (Note
-that \leanline{extend_content} in fact defines an additive content on
-a semiring.)
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+
   noncomputable def extendContent
@@
     AddContent C 
+\end{minted}
+
+\noindent
+In these formalizations, we introduced $m(\emptyset)=0$ as a separate
+hypothesis with \leanline{empty'}. Note that this also follows
+formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
+m(\emptyset)$. However, since \leanline{Finset}s are assumed to have
+distinct elements, we cannot have \leanline{∅} twice in \leanline{I}.
+
+For the concrete application we have in mind, we need to be more
+specific which semiring to use, based on a projective family of
+probability measures. We do this using \leanline{kolContent}, which is
+based on \leanline{kolmogorovFun}, which takes a family of measures
+and a cylinder set as input:
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  noncomputable def kolmogorovFun
+    (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
+    (s : Set (∀ i, α i))
+    (hs : s ∈ cylinders α) : ℝ≥0∞ :=
+    P (cylinders.finset hs) (cylinders.set hs)
 
@@
 
+Here, \leanline{kolmogorovFun_empty} and
+\leanline{kolmogorovFun_additive} give proofs for \leanline{m_empty}
+and \leanline{m_add} in \leanline{extendContent} as applied to a
+\leanline{kolmogorovFun}.
+
 \noindent
@@
 well as $\sigma$-additive $\Rightarrow$ $\sigma$-sub-additive in the
-proof of the Carathéodory extension, Theorem~\ref{T:masseind}.
+proof of the Carathéodory extension, Theorem~\ref{T:masseind}. The
+formalization of the next result comes in several lemmas, which we do
+not show here.
 
@@
 \noindent
-Although some material on outer measures was covered in {\tt mathlib4}
+Although some material on outer measures was covered in {\tt mathlib}
 already, the classical extension theorem (extending a set function $m$
@@
 H$. All statements on equality of $\mu$ and $m$ present in {\tt
-  mathlib4} at the time of writing have too many requirements.
+  mathlib} at the time of writing have too many requirements.
 
 The next result talks about an \leanline{induced_outer_measure} in
-terms of {\tt mathlib4}.
-
-Theorem 2.5
+terms of {\tt mathlib}.
 
@@
   Let \label{P:auss} $\mathcal H$ be a semi-ring and $m: \mathcal
-  H\to\mathbb R_+$ additiv. For $A\subseteq E$ let
+  H\to\mathbb R_+$ additive. For $A\subseteq E$ let
   $$ \mu(A) := \inf_{\mathcal G \in \mathcal U(A)}
@@
 space \leanline{OuterMeasure.caratheodory}, covered in {\tt
-  mathlib4}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}
+  mathlib}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}
 
@@
 \noindent
-\sloppy The second result, is not yet covered in {\tt
-  mathlib4}. However, for product spaces, in
-\leanline{measureTheory.Constructions.Pi}, there is
+\sloppy The second result states that for an outer measure induced by
+an additive content on a semi-ring, we have $\sigma(\mathcal H)
+\subseteq \mathcal F$. In particular, we then have defined a measure
+on $\sigma(\mathcal H)$. This result is not yet covered in {\tt
+  mathlib}. However, for product spaces, in
+\leanline{MeasureTheory.Constructions.Pi}, there is
 \leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
@@
 \end{theorem}
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def Measure.ofAddSubaddCaratheodory
-    (hC : SetSemiring C)
-    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
-    (m_empty : m ∅ hC.empty_mem = 0)
-    (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
-      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
-      (h_mem : ⋃₀ ↑I ∈ C), m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
-    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
-      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
-      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
-      @Measure α (inducedOuterMeasure m
-        hC.empty_mem m_empty).caratheodory
-\end{minted}
-
-\noindent
-Here, \leanline{MeasureTheory.OuterMeasure.caratheodory} is already
-implemented and gives the assertion that $\sigma(\mathcal
-H)\subseteq\mathcal F$.
-
-\noindent
-In the proof of Kolmogorov's extension theorem, we use the following
-definition, which is based on
-\leanline{Measure.ofAddSubaddCaratheodory}.
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def Measure.ofAddContent
-    [mα : MeasurableSpace α] (hC : SetSemiring C)
-    (hC_gen : MeasurableSpace.generateFrom C = mα)
-    (m : AddContent C)
-    (m_sigma_subadd :
-      ∀ ⦃f : ℕ → Set α⦄ (_hf : ∀ i, f i ∈ C) (_hf_Union : (⋃ i, f i) ∈ C),
-        m (⋃ i, f i) ≤ ∑' i, m (f i)) :
-        Measure α
-\end{minted}
 
@@
   \mathcal F$, since $\mathcal F$ is a $\sigma$-algebra.
-
 \end{proof}
+
+Here is a formalization of the measure resulting from
+Theorem~\ref{T:masseind}:
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  noncomputable def ofAddSubaddCaratheodory
+    (hC : SetSemiring C)
+    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
+    (m_empty : m ∅ hC.empty_mem = 0)
+    (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
+      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
+      (h_mem : ⋃₀ ↑I ∈ C),
+      m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
+    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
+      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
+      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
+      @Measure α (inducedOuterMeasure m
+        hC.empty_mem m_empty).caratheodory
+\end{minted}
+
+\noindent
+Here, \leanline{MeasureTheory.OuterMeasure.caratheodory} is already
+implemented and gives the assertion that $\mathcal F$ from
+Theorem~\ref{T:cara} is a $\sigma$-algebra. Note that
+\leanline{@Measure} is the same as \leanline{Measure}, but all
+variables have to be given explicitely. The second argument is a proof
+that measurable sets form a $\sigma$-algebra which cannot be inferred
+by the type class system of \leanline{Lean}.
+
+\noindent
+In the proof of Kolmogorov's extension theorem, we use a closely
+related definition of \leanline{Measure.ofAddContent}, based on
+\leanline{ [mα : MeasurableSpace α] (hC : SetSemiring C) (hC_gen :
+  MeasurableSpace.generateFrom C = mα)}, i.e.\ the semi-ring
+\leanline{C} generates the $\sigma$-algebra, and which uses an
+\leanline{m : AddContent} as well as \leanline{m_sigma_subadd} as
+above.
 
@@
 The above definition closely resembles its formalization in
-\leanline{mathlib4}:
+\leanline{mathlib}:
 
@@
 second countable extended pseudo metric space, which is
-\leanline{OpensMeasurable}. Then, from \leanline{mathlib4}, we use
+\leanline{OpensMeasurable}. Then, from \leanline{mathlib}, we use
 \leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
@@
 \leanline{isCompact_iff_totallyBounded_isComplete} from {\tt
-  mathlib4}, any subset of such a space is compact iff it is complete
+  mathlib}, any subset of such a space is compact iff it is complete
 and totally bounded. Since closed subsets of a complete space are
@@
 set-function on a semi-ring by an outer measure, not previously
-implemented in mathlib4. This could also be used to re-define
+implemented in mathlib. This could also be used to re-define
 Stieltjes measures (in particular Lebesgue measure) on $\mathbb R$, as
@@
 \end{document}
-
-Our formalization broadly follows the procedure described in \cite{hytonen2016analysis} and leads to a definition with the following type.
-\begin{lstlisting}
-def condexp (ℬ : measurable_space Ω)
-  { : measurable_space Ω}
-  (μ : measure Ω) (f : Ω → E) : Ω → E :=
-if hm : ℬ ≤ 
-  then if h : sigma_finite (μ.trim hm) ∧ integrable f μ
-    then if strongly_measurable[ℬ] f then f
-      else (@ae_strongly_measurable'_condexp_L1 _ _ _ _ _ ℬ  μ hm h.1 _).mk (@condexp_L1 _ _ _ _ _ _ _ hm μ h.1 f)
-    else 0
-  else 0
-\end{lstlisting}
-\begin{definition}\label{def:condexp}
-Let $(\Omega, \mathscr A, \mu)$ be a measure space and let $\mathscr B$ be a sub-$\sigma$-algebra of $\mathscr A$.
-Let $E$ be a Banach space (complete normed vector space) and $f : \Omega \to E$ be an i\
-ntegrable function (with respect to $\mu$).
-Then a function $g : \Omega \to E$ is said to be a \emph{conditional expectation} of $f$ with respect to $\mathscr B$ if it is integrable, $\mathscr B$-strong\
-ly measurable and for all $\mathscr B$-measurable sets $s$,
-\begin{align*}
-\int_{x \in s} g(x) d\mu(x) = \int_{x \in s} f(x) d\mu(x) \: .
-\end{align*}
-\end{definition}
-
-%\documentclass[sigplan,10pt,anonymous,review]{acmart}
-\documentclass[sigplan,10pt,screen]{acmart}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{enumitem}
-\setlist[enumerate]{%
-align=left, itemsep=0pt, leftmargin=0pt, labelindent=0pt, listparindent=0pt, labelwidth=0pt, itemindent=!
-}
-
-%\theoremstyle{plain}
-%
-\newtheorem{proposition}{Proposition}[section]
-%\newtheorem{theorem}[proposition]{Theorem}
-\newtheorem{lemma}[proposition]{Lemma}
-\newtheorem{corollary}[proposition]{Korollar}
-\newtheorem{definition}[proposition]{Definition}
-\newtheorem{remark}[proposition]{Remark}
-
-\pdfoutput=1
-
-% Definitions and packages for Lean code
-\usepackage{color}
-\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
-\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
-\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
-\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
-
-\usepackage{listings}
-\def\lstlanguagefiles{lstlean.tex}
-\lstset{language=lean, basicstyle=\small}
-
-% Other packages
-\usepackage[shortlabels]{enumitem}
-\usepackage{tikz}
-
-\input{quiver.tex}
-
-\documentclass{article}
-\usepackage[utf8x]{inputenc}
-
-\usepackage{hyperref}
-\usepackage{refcheck}
-\usepackage[centertags]{amsmath} 
-\usepackage{amssymb, amsthm}
-
-\usepackage{enumitem}
-\setlist[enumerate]{%
-align=left, itemsep=0pt, leftmargin=0pt, labelindent=0pt, listparindent=0pt, labelwidth=0pt, itemindent=!
-}
-
-\theoremstyle{plain}
-%
-\newtheorem{proposition}{Proposition}[section]
-\newtheorem{theorem}[proposition]{Theorem}
-\newtheorem{lemma}[proposition]{Lemma}
-\newtheorem{corollary}[proposition]{Korollar}
-\newtheorem{definition}[proposition]{Definition}
-\newtheorem{remark}[proposition]{Remark}
-
-\title{The Kolmogorov Extension Theorem in Lean} \author{Rémy Degenne,
-  Peter Pfaffelhuber}
-
-\begin{document}
-
-\maketitle
-
-The goal of the project is to state and prove the Kolmogorov extension
-theorem in {\tt Lean 4} and make a PR to {\tt mathlib4}.
-
-\input{formalization}
-
-\bibliographystyle{alpha}
-\bibliography{main}
-
-\end{document}
-
-\begin{thebibliography}{EPW06b}
-\bibitem[Bil95]{Billingsley1995} P.~Billingsley.  \newblock {\em
-  {Probability and Measure. 3rd ed.}}  \newblock {Wiley Series in
-  Probability and Statistics.}, 1995.
-\bibitem[Imm12]{Immler2012} F.~Immler.  \newblock {\em {Generic
-    Construction of Probability Spaces for Paths of Stochastic
-    Processes in Isabelle/HOL.}} \newblock {Master’s Thesis in
-  Informatik, TU Munich},
-  \url{https://saloranta.de/immler/fabian/mastersthesis/thesis.pdf},
-  2012.
-\bibitem[Kal20]{Kallenberg2020} O.~Kallenberg.  \newblock {\em
-  {Foundations of modern probability. 3rd ed.}}  \newblock
-  {Probability and Its Applications. New York, NY: Springer.}, 2020.
-\bibitem[Kle13]{Klenke2013} A.~Klenke.  \newblock {\em {Probability
-    Theory. A Comprehensive Course. 2nd ed.}}  \newblock {Springer.},
-  2013.
-\end{thebibliography}

### 2023-09-10 Manuscript/proof_outline.tex
updates on the manuscript

@@
   defined on $\mathcal A := \bigcup_{J \subseteq_f \iota} \mathcal
-  F_J$ given by $P(A) = P_J(A)$ for $A\in\mathcal F_J$. Then, there is
-  a unique extension of $P$ to $\sigma(\mathcal A)$.
+  F_J$ given by $P(A) = P_J(\pi_J A)$ for $A\in\mathcal F_J$. Then,
+  there is a unique extension of $P$ to $\sigma(\mathcal A)$.
 \end{theorem}
@@
 In the proof of Kolmogorov's extension theorem, we use a closely
-related definition of \leanline{Measure.ofAddContent}, based on
-\leanline{ [mα : MeasurableSpace α] (hC : SetSemiring C) (hC_gen :
+related definition of \leanline{Measure.ofAddContent}, with variables
+\leanline{[mα : MeasurableSpace α] (hC : SetSemiring C) (hC_gen :
   MeasurableSpace.generateFrom C = mα)}, i.e.\ the semi-ring
@@
 \subsection{$\sigma$-additivity of set functions}
-For the proof of Theorem~\ref{T1}, note that $P$ as given in the
-result is a finite additive set function on the ring $\mathcal A$. In
-order to use Theorem~\ref{T:masseind}, we therefore have to show
-$\sigma$-additivity. For this, we will use Lemma~\ref{l:stetigcompact}
-below, i.e.\ inner regularity of $P$ with respect to compact
-sets. Before, we will show useful alternative conditions for
-$\sigma$-additivity, which do not make use of any topological
-structure of the underlying space.
+For the proof of Kolmogorov's extension Theorem, note that $P$ as
+given in Theorem~\ref{T1} is a finite additive set function on the
+ring $\mathcal A$. In order to use Theorem~\ref{T:masseind}, we
+therefore have to show $\sigma$-additivity. For this, we will use
+Lemma~\ref{l:stetigcompact} below, i.e.\ inner regularity of $P$ with
+respect to compact sets. Before, we will show useful alternative
+conditions for $\sigma$-additivity, which do not make use of any
+topological structure of the underlying space.
 
@@
     \item $m$ is continuous from below, i.e.\ for $A, A_1, A_2,... \in
-      \mathcal R$ with $A_1 \subseteq A_2 \subseteq \cdots$, we have
-      $m(A) = \lim_{n\to\infty} m(A_n)$.
+      \mathcal R$ with $A_1 \subseteq A_2 \subseteq \cdots$ and $A =
+      \bigcup_{n=1}^\infty A_n$, we have $m(A) = \lim_{n\to\infty}
+      m(A_n)$.
     \item $m$ is continuous from above in $\emptyset$, i.e.\ for $A_1,
@@
   (with $A_0 := \emptyset$)
-  $$ m(A) = m\Big( \biguplus_{n=1}^\infty A_n \setminus
-  \Big(\bigcup_{l=1}^{k-1} A_l\Big)\Big) = \lim_{n\to\infty} m\Big(
-  \biguplus_{k=1}^n A_k \setminus \Big(\bigcup_{l=1}^{k-1}
-  A_l\Big)\Big) = \lim_{n\to\infty} m(A_n).$$
+  \begin{align*}
+     m(A) & % = m\Big( \biguplus_{n=1}^\infty \Big(A_n \setminus
+     %\Big(\bigcup_{l=1}^{k-1} A_l\Big)\Big)\Big)
+     = \lim_{n\to\infty}
+     m\Big( \biguplus_{k=1}^n A_k \setminus \Big(\Big(\bigcup_{l=1}^{k-1}
+     A_l\Big)\Big)\Big) = \lim_{n\to\infty} m(A_n).
+  \end{align*}
 
@@
   \bigcup_{k=1}^n A_k$, such that $B_1, B_2,...$ satisfy 3. Hence,
-  $$ m\Big( \biguplus_{n=1}^\infty A_n\Big) = m\Big(
+  \begin{align*}
+    m\Big( \biguplus_{n=1}^\infty A_n\Big) & = m\Big(
     \bigcup_{n=1}^\infty B_n\Big) = \lim_{n\to\infty} m\Big(
-  \bigcup_{k=1}^n B_n\Big) = \lim_{n\to\infty} m\Big(
-  \biguplus_{k=1}^n A_n\Big) = \sum_{n=1}^\infty m(A_n).$$
+    \bigcup_{k=1}^n B_n\Big) \\ & = \lim_{n\to\infty} m\Big(
+    \biguplus_{k=1}^n A_n\Big) = \sum_{n=1}^\infty m(A_n).
+  \end{align*}
 
@@
 As an example, we give the formalization of $4.\Rightarrow 1.$:
+
+xxx Rename the todo!
 
@@
       Antitone s → (⋂ n, s n) = ∅ →
-      Tendsto (fun n => m (s n) (hs n)) atTop (𝓝 0))
+      Tendsto (fun n => m (s n) (hs n)) atTop (nhds 0))
       ⦃f : ℕ → Set α⦄ (h : ∀ i, f i ∈ C) (hUf : (⋃ i, f i) ∈ C)
@@
   \item If $q(A) = \text{true}$ iff $A$ is measurable, we neglect the
-    {\em and $q$}. If $p(A) = \text{true}$ iff $A$ is closed (ompact,
+    {\em and $q$}. If $p(A) = \text{true}$ iff $A$ is closed (compact,
     closed and compact), we say that $m$ is inner regular with respect
@@
 $\bigcap_{n=1}^\infty C_n = \emptyset$, there is some $N$ with
-$\bigcap_{n=1}^N C_n = \emptyset$.
+$\bigcap_{n=1}^N C_n = \emptyset$. More generally, compact sets form a
+compact system, which is defined as follows:
+
+\begin{definition}
+  Let $\mathcal C \subseteq 2^\alpha$. If, for all $C_1, C_2,...$ with
+  $\bigcap_{n=1}^\infty C_n = \emptyset$, there is some $N$ with
+  $\bigcap_{n=1}^N C_n = \emptyset$, we call $\mathcal C$ a {\em
+    compact system}.
+\end{definition}
 
@@
   function on a ring $\mathcal R$ contained by the Borel
-  $\sigma$-algebra, and which is inner regular with respect to compact
-  sets. Then, $\mu$ is $\sigma$-additive.
+  $\sigma$-algebra, and which is inner regular with respect to a
+  compact system. Then, $\mu$ is $\sigma$-additive.
 \end{lemma}
@@
   $\mu$ in $\emptyset$, so let $A_1 \supseteq A_2 \supseteq... \in
-  \mathcal R$ satisfy $\bigcap_{n=1}^\infty = \emptyset$ and
-  $\varepsilon>0$. Let $\delta_1, \delta_2,..>0$ with
-  $\sum_{n=1}^\infty \delta_n < \varepsilon$. For each $n$, let $C_n
-  \subseteq A_n \in \mathcal R$ be compact with $\mu(A_n) \leq
-  \mu(C_n) + \delta_n$. We have that $\bigcap_{n=1}^\infty C_n
+  \mathcal R$ satisfy $\bigcap_{n=1}^\infty A_n = \emptyset$ and
+  $\varepsilon>0$. Let $\delta_1, \delta_2,... >0$ with
+  $\sum_{n=1}^\infty \delta_n < \varepsilon$. For each $n$, let
+  $\mathcal C \ni C_n \subseteq A_n \in \mathcal R$ with $\mu(A_n)
+  \leq \mu(C_n) + \delta_n$. We have that $\bigcap_{n=1}^\infty C_n
   \subseteq \bigcap_{n=1}^\infty A_n = \emptyset$, so there is $N \in
-  \mathbb N$ with $\bigcap_{n=1}^N C_n = \emptyset$. So, for any
-  $m>N$, we have that
-  $$ \mu(A_m) = \mu\Big( \Big(\bigcap_{n=1}^m A_n\Big) \setminus
+  \mathbb N$ with $\bigcap_{n=1}^N C_n = \emptyset$ since $\mathcal C$
+  is a compact system. So, for any $m>N$, we have that
+  \begin{align*}
+     \mu(A_m) & = \mu\Big( \Big(\bigcap_{n=1}^m A_n\Big) \setminus
      \Big(\bigcap_{n=1}^m C_n\Big)\Big) \leq \sum_{n=1}^m \mu(A_n
-  \setminus C_n) \leq \sum_{n=1}^m \delta_n < \varepsilon.$$ This
-  concludes the proof.
+     \setminus C_n) \\ & \leq \sum_{n=1}^m \delta_n < \varepsilon.
+  \end{align*}
+  This concludes the proof.
 \end{proof}
 
-Here is the formalization using a \leanline{kol_content}:
+xxx Work on this, of time permits!
+
+In our formalization, we are less general and use a direct argument
+for a \leanline{kol_content}. More precisely, we show that a
+\leanline{kolContent}, which is based on a projective family which is
+inner regular with respect to compact and closed sets is
+$\sigma$-additive.
 
@@
 
-\sloppy Using this, we also have
-\leanline{kolContent_countably_subadditive_of_innerRegular} by using
-\leanline{countably_subadditive_of_countably_additive}.
-
-Recall that a finite measure on the Borel $\sigma$-algebra is inner
-regular with respect to closed sets, but also outer regular with
-respect to open sets.
+For using the more general version, we would need to show that
+$\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact}\}$ as
+well as $\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact
+  and closed}\}$ are compact systems. (Note that compact sets are
+closed in Hausdorff spaces, but we do not have this property since we
+are working with pseudo-metric spaces.) In addition, the assumption
+\leanline{hP_inner} from above translates directly to inner regularity
+of the \leanline{kolContent}, which is defined through the projective
+family \leanline{P}. Then, Lemma~\ref{l:stetigcompact} gives the same
+result as the formalization. Moreover, since the Carathéodory
+extension theorem requires $\sigma$-subadditivity (rather than
+$\sigma$-additivity), we can use
+\leanline{countably_subadditive_of_countably_additive} in order to
+show \leanline{kolContent_countably_subadditive_of_innerRegular},
+which shows that the assumptions of the last result also imply
+$\sigma$-subadditivity.
+
+What remains to be done is to show conditions under which the
+projective family is inner regular with respect to compact (and
+closed) sets. For this, we need some assumption on the underlying
+spaces. The following lemma provides the generalizations to extended
+pseudo-metric spaces:
+
+The next result is already implemented in \leanline{mathlib}: Use
+\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
+that every open set is inner regular with respect to closed sets, and
+\leanline{Measure.InnerRegular.measurableSet_of_open} to show that
+every measurable set is inner regular with respect to closed sets.
+The claimed outer reularity is part of
+\leanline{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.
 
@@
   with respect to closed sets. In fact, we have also outer regularity
-  with respect to open sets, i.e.
+  with respect to open sets, i.e.\ for all $A\in \mathcal B$,
   $$ \mu(A) = \inf \{ \mu(O) : A \subseteq O \text{ open}\}.$$
@@
   \delta_n$. Then, $O := \bigcup_{n=1}^\infty O_n$ is open and
-  $$ 0 \leq \mu(O \setminus A) = \mu(O) - \mu(A) \leq
-  \sum_{n=1}^\infty \mu(O_n \setminus A) \leq \sum_{n=1}^\infty
-  \mu(O_n \setminus A_n) < \varepsilon.$$ This shows outer regularity
-  of $\mu$ with respect to open sets at $A$. It remains to show inner
-  regularity with respect to closed sets. For this, note that
-  $\bigcup_{n=1}^N F_n$ is closed for all $N$ and recall from
-  Proposition~\ref{P:stetigmass} that $\mu$ is continuous from
-  below. Therefore, setting $F := \bigcup_{n=1}^\infty F_n$,
-  $$ 0 \leq \mu(A) - \lim_{N\to\infty} \mu\Big(\bigcup_{n=1}^N
+  $$ 0 \leq \mu(O) - \mu(A) \leq \sum_{n=1}^\infty \mu(O_n \setminus
+  A) \leq \sum_{n=1}^\infty \mu(O_n \setminus A_n) < \varepsilon.$$
+  This shows outer regularity of $\mu$ with respect to open sets at
+  $A$. It remains to show inner regularity with respect to closed
+  sets. For this, note that $\bigcup_{n=1}^N F_n$ is closed for all
+  $N$ and recall from Proposition~\ref{P:stetigmass} that $\mu$ is
+  continuous from below. Therefore, setting $F := \bigcup_{n=1}^\infty
+  F_n$,
+  \begin{align*}
+    0 & \leq \mu(A) - \lim_{N\to\infty} \mu\Big(\bigcup_{n=1}^N
     F_n\Big) = \mu(A \setminus F) \leq \sum_{n=1}^\infty \mu(A_n
-  \setminus F) \leq \sum_{n=1}^\infty \mu(A_n \setminus F_n) <
-  \varepsilon.$$ This shows inner regularity of $\mu$ with respect to
-  closed sets at $A$.
+    \setminus F) \\ & \leq \sum_{n=1}^\infty \mu(A_n \setminus F_n) <
+    \varepsilon.
+  \end{align*}
+  This shows inner regularity of $\mu$ with respect to closed sets at
+  $A$.
 
@@
   \subseteq \alpha$, $A^\varepsilon := \{x : \exists y \in A,
-  r(x,y<\varepsilon\}$. Then, for $A$ closed, we have that $A =
+  r(x,y)<\varepsilon\}$. Then, for $A$ closed, we have that $A =
   \bigcap_{n=1}^\infty A^{\varepsilon_n}$ and $A^{\varepsilon_n}$ is
   open for all $n$. Clearly, $\mu(A) = \sup\{\mu(F): F\text{ closed},
-  F \subseteq A\}$. By continuity of $\mu$, we find that
-  $\mu(A^{\varepsilon_n}) \xrightarrow{n\to\infty} \mu(A)$, therefore
-  $\mu(A) = \inf\{\mu(O): F\text{ open}, A \subseteq O\}$, and we have
-  shown that $A \in \mathcal A$.
+  F \subseteq A\}$, since $A$ is closed. By continuity of $\mu$, we
+  find that $\mu(A^{\varepsilon_n}) \xrightarrow{n\to\infty} \mu(A)$,
+  therefore $\mu(A) = \inf\{\mu(O): F\text{ open}, A \subseteq O\}$,
+  and we have shown that $A \in \mathcal A$.
 \end{proof}
-
-For the formalization, we use a new result: \leanline{P.InnerRegular
-  (fun s => IsCompact s ∧ IsClosed s) IsClosed}, which holds for every
-second countable extended pseudo metric space, which is
-\leanline{OpensMeasurable}. Then, from \leanline{mathlib}, we use
-\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
-that every open set is inner regular with respect to closed sets, and
-\leanline{Measure.InnerRegular.measurableSet_of_open} to show the
-above lemma.
 
@@
 In the next lemma, we will need that a closed subset of a compact set
-is compact. (See \leanline{isCompact_of_isClosed_subset}.)
+is compact. (See \leanline{isCompact_of_isClosed_subset}.) We describe
+the formalization of the next lemma only in conjunction with
+Lemma~\ref{L:relcoPol}.
 
@@
   lemma. In addition, by $\sigma$-sub-additivity of $\mu$,
-  $$ \mu((\overline A)^c) \leq \mu(A^c) = \mu\Big(
+  \begin{align*}
+     \mu((\overline A)^c) & \leq \mu(A^c) = \mu\Big(
      \bigcup_{n=1}^\infty \Big(\Big(\bigcup_{k=1}^{N_n}
-  B_{\varepsilon_n}(x_k)\Big)^c\Big)\Big) \leq \sum_{n=1}^\infty
-  \mu\Big( \Big(\bigcup_{k=1}^{N_n}
-  B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon,$$ and we are done.
+     B_{\varepsilon_n}(x_k)\Big)^c\Big)\Big) \\ & \leq
+     \sum_{n=1}^\infty \mu\Big( \Big(\bigcup_{k=1}^{N_n}
+     B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon,
+  \end{align*}
+ and we are done.
 \end{proof}
 
-Let us combine the last two results in one combined formalized result:
+Let us combine the last two results in one combined formalized result.
+This the gives the desired result that a measure on a second
+countable, complete extended pseudo-metric space is inner regular with
+respect to closed compact sets.
 
@@
 
-\noindent
+%\begin{lemma}\label{l:totalBoundcompact}
+%  Let $E$ be a uniform topologocal space\footnote{This is more general
+%    than a metric space, i.e.\ every metric space is uniform.} Then,
+%  $A\subseteq E$ is compact iff it is totally bounded\footnote{In a
+%    metric space, this means that forall $\varepsilon>0$ you find a
+%    finite number of balls covering the set} and
+%  complete\footnote{i.e.,\ Cauchy filters converge.}.
+%\end{lemma}
+
+%\noindent
+%The usual way to apply this result is in a Polish space. Here, $A$ is
+%relatively compact iff it is totally bounded.
+
+%\noindent
+%Although, regularity of measures is covered in {\tt mathlib}, the next
+%statement is not yet part of the library. Similar constructions will
+%be important for Prohorov's theorem in the theory of weak convergence.
+%In Lean, we can use \verb+ measure_theory.measure.inner_regular+ for
+%the statement of this result. In the proof of the lemma, we will be
+%using the last fact
+
+\subsection{Proof of Kolmogorov's extension theorem}
 With the above results, the strategy of the proof of Theorem~\ref{T1}
@@
 
-%\begin{lemma}\label{l:totalBoundcompact}
-%  Let $E$ be a uniform topologocal space\footnote{This is more general
-%    than a metric space, i.e.\ every metric space is uniform.} Then,
-%  $A\subseteq E$ is compact iff it is totally bounded\footnote{In a
-%    metric space, this means that forall $\varepsilon>0$ you find a
-%    finite number of balls covering the set} and
-%  complete\footnote{i.e.,\ Cauchy filters converge.}.
-%\end{lemma}
-
-%\noindent
-%The usual way to apply this result is in a Polish space. Here, $A$ is
-%relatively compact iff it is totally bounded.
-
-%\noindent
-%Although, regularity of measures is covered in {\tt mathlib}, the next
-%statement is not yet part of the library. Similar constructions will
-%be important for Prohorov's theorem in the theory of weak convergence.
-%In Lean, we can use \verb+ measure_theory.measure.inner_regular+ for
-%the statement of this result. In the proof of the lemma, we will be
-%using the last fact
-
-\subsection{Proof of Kolmogorov's extension theorem}
 Now we can put everything together: According to
@@
 \section{Concluding thoughts}
+We will highlight some paths we imagine can be taken, leading to
+further extensions of \leanline{mathlib}, which build on our
+contribution.
+
 {\em Using semi-rings in measure theory:} Frequently, the construction
@@
 set-function on a semi-ring by an outer measure, not previously
-implemented in mathlib. This could also be used to re-define
-Stieltjes measures (in particular Lebesgue measure) on $\mathbb R$, as
-well as product measures (on finite products). More precisely,
+implemented in mathlib. This could also be used to re-define Stieltjes
+measures (in particular Lebesgue measure) on $\mathbb R$, as well as
+product measures (on finite products). More precisely,
 e.g.\ \leanline{StieltjesFunction.outer_Ioc} and
@@
 
-xxx design decisions?
+~~
 

### 2023-09-11 Manuscript/proof_outline.tex
update manuscript

@@
 \maketitle
-
-{\color{blue}
-  \noindent
-  TODO
-  \begin{itemize}
-  \item Make sure that the line of argument for code and mauscript are in line.
-  \end{itemize}
-}
 
@@
   Let $\mathcal H$ be a semi-ring and $A_1,...,A_m \in \mathcal
-  H$. Then, there are $\mathcak K_1,...,\mathcal K_m \subseteq_f
+  H$. Then, there are $\mathcal K_1,...,\mathcal K_m \subseteq_f
   \mathcal H$ disjoint such that $\bigcup_{n=1}^m \mathcal K_n$
@@
 In the proof of Kolmogorov's extension theorem, we use a closely
-related definition of \leanline{Measure.ofAddContent}, with variables
-\leanline{[mα : MeasurableSpace α] (hC : SetSemiring C) (hC_gen :
-  MeasurableSpace.generateFrom C = mα)}, i.e.\ the semi-ring
-\leanline{C} generates the $\sigma$-algebra, and which uses an
-\leanline{m : AddContent} as well as \leanline{m_sigma_subadd} as
-above.
+related definition:
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+noncomputable def Measure.ofAddContent
+  [mα : MeasurableSpace α] (hC : SetSemiring C)
+  (hC_gen : MeasurableSpace.generateFrom C = mα)
+  (m : AddContent C)
+  (m_sigma_subadd : ...) :
+  Measure α
+\end{minted}
+with \leanline{m_sigma_subadd} as above.
 
@@
   \end{align*}
-
-  ~
 
@@
   kolContent hP (⋃ i, f i) = ∑' i, kolContent hP (f i)
+
+  theorem kolContent_countably_subadditive_of_innerRegular
+    (hP : IsProjectiveMeasureFamily P)
+    (hP_inner : ∀ J,
+      (P J).InnerRegular (fun s => IsCompact s ∧ IsClosed s)
+      MeasurableSet)
+    ⦃f : ℕ → Set (∀ i, α i)⦄
+    (hf : ∀ i, f i ∈ cylinders α)
+    (hf_Union : (⋃ i, f i) ∈ cylinders α) :
+    kolContent hP (⋃ i, f i) ≤ ∑' i, kolContent hP (f i)
 \end{minted}
@@
 We apply this result by using that 1.\ is satisfied for complete and
-separable extended pseudo-metric spaces:
+separable extended pseudo-metric spaces. Here, for the generalization
+to extended pseudo-metric spaces, we use that any subset of such a
+space is compact iff it is complete and totally bounded (see
+\leanline{isCompact_iff_totallyBounded_isComplete} in {\tt
+  mathlib}). Since closed subsets of a complete space are complete,
+the closure of a totally bounded set is (still totally bounded, hence)
+compact. (This fact is used in the proof below.)
 
@@
 \subsection{Proof of Kolmogorov's extension theorem}
-With the above results, the strategy of the proof of Theorem~\ref{T1}
-is as follows: We have to show that under the given conditions,
-condition~1 of Lemma~\ref{l:tight} holds. By the equivalent~2., we
-find that $P$ is inner regular with respect to closed compact sets. In
-particular, the conditions of Lemma~\ref{l:stetigcompact} are
-satisfied, showing that $P$ is $\sigma$-additive. Then, we proceed as
-described right before Proposition~\ref{P:stetigmass}. So, it remains
-to show condition~1 of Lemma~\ref{l:tight} for complete and separable
-extended pseudo-metric spaces. Note that -- as a special case of
-\leanline{isCompact_iff_totallyBounded_isComplete} from {\tt
-  mathlib}, any subset of such a space is compact iff it is complete
-and totally bounded. Since closed subsets of a complete space are
-complete, the closure of a totally bounded set is (still totally
-bounded, hence) compact.
-
-Now we can put everything together: According to
-Lemma~\ref{L:relcoPol}, condition~1 of Lemma~\ref{l:tight} is
-satisfied for the set function $P$ on the ring $\mathcal A$ from
-Theorem~\ref{T1}. The same lemma shows that the conditions for
-Lemma~\ref{l:stetigcompact} are satisfied, so $P$ is $\sigma$-additive
-on $\mathcal A$. As a consequence, we can use the Carathéodory
-extension from Theorem~\ref{T:masseind}. This shows all other
-assertions. The formalization of this proof resembles these
-arguments. We leave out all instances in the reformulation of the
-result and its proof (see below Theorem~\ref{T1} for a full
-formulation):
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def projectiveLimitWithWeakestHypotheses
-    [∀ i, PseudoEMetricSpace (α i)]
-    [∀ i, BorelSpace (α i)]
-    [∀ i, TopologicalSpace.SecondCountableTopology (α i)]
-    [∀ i, CompleteSpace (α i)] [Nonempty (∀ i, α i)]
+We now describe the proof of Kolmogorov's extension theorem as well as
+its formalization: For the proof of Theorem~\ref{T1}, we 
+\begin{enumerate}
+  \item apply Theorem~\ref{T:masseind} for the ring (hence semi-ring)
+    $\mathcal A$ and the set-function $P$ as given in
+    Theorem~\ref{T1};
+  \item show $\sigma$-additivity (as assumed in
+    Theorem~\ref{T:masseind}) of $P$ by using
+    Lemma~\ref{l:stetigcompact}. For the latter, we need to show that
+    $P$ is inner regular with respect to a compact system. Here, note
+    that $\{\pi_J^{-1}C : C \in \prod_{j\in J} \alpha_j \text{ compact
+      and closed}\}$ is a compact system and $P(\pi_J^{-1}C) =
+    P_J(C)$, so we need to show that $P_J$ is inner regular with
+    respect to compact and closed sets, $J\subseteq_f \iota$;
+  \item use Lemma~\ref{L:relcoPol} in combination with
+    Lemma~\ref{l:tight} (1.$\Rightarrow$2.) and the properties of
+    completeness, separability of the underlying extended
+    pseudo-metric spaces in order to see that every $P_J$ has the
+    desired property of being inner regular wrt compact and closed
+    sets. This concludes the proof of Theorem~\ref{T1}. 
+\end{enumerate}
+The formalization of this proof resembles these arguments. We leave
+out all instances in the reformulation of the result and its proof
+(see below Theorem~\ref{T1} for a full formulation):
+
+\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+noncomputable def
+  projectiveLimitWithWeakestHypotheses
   (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
-    [∀ i, IsFiniteMeasure (P i)]
   (hP : IsProjectiveMeasureFamily P) :
@@
 \end{minted}
+Let us look closer at the formalized proof:
+\leanline{kolContent_countably_subadditive_of_innerRegular} has the
+same hypothesis as connected to
+\leanline{kolContent_sigma_additive_of_innerRegular}, except
+\leanline{h_disj}. So, the content of the last bracket shows
+$\sigma$-sub-additivity of the content $P$ from
+Theorem~\ref{T1}. Then, \leanline{generateFrom_cylinders (kolContent
+  hP)} gives the \leanline{MeasurableSpace} on which we define the
+\leanline{addContent}.
 
@@
 \noindent
-{\em Prohorovs theorem:} The proof that single finite measures on a Polish space is tight (i.e.\ inner regular with respect to compact sets) is a special case of Prohorov's theorem, which states that a set $\mathcal M$ of finite measures is relatively compact (in the weak topology on the set of finite measures, i.e.\ every sequence of measures in the set has a weakly convergent sub-sequence) if and only if $\mathcal M$ is tight (i.e.\ for all $\varepsilon>0$ there is some compact $K$ such that $\sup_{m \in \mathcal M} m(K^c) < \varepsilon$).
-We treat here the special case of $\mathcal M$ being a singleton (hence is a compact set).
+{\em Prohorovs theorem:} The proof that single finite measures on a
+Polish space is tight (i.e.\ inner regular with respect to compact
+sets) is a special case of Prohorov's theorem, which states that a set
+$\mathcal M$ of finite measures is relatively compact (in the weak
+topology on the set of finite measures, i.e.\ every sequence of
+measures in the set has a weakly convergent sub-sequence) if and only
+if $\mathcal M$ is tight (i.e.\ for all $\varepsilon>0$ there is some
+compact $K$ such that $\sup_{m \in \mathcal M} m(K^c) <
+\varepsilon$). We treat here the special case of $\mathcal M$ being a
+singleton (hence is a compact set).
 

### 2023-09-11 Manuscript/proof_outline.tex
make small changes, add todos

@@
 %\usepackage{amsmath}
+
+\usepackage[textsize=tiny]{todonotes}
 
@@
 \begin{abstract}
-  We present a formalization of the Kolmogorov extension theorem,
-  which is a main building block in construction the (distribution of)
-  stochastic processes with arbitrary index sets in Lean. Our
+  We present a formalization in Lean of the Kolmogorov extension theorem,
+  which is a main building block in the construction of the (distribution of)
+  stochastic processes with arbitrary index sets. Our
   approach is based on mathlib, the mathematical library for Lean. On
@@
   allows us to construct the (distribution of) a stochastic process on
-  a complete and separable pseudo-metric space (rather than a metric
-  space).
+  a complete and separable pseudo-metric space rather than a metric
+  space.
 \end{abstract}
@@
 \maketitle
+
+\todo[inline]{Clearly differentiate refs to mathlib vs refs to new lemmas?}
+
+\todo[inline]{Mention a precise commit of Mathlib}
+
+\todo[inline]{write about sets or types?}
+
+\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}
+
+\todo[inline]{omit all noncomputable? Otherwise we need to explain what it means}
 
@@
 The resulting family of probability measures $(P_J)_{J \subseteq \iota \text{ finite}}$ has to be {\em projective} in the sense that the projection of $P_J$ to $H\subseteq J$ has to be equal to $P_H$.
-In other words, when describing the distribution of the stochastic process at all times in $J$, and then forgetting (i.e.\ projecting) all properties for times in $J\setminus H$, results in the description of properties at times in $H$.
+In other words, when describing the distribution of the stochastic process at all times in $J$, and then forgetting all properties for times in $J\setminus H$, results in the description of properties at times in $H$.
 One may then ask if this already gives a complete description of the process for all times.
@@
 However, it is the achievement of Kolmogorov that the finite-dimensional distributions in fact provide a unique description of the distribution of a stochastic process, as long as the underlying family of state spaces $(\alpha_t)_{t\in\iota}$ is nice enough (Polish, i.e.\ a separable topological space which can be metrized by a complete metric, for example).
-Assuming a family of topological space $(\alpha_t)_{t\in\iota}$, this distribution is a measure on the product-$\sigma$-field $\mathcal F :=\bigotimes_{t\in\iota} \mathcal B(\alpha_t)$ (where $\mathcal B(\alpha_t)$ is the Borel $\sigma$-algebra on $\alpha_t$).
+Assuming a family of topological spaces $(\alpha_t)_{t\in\iota}$, this distribution is a measure on the product-$\sigma$-field $\mathcal F :=\bigotimes_{t\in\iota} \mathcal B(\alpha_t)$ (where $\mathcal B(\alpha_t)$ is the Borel $\sigma$-algebra on $\alpha_t$).
 Here, $\mathcal F$ is generated by finite projections and hence any element of $\mathcal F$ may only depend on at most countably many $t\in \iota$, making this a rather coarse $\sigma$-algebra.
@@
 Here, we are going to use \leanline{Lean4}, since \leanline{mathlib} \cite{mathlib2020lean}, the mathematical library of \leanline{Lean} has recently been ported to \leanline{Lean4}.
-Subsequently, we aim to make a pull request for the publication of our code in \leanline{mathlib}.
+%Subsequently, we aim to make a pull request for the publication of our code in \leanline{mathlib}.
 The Kolmogorov extension theorem is on the interface between measure theory and probability theory.
@@
 So, our work complements \cite{ying2023formalization} in the sense that probability spaces $(\Omega, \mathcal F, \mathbf P)$ on which you can define a martingale exist.
-However, we do not give any specific example for the construction of a specific instance of such a space.
+However, we do not give any specific example for the construction of a specific instance of such a space\todo{could we give one?}.
 
@@
 \item a formalization of (the set-system of) semi-rings (see Definition~\ref{def:semi} below); 
-\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
+\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset\todo{for a measure, this is already in mathlib?} (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
 \item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below);
-\item the classical Carathéodory extension theorem, providing us a candidate for the measure which we want to construct (see Theorem~\ref{T:masseind}); 
+\item the classical Carathéodory extension theorem, providing us with a candidate for the measure which we want to construct (see Theorem~\ref{T:masseind}); 
 \item the Kolmogorov extension theorem, as based on the previous steps (See Theorem~\ref{T1}).
@@
 The Kolmogorov extension theorem has been previously formalized in Isabelle/HOL \cite{Immler2012}.
-This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical.
+This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical\todo{Lean has dependent types, Isabelle doesn't}.
 However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations in these two interactive theorem provers.
@@
 {\em Instances of stochastic processes:} An obvious application of Kolmogorov's extension theorem is the construction of basic stochastic processes like Markov chains \cite[Section~11]{Kallenberg2020}, the Poisson process \cite[Section~13]{Kallenberg2020} and Brownian motion \cite[Section~14]{Kallenberg2020}.
-(Note that other Gaussian processes -- indexed by $[0, \infty)$ or $\mathbb R$ -- might as well be constructed the same way.
-In addition, fields like the Gaussien free field indexed by $\mathbb R^d$ \cite{sheffield2007gaussian} can also be given.) For these tasks, we would have to define the finite-dimensional distributions (Poisson and normally distributed, respectively), and apply the extension theorem.
+Other Gaussian processes -- indexed by $[0, \infty)$ or $\mathbb R$ -- might as well be constructed the same way.
+In addition, fields like the Gaussien free field indexed by $\mathbb R^d$ \cite{sheffield2007gaussian} can also be given.
+For these tasks, we would have to define the finite-dimensional distributions (Poisson and normally distributed, respectively), and apply the extension theorem.
 This task requires the formalization of multi-dimensional Poisson and normal distributions, which -- in textbooks -- is usually done using characteristic functions.
@@
 {\em Sample-path properties:} The Kolmogorov extension gives the existence of a distribution of a stochastic process $(X_t)_{t\in\iota}$ with certain properties.
-Extra work is needed in order to show that -- on the same probability space -- we can as well define a version (i.e.\ another process $(Y_t)_{t\in\iota}$ with $\mathbf P(X_t = Y_t) = 1$ for all $t\in\iota$) which is right-continuous with left limits (for the Poisson process) or continuous (for Brownian motion), respectively.
+Extra work is needed in order to show that -- on the same probability space -- we can as well define a version (i.e.\ another process $(Y_t)_{t\in\iota}$ with $\mathbf P(X_t = Y_t) = 1$ for all $t\in\iota$) which is right-continuous with left limits (for the Poisson process) or continuous (for Brownian motion).
 For the former, this follows from some general principles of Markov processes (e.g.\ \cite[Theorem 4.3.6]{EthierKurtz1986}).
@@
 The proof of the latter, however, uses transition kernels (and an inductive construction of a content, which can be extended to a measure) rather than projective families.
-However, continuity of a measure at $\emptyset$ is used as well and a formalization would complement our work.
+However, continuity of a measure at $\emptyset$\todo{already in mathlib} is used as well and a formalization would complement our work.
 In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} without requirements on $(\alpha_t)_{t\in\iota}$.
@@
 \end{definition}
-
+\todo{reorder that next remark? I'll make a proposal}
 \begin{remark}[Metric and topological spaces in {\tt mathlib}] \mbox{}
@@
       PseudoEMetricSpace (α : Type u) extends EDist}, where
-      \leanline{EDist} comes with an extended distance \leanline{edist
-        : α → α → ENNReal}, and the extension to
+      \leanline{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}
+      (the notation \leanline{ℝ≥0∞} stands for $[0, \infty]$), and the extension to
       \leanline{PseudoEMetricSpace} consists of the components,
@@
       \cite{melikhov2011metrizable}.
-    \item \sloppy For topological spaces, we have \leanline{class
-      TopologicalSpace(α : Type u) : Type u}, which comes with
-      \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen
-        Set.univ}, \leanline{isOpen_inter} and
+    \item \sloppy For topological spaces, we have \leanline{class TopologicalSpace(α : Type u) : Type u}, which comes with
+      \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
       \leanline{isOpen_sUnion}, which are exactly the properties of a
@@
     on $2^\alpha$, whereas a measure is only defined on a subset (the
-    $\sigma$-algebra of measurable sets). More precisely: A
+    $\sigma$-algebra of measurable sets). More precisely\todo{I don't understand why all these details are here. We never use MeasureSpace}: A
     \leanline{MeasureSpace α} extends a \leanline{MeasurableSpace α}
@@
     measurable (\leanline{AEMeasurable}) if it coincices with some
-    \leanline{Measurable g}, \leanline{m}-almost everywhere.\\ In this
+    \leanline{Measurable g}, \leanline{m}-almost everywhere\todo{do we need aemeasurable? could it be omitted for simplicity?}.\\ In this
     case, if $\mu$ is a measure on $\mathcal F$, the measure $\nu:
@@
 
-We now come to our contribution, which we will PR to mathlib: The
-extension theorem is on finite measures on product spaces. In
+We now come to our contribution%, which we will PR to mathlib
+: the extension theorem is on finite measures on product spaces\todo{I don't understand this sentence}. In
 particular, the product is taken over a set of arbitrary
 cardinality. The next definition covers the important concept of a
-projective family of measures. In short term, we define measures on
-any finite subset of indices in a consistent way. Recall that finite
+projective family of measures. In short, we define measures on
+any finite subset of indices in a consistent way. Finite
 sets (of some type \leanline{ι}) are formalized with \leanline{Finset
@@
 quotient type of the \leanline{List}-type with respect to permutations
-of the list) together with a proof that it contains no duplicates.
+of the list) together with a proof that it contains no duplicates\todo{remove this implementation detail?}.
 
@@
 \noindent
-With this, we can define the projetive family as follows. Note that
-\leanline{Lean} automatically uses the subset relation as
+With this, we can define the projetive family as follows. The typeclass inference system of
+\leanline{Lean} automatically uses the subset relation to generate
 \leanline{[Preorder (Finset ι)]} when \leanline{isProjective} is
@@
     [∀ i, BorelSpace (α i)]
-    [∀ i, TopologicalSpace.SecondCountableTopology (α i)]
-    [∀ i, CompleteSpace (α i)] [Nonempty (∀ i, α i)]
+    [∀ i, SecondCountableTopology (α i)]
+    [∀ i, CompleteSpace (α i)] [∀ i, Nonempty (α i)]
     (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
@@
 Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
-open balls around $x$ and $y$ overlap.
+open balls around $x$ and $y$ overlap\todo{it would be nice to explain how the formalization led us to remove that hypothesis}.
 
@@
 \noindent
-Using \leanline{variable {α : Type _} {C : Set (Set α)} {s t : Set α}
+Any ring is a semi-ring since $A\cap B = A \setminus (A \setminus B)$,
+i.e.\ every ring is a $\pi$-system.
+Using \leanline{variable {α : Type _} {s t : Set α}
 }, we define a semi-ring as follows:
@@
 \noindent
-We do not show the formalization of rings here. Note, however, that
-any ring is a semi-ring since $A\cap B = A \setminus (A \setminus B)$,
-i.e.\ every ring is a $\pi$-system.
+We do not show the formalization of rings here.
 %In our formalization, we also use the notion of a field $\mathcal H$,
@@
 %\mathcal H$.
-Since we are dealing with product spaces, we will use cylinders, which
+Since we are dealing with product spaces, we will use cylinders\todo{R: rewrite this}, which
 are defined as follows:
@@
 From the field/ring/semi-ring of cylinders, we have to define the
-generated $\sigma$-alebra. This uses the following:
+generated $\sigma$-alebra. This uses the following\todo{one of those 2 is ours, the other is from mathlib. How do we show the difference?}:
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
 an element from an $\exists$-statement. In addition, we do not allow
-$\emptyset \in \mathcal K$ without loss of generality:
+$\emptyset \in \mathcal K$ without loss of generality\todo{that exclusion was useful for something. Find out what that is}:
 
@@
 members of a semi-ring as a finite union of disjoint sets.
+
+\todo[inline]{Is this section going into too much detail?}
 
@@
   \item If $\mathcal H$ is a $\sigma$-algebra and $m$ is
-    $\sigma$-additive, we call $m$ a measure.
+    $\sigma$-additive, we call $m$ a measure\todo{need 0 at the empty set? Otherwise could be infinite everywhere}.
   \item If $\mathcal H = 2^\alpha$ and $m$ is monotone and
@@
 \noindent
-For additive contents, we need two definitions. \leanline{AddContent}
+For additive contents\todo{"content" not defined?}, we need two definitions. \leanline{AddContent}
 defines a set-function \leanline{tuFun} on \leanline{Set α}, whereas
@@
 formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
-m(\emptyset)$. However, since \leanline{Finset}s are assumed to have
+m(\emptyset)$\todo{not if m is infinite everywhere}. However, since \leanline{Finset}s are assumed to have
 distinct elements, we cannot have \leanline{∅} twice in \leanline{I}.
 
-For the concrete application we have in mind, we need to be more
-specific which semiring to use, based on a projective family of
-probability measures. We do this using \leanline{kolContent}, which is
-based on \leanline{kolmogorovFun}, which takes a family of measures
-and a cylinder set as input:
+For the concrete application we have in mind, we introduce a definition which uses a specific semiring, the measurable cylinders, based on a projective family of measures.
+That is the goal of the two definitions \leanline{kolmogorovFun} and \leanline{kolContent} below:
 
@@
 
+\todo{cylinders.finset not defined?}
+
 Here, \leanline{kolmogorovFun_empty} and
@@
 well as $\sigma$-additive $\Rightarrow$ $\sigma$-sub-additive in the
-proof of the Carathéodory extension, Theorem~\ref{T:masseind}. The
+proof of the Carathéodory extension, Theorem~\ref{T:masseind}\todo{I don't understand this sentence}. The
 formalization of the next result comes in several lemmas, which we do
@@
   Next, we claim that for $\biguplus_{I\in\mathcal I} I \subseteq A$
-  (and all sets belong to $\mathcal H$, we have $\sum_{I \in \mathcal
+  with all sets belonging to $\mathcal H$, we have $\sum_{I \in \mathcal
     I} m(I) \leq m(A)$. For this, write $A \setminus \biguplus_{I\in
@@
 H$. All statements on equality of $\mu$ and $m$ present in {\tt
-  mathlib} at the time of writing have too many requirements.
-
-The next result talks about an \leanline{induced_outer_measure} in
-terms of {\tt mathlib}.
+  mathlib} at the time of writing have too many requirements: they all require $m$ to be a $\sigma$-additive function defined on a $\sigma$-algebra.
+
+%The next result talks about an \leanline{induced_outer_measure} in
+%terms of {\tt mathlib}.
+The next result extends a set function on a semi-ring\todo{check that we are consistend with semiring vs semi-ring} to an outer measure.
 
@@
   Then, $\mathcal F$ is a $\sigma$-Algebra and $\mu|_{\mathcal F}$ is
-  a measure. In addition, $\mathcal N:=\{N\subseteq \Omega:
+  a measure. In addition, $\{N\subseteq \Omega:
   \mu(N)=0\}\subseteq \mathcal F$, i.e.\ $\mathcal F$ is complete.
@@
 \leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
-\subseteq \mathcal F$ in a special case. In addition, there is
+\subseteq \mathcal F$ in a special case\todo{which special case?}. In addition, there is
 \leanline{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
@@
 \begin{proof}
-  First, note that $m$ is $\sigma$-sub-additive by
+  First, $m$ is $\sigma$-sub-additive by
   Lemma~\ref{l:halbextRing}.
@@
 \end{minted}
-
-\noindent
-Here, \leanline{MeasureTheory.OuterMeasure.caratheodory} is already
+\todo{is this the right lemma? Do we have a similar one about contents?}
+
+\noindent
+Here, \leanline{MeasureTheory.OuterMeasure.caratheodory}\todo{used in dot notation... but we don't want to explain this} is already
 implemented and gives the assertion that $\mathcal F$ from
@@
 In the proof of Kolmogorov's extension theorem, we use a closely
-related definition:
+related definition\todo{should we present only this one and omit the lemma above?}:
 
@@
 
-xxx Rename the todo!
+\todo[inline]{Rename the todo! And replace by the version about contents?}
 
@@
 \end{definition}
+\todo{no definition for this in the code for now}
 
@@
 
-xxx Work on this, of time permits!
+\todo[inline]{Work on this, if time permits!}
 
@@
 well as $\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact
-  and closed}\}$ are compact systems. (Note that compact sets are
+  and closed}\}$ are compact systems. Note that compact sets are
 closed in Hausdorff spaces, but we do not have this property since we
-are working with pseudo-metric spaces.) In addition, the assumption
+are working with pseudo-metric spaces. In addition, the assumption
 \leanline{hP_inner} from above translates directly to inner regularity
@@
 
+\todo{too much detail? omit the proof?}
+
 \begin{proof}
@@
   \]
-  which by construction is totally bounded (for $r>0$ take $n$ such
+  which by construction is totally bounded\todo{totally bounded not defined?} (for $r>0$ take $n$ such
   that $\varepsilon_n < r$, such that $B_r(x_1),...,B_r(x_{N_n})$
   cover $A$), and therefore has a compact closure as noted before the
-  lemma. In addition, by $\sigma$-sub-additivity of $\mu$,
+  lemma\todo{is it noted before the lemma? where?}. In addition, by $\sigma$-sub-additivity of $\mu$,
   \begin{align*}
@@
     [PseudoEMetricSpace α] [CompleteSpace α]
-    [TopologicalSpace.SecondCountableTopology α]
+    [SecondCountableTopology α]
     [BorelSpace α] (P : Measure α) [IsFiniteMeasure P] :
@@
 \leanline{kolContent_countably_subadditive_of_innerRegular} has the
-same hypothesis as connected to
+same hypotheses as
 \leanline{kolContent_sigma_additive_of_innerRegular}, except
-\leanline{h_disj}. So, the content of the last bracket shows
+\leanline{h_disj}. So, the lemma in the last brackets shows
 $\sigma$-sub-additivity of the content $P$ from
-Theorem~\ref{T1}. Then, \leanline{generateFrom_cylinders (kolContent
-  hP)} gives the \leanline{MeasurableSpace} on which we define the
+Theorem~\ref{T1}. Then, \leanline{generateFrom_cylinders} gives the \leanline{MeasurableSpace} on which we define the
 \leanline{addContent}.

### 2023-09-12 Manuscript/proof_outline.tex
fix leanline spacing

@@
   \begin{enumerate}
-    \item \sloppy In \leanline{Lean}, we have the \leanline{class
-      PseudoEMetricSpace (α : Type u) extends EDist}, where
+    \item \sloppy In \leanline{Lean}, we have the \leanline{class PseudoEMetricSpace (α : Type u) extends EDist}, where
       \leanline{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}
@@
       connects the \leanline{uniformity} with the topology using
-      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p |
-          p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
+      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
       uniform space defines a topological space, which can be used by
@@
     \leanline{measurableSet_empty : MeasurableSet' ∅},
-    \leanline{measurableSet_compl : (s : Set α) → MeasurableSet' s →
-      MeasurableSet' sᶜ}, and \leanline{measurableSet_iUnion},
+    \leanline{measurableSet_compl : (s : Set α) → MeasurableSet' s → MeasurableSet' sᶜ}, and \leanline{measurableSet_iUnion},
     i.e.\ properties (i)-(iii) from Definition~\ref{def:sigma}.1.
   \item We are using outer measures in our construction. In
-    \leanline{mathlib}, this is \leanline{structure
-      MeasureTheory.OuterMeasure α} which is a function
-    \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf
-      ∅ = 0} (i.e.\ the empty set has measure~$0$), \leanline{∀ {s₁ s₂
-        : Set α}, s₁ ⊆ s₂ → measureOf s₁ ≤ measureOf s₂}
-    (i.e.\ monotonicity) and \leanline{∀ (s : ℕ → Set α), measureOf (⋃
-      (i : ℕ), s i) ≤ ∑' (i : ℕ), measureOf (s i)} (which we call
+    \leanline{mathlib}, this is \leanline{structure MeasureTheory.OuterMeasure α} which is a function
+    \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf ∅ = 0} (i.e.\ the empty set has measure~$0$),
+    \leanline{∀ {s₁ s₂ : Set α}, s₁ ⊆ s₂ → measureOf s₁ ≤ measureOf s₂}
+    (i.e.\ monotonicity) and \leanline{∀ (s : ℕ → Set α), measureOf (⋃ (i : ℕ), s i) ≤ ∑' (i : ℕ), measureOf (s i)} (which we call
     $\sigma$-subadditivity). \\ Note that an outer measure is defined
@@
     by a \leanline{Measure α} (which is usually called
-    \leanline{volume}). For the measure, we have: \leanline{structure
-      MeasureTheory.Measure α [MeasurableSpace α] extends
-      MeasureTheory.OuterMeasure}. Here $\sigma$-additivity of a
+    \leanline{volume}). For the measure, we have: \leanline{structure MeasureTheory.Measure α [MeasurableSpace α] extends MeasureTheory.OuterMeasure}.
+    Here $\sigma$-additivity of a
     \leanline{Measure α} reads
@@
 any finite subset of indices in a consistent way. Finite
-sets (of some type \leanline{ι}) are formalized with \leanline{Finset
-  ι}. Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
+sets (of some type \leanline{ι}) are formalized with \leanline{Finset ι}. Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
 quotient type of the \leanline{List}-type with respect to permutations
@@
 statement) to the image measure on \leanline{J}, which is
-\leanline{μ.map (fun x : ∀ i : I, α i => fun j => x ⟨j, hJI
-    j.2⟩)}. The map defined by the \leanline{=>}-notation maps every
+\leanline{μ.map (fun x : ∀ i : I, α i => fun j => x ⟨j, hJI j.2⟩)}.
+The map defined by the \leanline{=>}-notation maps every
 $(x_i)_{i\in I}$ to a function of \leanline{j}, whose type is the
@@
 i.e.\ every ring is a $\pi$-system.
-Using \leanline{variable {α : Type _} {s t : Set α}
-}, we define a semi-ring as follows:
+Using \leanline{variable {α : Type _} {s t : Set α}}, we define a semi-ring as follows:
 
@@
 \leanline{↑}, which in this case is from \leanline{Finset} to
-\leanline{Set} and (ii) we have \leanline{PairwiseDisjoint (s : Set ι)
-  (f : ι → α)} iff the images of any distinct elements of \leanline{ι}
+\leanline{Set} and (ii) we have \leanline{PairwiseDisjoint (s : Set ι) (f : ι → α)}
+iff the images of any distinct elements of \leanline{ι}
 under \leanline{f} are different. (Hence, if \leanline{f = id}, the
@@
   n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
-is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin
-  (card J)) : Finset (Set α)} is the set consisting of the
+is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)}
+is the set consisting of the
 \leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
-\leanline{finsetLT_subset'} gives a proof that \leanline{↑(finsetLT J
-  n) ⊆ C}. So, in terms of Lemma~\ref{l2}, this gives, for
+\leanline{finsetLT_subset'} gives a proof that \leanline{↑(finsetLT J n) ⊆ C}.
+So, in terms of Lemma~\ref{l2}, this gives, for
 $n=1,...,m$, a construction for $\mathcal K_n \subseteq_f \mathcal H$.
@@
 defines a set-function \leanline{tuFun} on \leanline{Set α}, whereas
-\leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C
-  → ℝ≥0∞}.
+\leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.
 

### 2023-09-12 Manuscript/proof_outline.tex
remove noncomputable

@@
 \todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}
-
-\todo[inline]{omit all noncomputable? Otherwise we need to explain what it means}
 
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def
-    projectiveLimitWithWeakestHypotheses
+def projectiveLimitWithWeakestHypotheses
   [∀ i, PseudoEMetricSpace (α i)]
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def diff₀ (hC : SetSemiring C)
+def diff₀ (hC : SetSemiring C)
     (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def indexedDiff₀ (hC : SetSemiring C)
+def indexedDiff₀ (hC : SetSemiring C)
     (J : Finset (Set α)) (hJ : ↑J ⊆ C)
@@
 
-  noncomputable def extendContent
+def extendContent
     (hC : SetSemiring C)
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def kolmogorovFun
+def kolmogorovFun
     (P : ∀ J : Finset ι, Measure (∀ j : J, α j))
@@
 
-  noncomputable def kolContent
+def kolContent
     (hP : IsProjectiveMeasureFamily P) :
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  noncomputable def ofAddSubaddCaratheodory
+def ofAddSubaddCaratheodory
     (hC : SetSemiring C)
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-noncomputable def Measure.ofAddContent
+def Measure.ofAddContent
     [mα : MeasurableSpace α] (hC : SetSemiring C)
@@
 \begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-noncomputable def
-  projectiveLimitWithWeakestHypotheses
+def projectiveLimitWithWeakestHypotheses
     (P : ∀ J : Finset ι, Measure (∀ j : J, α j))

### 2023-09-12 Manuscript/proof_outline.tex
remove some details

@@
 \item a formalization of (the set-system of) semi-rings (see Definition~\ref{def:semi} below); 
-\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset\todo{for a measure, this is already in mathlib?} (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
+\item a definition of additive contents over semi-rings\todo{R: I added that point but we don't talk much about contents};
+%\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset\todo{for a measure, this is already in mathlib?} (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
 \item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below);
@@
 This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical\todo{Lean has dependent types, Isabelle doesn't}.
-However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations in these two interactive theorem provers.
+However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations\todo{no, the type theory is not the same} in these two interactive theorem provers.
 
@@
 \noindent
-{\em Theorem of Ionescu-Tulcea and related results:} While Kolmogorov's extension Theorem gives a result for arbitrary $\iota$, but requires some properties of the family $(\alpha_t)_{t\in\iota}$, the Theorem of Ionescu-Tulcea \cite[Theorem 8.24]{Kallenberg2020} can only deal with countable $\iota$, but has no restrictions on $(\alpha_t)_{t\in\iota}$.
+{\em Theorem of Ionescu-Tulcea and related results:}
+While Kolmogorov's extension Theorem gives a result for arbitrary $\iota$, but requires some properties of the family $(\alpha_t)_{t\in\iota}$, the Theorem of Ionescu-Tulcea \cite[Theorem 8.24]{Kallenberg2020} can only deal with countable $\iota$, but has no restrictions on $(\alpha_t)_{t\in\iota}$.
 The proof of the latter, however, uses transition kernels (and an inductive construction of a content, which can be extended to a measure) rather than projective families.
-However, continuity of a measure at $\emptyset$\todo{already in mathlib} is used as well and a formalization would complement our work.
+However, the structure of the argument is similar and could use some of our results, like the proof of $\sigma$-additivity from continuity at the empty set (see Lemma~\ref{P:stetigmass}).
+A formalization would complement our work.
 In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} without requirements on $(\alpha_t)_{t\in\iota}$.
@@
     on $2^\alpha$, whereas a measure is only defined on a subset (the
-    $\sigma$-algebra of measurable sets). More precisely\todo{I don't understand why all these details are here. We never use MeasureSpace}: A
-    \leanline{MeasureSpace α} extends a \leanline{MeasurableSpace α}
-    by a \leanline{Measure α} (which is usually called
-    \leanline{volume}). For the measure, we have: \leanline{structure MeasureTheory.Measure α [MeasurableSpace α] extends MeasureTheory.OuterMeasure}.
-    Here $\sigma$-additivity of a
-    \leanline{Measure α} reads
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  m_iUnion : ∀ ⦃f : ℕ → Set α⦄,
-    (∀ (i : ℕ), MeasurableSet (f i)) →
-    Pairwise (Disjoint on f) →
-    ↑toOuterMeasure (⋃ (i : ℕ), f i) =
-    ∑' (i : ℕ), ↑toOuterMeasure (f i)
-\end{minted}    
+    $\sigma$-algebra of measurable sets).
+  \item A measure defined on a type \leanline{α} on which we have a \leanline{MeasurableSpace α} is denoted by the type \leanline{MeasureTheory.Measure α}
+    %More precisely\todo{I don't understand why all these details are here. We never use MeasureSpace}: A
+    %\leanline{MeasureSpace α} extends a \leanline{MeasurableSpace α}
+    %by a \leanline{Measure α} (which is usually called
+    %\leanline{volume}). For the measure, we have: \leanline{structure MeasureTheory.Measure α [MeasurableSpace α] extends MeasureTheory.OuterMeasure}.
+    %Here $\sigma$-additivity of a
+    %\leanline{Measure α} reads
+%\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+%  m_iUnion : ∀ ⦃f : ℕ → Set α⦄,
+%    (∀ (i : ℕ), MeasurableSet (f i)) →
+%    Pairwise (Disjoint on f) →
+%    ↑toOuterMeasure (⋃ (i : ℕ), f i) =
+%    ∑' (i : ℕ), ↑toOuterMeasure (f i)
+%\end{minted}    
   \end{enumerate}
@@
     $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib},
-    see \leanline{Measurable} and note that this notion usually comes
-    with some \leanline{m : Measure α} and $f$ is almost everywhere
-    measurable (\leanline{AEMeasurable}) if it coincices with some
-    \leanline{Measurable g}, \leanline{m}-almost everywhere\todo{do we need aemeasurable? could it be omitted for simplicity?}.\\ In this
+    see \leanline{Measurable}.
+    %and note that this notion usually comes
+    %with some \leanline{m : Measure α} and $f$ is almost everywhere
+    %measurable (\leanline{AEMeasurable}) if it coincices with some
+    %\leanline{Measurable g}, \leanline{m}-almost everywhere\todo{do we need aemeasurable? could it be omitted for simplicity?}.
+    \\ In this
     case, if $\mu$ is a measure on $\mathcal F$, the measure $\nu:
@@
 any finite subset of indices in a consistent way. Finite
-sets (of some type \leanline{ι}) are formalized with \leanline{Finset ι}. Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
-quotient type of the \leanline{List}-type with respect to permutations
-of the list) together with a proof that it contains no duplicates\todo{remove this implementation detail?}.
+sets (of some type \leanline{ι}) are formalized with \leanline{Finset ι}. % Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
+%quotient type of the \leanline{List}-type with respect to permutations
+%of the list) together with a proof that it contains no duplicates\todo{remove this implementation detail?}.
 

### 2023-09-12 Manuscript/proof_outline.tex
minor rewrite
add leanlinem command to mark mathlib lemmas

@@
 \newcommand{\leanline}{\mintinline[breaklines, breakafter=., breaklines]{Lean}}
+\newcommand{\leanlinem}[1]{\mintinline[breaklines, breakafter=., breaklines]{Lean}{#1}\textsuperscript{\textcolor{gray}{\tiny(m)}}}
 %\usepackage{refcheck}
@@
 
-\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}
+\todo[inline]{\leanline{∀ j : J, α j} or \leanlinem{(j : J) -> α j} ?}
+
+\todo[inline]{explain leanlinem}
 
@@
   \begin{enumerate}
-    \item \sloppy In \leanline{Lean}, we have the \leanline{class PseudoEMetricSpace (α : Type u) extends EDist}, where
-      \leanline{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}
+    \item \sloppy In \leanline{Lean}, we have the \leanlinem{class PseudoEMetricSpace (α : Type u) extends EDist}, where
+      \leanlinem{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}
       (the notation \leanline{ℝ≥0∞} stands for $[0, \infty]$), and the extension to
-      \leanline{PseudoEMetricSpace} consists of the components,
+      \leanlinem{PseudoEMetricSpace} consists of the components,
       \leanline{edist_self}, \leanline{edist_comm},
@@
       pseudo-metric-space structure that generates the same
-      uniformity; see \leanline{UniformSpace.pseudoMetrizableSpace},
+      uniformity; see \leanlinem{UniformSpace.pseudoMetrizableSpace},
       which formalizes a result stated in
       \cite{melikhov2011metrizable}.
-    \item \sloppy For topological spaces, we have \leanline{class TopologicalSpace(α : Type u) : Type u}, which comes with
+    \item \sloppy For topological spaces, we have \leanlinem{class TopologicalSpace(α : Type u) : Type u}, which comes with
       \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
@@
       above, every extended pseudo-metric defines a uniform space, and
-      a uniform space is a \leanline{class UniformSpace(α : Type u)
+      a uniform space is a \leanlinem{class UniformSpace(α : Type u)
         extends TopologicalSpace , UniformSpace.Core} (where the
-      uniformity is defined in \leanline{UniformSpace.Core}), which
+      uniformity is defined in \leanlinem{UniformSpace.Core}), which
       connects the \leanline{uniformity} with the topology using
@@
       \mathcal H \text{ topology}} \mathcal F$; see
-    \leanline{TopologicalSpace.generateFrom}. This is called the
+    \leanlinem{TopologicalSpace.generateFrom}. This is called the
     topology generated by $\mathcal H$. If $\mathcal H$ is closed
     under finite intersections, we call $\mathcal H$ a basis for
-    $\mathcal O$; see \leanline{TopologicalSpace.IsTopologicalBasis}.
+    $\mathcal O$; see \leanlinem{TopologicalSpace.IsTopologicalBasis}.
   \item \sloppy We call the topology (generated from an extended
     pseudo-metric) separable (see
-    \leanline{TopologicalSpace.SeparableSpace}) if there is a
+    \leanlinem{TopologicalSpace.SeparableSpace}) if there is a
     countable $s \subseteq \iota$ such that $\inf\{r(x,y) : y \in s\}
     = 0$ for all $x\in\alpha$. (More generally, a
-    \leanline{TopologicalSpace α} is a \leanline{SeparableSpace} iff
+    \leanline{TopologicalSpace α} is a \leanlinem{SeparableSpace} iff
     \leanline{∃ s, Set.Countable s ∧ Dense s}, where the latter means
@@
     is separable; see
-    \leanline{TopologicalSpace.SecondCountableTopology.to_separableSpace}.
+    \leanlinem{TopologicalSpace.SecondCountableTopology.to_separableSpace}.
   \end{enumerate}
@@
   \begin{enumerate}
-  \item \sloppy The \leanline{class MeasurableSpace α} is very similar
+  \item \sloppy The \leanlinem{class MeasurableSpace α} is very similar
     to a topological space in \leanline{mathlib}, since it comes with
@@
   \item We are using outer measures in our construction. In
-    \leanline{mathlib}, this is \leanline{structure MeasureTheory.OuterMeasure α} which is a function
+    \leanline{mathlib}, this is \leanlinem{structure MeasureTheory.OuterMeasure α} which is a function
     \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf ∅ = 0} (i.e.\ the empty set has measure~$0$),
@@
     $\sigma$-algebra of measurable sets).
-  \item A measure defined on a type \leanline{α} on which we have a \leanline{MeasurableSpace α} is denoted by the type \leanline{MeasureTheory.Measure α}
+  \item A measure defined on a type \leanline{α} on which we have a \leanline{MeasurableSpace α} is denoted by the type \leanlinem{MeasureTheory.Measure α}
     %More precisely\todo{I don't understand why all these details are here. We never use MeasureSpace}: A
@@
     \mathcal F$. This is called the $\sigma$-algebra generated from
-    $\mathcal H$; see \leanline{MeasurableSpace.generateFrom}.  \\ In
+    $\mathcal H$; see \leanlinem{MeasurableSpace.generateFrom}.  \\ In
     particular, if $\mathcal O$ is a topology on $\alpha$, we call
@@
     $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib},
-    see \leanline{Measurable}.
+    see \leanlinem{Measurable}.
     %and note that this notion usually comes
@@
     the image (or push-forward) measure of $\mu$ under $f$. We write
-    $\nu := f_\ast \mu$; see \leanline{MeasureTheory.Measure.map}.
+    $\nu := f_\ast \mu$; see \leanlinem{MeasureTheory.Measure.map}.
   \end{enumerate}
@@
 any finite subset of indices in a consistent way. Finite
-sets (of some type \leanline{ι}) are formalized with \leanline{Finset ι}. % Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
+sets (of some type \leanline{ι}) are formalized with \leanlinem{Finset ι}. % Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
 %quotient type of the \leanline{List}-type with respect to permutations
@@
 \leanline{↑}, which in this case is from \leanline{Finset} to
-\leanline{Set} and (ii) we have \leanline{PairwiseDisjoint (s : Set ι) (f : ι → α)}
+\leanline{Set} and (ii) we have \leanlinem{PairwiseDisjoint (s : Set ι) (f : ι → α)}
 iff the images of any distinct elements of \leanline{ι}
@@
 %\mathcal H$.
-Since we are dealing with product spaces, we will use cylinders\todo{R: rewrite this}, which
+The import ring of sets in our formalization is the ring of measurable cylinders on a product space, which
 are defined as follows:
@@
 The existence-statement of the above lemma actually gives rise to a
-definition. Here, we use \leanline{Exists.choose} in order to extract
+definition. Here, we use \leanlinem{Exists.choose} in order to extract
 an element from an $\exists$-statement. In addition, we do not allow
@@
 
-Here, recall that \leanline{Fin n} is the subtype of \leanline{Nat}
+Here, recall that \leanlinem{Fin n} is the subtype of \leanline{Nat}
 consisting of all numbers \leanline{<n}. In addition,
@@
 For additive contents\todo{"content" not defined?}, we need two definitions. \leanline{AddContent}
-defines a set-function \leanline{tuFun} on \leanline{Set α}, whereas
+defines a set-function \leanline{toFun} on \leanline{Set α}, whereas
 \leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.
@@
 Carathéodory. The first leads to the definition of the measurable
-space \leanline{OuterMeasure.caratheodory}, covered in {\tt
+space \leanlinem{OuterMeasure.caratheodory}, covered in {\tt
   mathlib}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}
@@
   mathlib}. However, for product spaces, in
-\leanline{MeasureTheory.Constructions.Pi}, there is
-\leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
+\leanlinem{MeasureTheory.Constructions.Pi}, there is
+\leanlinem{pi_caratheodory}, which shows that $\sigma(\mathcal H)
 \subseteq \mathcal F$ in a special case\todo{which special case?}. In addition, there is
-\leanline{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
+\leanlinem{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
 H$ in the same special case. See also \cite[Theorem
@@
 \noindent
-Here, \leanline{MeasureTheory.OuterMeasure.caratheodory}\todo{used in dot notation... but we don't want to explain this} is already
+Here, \leanlinem{MeasureTheory.OuterMeasure.caratheodory}\todo{used in dot notation... but we don't want to explain this} is already
 implemented and gives the assertion that $\mathcal F$ from
@@
 In our formalization, we are less general and use a direct argument
-for a \leanline{kol_content}. More precisely, we show that a
+for a \leanline{kolContent}. More precisely, we show that a
 \leanline{kolContent}, which is based on a projective family which is
@@
 The next result is already implemented in \leanline{mathlib}: Use
-\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
+\leanlinem{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
 that every open set is inner regular with respect to closed sets, and
-\leanline{Measure.InnerRegular.measurableSet_of_open} to show that
+\leanlinem{Measure.InnerRegular.measurableSet_of_open} to show that
 every measurable set is inner regular with respect to closed sets.
 The claimed outer reularity is part of
-\leanline{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.
+\leanlinem{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.
 
@@
 In the next lemma, we will need that a closed subset of a compact set
-is compact. (See \leanline{isCompact_of_isClosed_subset}.) We describe
+is compact. (See \leanlinem{isCompact_of_isClosed_subset}.) We describe
 the formalization of the next lemma only in conjunction with
@@
 space is compact iff it is complete and totally bounded (see
-\leanline{isCompact_iff_totallyBounded_isComplete} in {\tt
+\leanlinem{isCompact_iff_totallyBounded_isComplete} in {\tt
   mathlib}). Since closed subsets of a complete space are complete,
@@
 product measures (on finite products). More precisely,
-e.g.\ \leanline{StieltjesFunction.outer_Ioc} and
-\leanline{MeasureTheory.constructions.pi.pi_pi_aux} are specific
+e.g.\ \leanlinem{StieltjesFunction.outer_Ioc} and
+\leanlinem{MeasureTheory.constructions.pi.pi_pi_aux} are specific
 examples of the more general statement, that the outer measure extends

### 2023-09-12 Manuscript/proof_outline.tex
semiring -> semi-ring

@@
 \begin{lemma}\label{l1}
-  Let $\mathcal H$ be a semiring, $\mathcal I \subseteq_f \mathcal H$,
+  Let $\mathcal H$ be a semi-ring, $\mathcal I \subseteq_f \mathcal H$,
   $A \in \mathcal H$. Then, there is $\mathcal K \subseteq_f \mathcal
@@
   We proceed by induction on $\mathcal I$. If $\mathcal I$ is a
-  singleton, the assertion is true by the definition of a semiring. If
+  singleton, the assertion is true by the definition of a semi-ring. If
   it holds for some $\mathcal I$ (i.e.\ there is $\mathcal K
@@
 
-For the concrete application we have in mind, we introduce a definition which uses a specific semiring, the measurable cylinders, based on a projective family of measures.
+For the concrete application we have in mind, we introduce a definition which uses a specific semi-ring, the measurable cylinders, based on a projective family of measures.
 That is the goal of the two definitions \leanline{kolmogorovFun} and \leanline{kolContent} below:
@@
 already, the classical extension theorem (extending a set function $m$
-on a semiring $\mathcal H$)) was not provided yet. In particular, this
+on a semi-ring $\mathcal H$)) was not provided yet. In particular, this
 result states that the outer measure coincides with $m$ on $\mathcal
@@
 %terms of {\tt mathlib}.
-The next result extends a set function on a semi-ring\todo{check that we are consistend with semiring vs semi-ring} to an outer measure.
+The next result extends a set function on a semi-ring to an outer measure.
 
@@
 \begin{theorem}[Carathéodory extension]\label{T:masseind}
-  Let $\mathcal H$ be a semiring and $m: \mathcal H\to\mathbb R_+$
+  Let $\mathcal H$ be a semi-ring and $m: \mathcal H\to\mathbb R_+$
   $\sigma$-finite and $\sigma$-additive. Furthermore, let $\mu$ be the
@@
   1.$\Leftrightarrow$2.\ was already shown in
-  Lemma~\ref{l:halbextRing}, since $\mathcal R$ is a semiring.
+  Lemma~\ref{l:halbextRing}, since $\mathcal R$ is a semi-ring.
 

### 2023-09-13 Manuscript/proof_outline.tex
updated manuscript

--- 2023-09-12 Manuscript/proof_outline.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage{ulem}
 %\usepackage{amsmath}
@@
 \definecolor{mygray}{rgb}{0.97,0.97,0.97}
-\newcommand{\leanline}{\mintinline[breaklines, breakafter=., breaklines]{Lean}}
+\definecolor{LightCyan}{rgb}{0.8784,1,1}
+\newcommand{\leanline}[1]{\mintinline[breaklines, breakafter=., breaklines]{Lean}{#1}}%\textsuperscript{\textcolor{gray}{\tiny(m)}}}
+\newcommand{\leanlinecolor}{\mintinline[breaklines, breakafter=., breaklines, bgcolor=LightCyan]{Lean}}
 %\usepackage{refcheck}
@@
 
-\todo[inline]{Clearly differentiate refs to mathlib vs refs to new lemmas?}
+%\todo[inline]{Clearly differentiate refs to mathlib vs refs to new lemmas?}
 
@@
 \todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}
+
+%\todo[inline]{explain leanlinem}
 
@@
 So, our work complements \cite{ying2023formalization} in the sense that probability spaces $(\Omega, \mathcal F, \mathbf P)$ on which you can define a martingale exist.
-However, we do not give any specific example for the construction of a specific instance of such a space\todo{could we give one?}.
+{\color{blue}As an application of our implementation, we construct a probability measure on an infinite product space where all coordinates are independent; see Section~\ref{ss:concl}.} % However, we do not give any specific example for the construction of a specific instance of such a space\todo{could we give one?}.
 
@@
 \begin{enumerate}
-\item a formalization of (the set-system of) semi-rings (see Definition~\ref{def:semi} below); 
-\item a definition of additive contents over semi-rings\todo{R: I added that point but we don't talk much about contents};
+\item a formalization of (the set-system of) semi-rings (see
+  Definition~\ref{def:semi});
+\item {\color{blue}a definition of additive contents over semi-rings (see Definition~\ref{def:content});}
 %\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset\todo{for a measure, this is already in mathlib?} (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
@@
 The Kolmogorov extension theorem has been previously formalized in Isabelle/HOL \cite{Immler2012}.
-This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical\todo{Lean has dependent types, Isabelle doesn't}.
-However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations\todo{no, the type theory is not the same} in these two interactive theorem provers.
+This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical.
+In addition, {\color{blue}Isabelle works with simple type theory, while Lean is an implementation of dependent theory.} %However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations\todo{no, the type theory is not the same} in these two interactive theorem provers.
 
@@
 \todo{reorder that next remark? I'll make a proposal}
-\begin{remark}[Metric and topological spaces in {\tt mathlib}] \mbox{}
+\begin{remark}[Metric and topological spaces in {\tt mathlib}] \label{rem:uniform}\mbox{}
   \begin{enumerate}
-    \item \sloppy In \leanline{Lean}, we have the \leanline{class PseudoEMetricSpace (α : Type u) extends EDist}, where
-      \leanline{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}
-      (the notation \leanline{ℝ≥0∞} stands for $[0, \infty]$), and the extension to
+    \item \sloppy In \leanline{Lean}, we have the \leanline{class
+      PseudoEMetricSpace (α : Type u) extends EDist}, where
+      \leanline{EDist} comes with an extended distance
+      \leanline{edist: α → α → ℝ≥0∞} (the notation \leanline{ℝ≥0∞}
+      stands for $[0, \infty]$), and the extension to
       \leanline{PseudoEMetricSpace} consists of the components,
@@
       \cite{melikhov2011metrizable}.
-    \item \sloppy For topological spaces, we have \leanline{class TopologicalSpace(α : Type u) : Type u}, which comes with
-      \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
+    \item \sloppy For topological spaces, we have \leanline{class
+      TopologicalSpace(α : Type u) : Type u}, which comes with
+      \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen
+        Set.univ}, \leanline{isOpen_inter} and
       \leanline{isOpen_sUnion}, which are exactly the properties of a
@@
 
-We now come to our contribution%, which we will PR to mathlib
-: the extension theorem is on finite measures on product spaces\todo{I don't understand this sentence}. In
-particular, the product is taken over a set of arbitrary
-cardinality. The next definition covers the important concept of a
-projective family of measures. In short, we define measures on
-any finite subset of indices in a consistent way. Finite
-sets (of some type \leanline{ι}) are formalized with \leanline{Finset ι}. % Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
+We now come to our contribution. 
+%which we will PR to mathlib :
+{\color{blue}The extension theorem is a statement about extending a
+  set-function on a product space to a (finite) measure, where the
+  product space can come with an arbitraty index set.}  The next
+definition covers the important concept of a projective family of
+measures. In short, we define measures on any finite subset of indices
+in a consistent way. Finite sets (of some type \leanline{ι}) are
+formalized with \leanline{Finset ι}. % Such a \leanline{Finset} is
+defined as a \leanline{Multiset} (a
 %quotient type of the \leanline{List}-type with respect to permutations
@@
 
-In the formalization, we use \leanline{variable {ι : Type _} {α : ι →
-    Type _} }, which fixes the index set $\iota$ and all spaces
+{\color{blue}In contrast to statements already part of
+  \leanline{mathlib}, in the sequel we \leanlinecolor{highlight}
+  definitions and statements which are part of our own code.} In the
+formalization of the above, we use \leanline{variable {ι : Type _} {α
+    : ι → Type _} }, which fixes the index set $\iota$ and all spaces
 $\alpha_t, t\in\iota$ as global variables. In addition, note that the
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def IsProjective [Preorder ι]
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  def IsProjective
+    [Preorder ι]
     (P : ∀ j : ι, α j) (π : ∀ {i j : ι}, j ≤ i → α i → α j) : Prop :=
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   def IsProjectiveMeasureFamily
@@
 It is worth understanding the precise connection of
-\leanline{isProjective} and \leanline{isProjectiveMeasureFamily}. In
-the latter, the first variable of \leanline{isProjective} is the
-family \leanline{P} of finite measures for all finite subsets of
-\leanline{ι}. The second variable is the functions which maps two sets
-\leanline{I J : Finset ι} and a proof \leanline{hJI} of \leanline{J ⊆ I}
-together with \leanline{P I} (which is \leanline{μ} in the
-statement) to the image measure on \leanline{J}, which is
-\leanline{μ.map (fun x : ∀ i : I, α i => fun j => x ⟨j, hJI j.2⟩)}.
-The map defined by the \leanline{=>}-notation maps every
-$(x_i)_{i\in I}$ to a function of \leanline{j}, whose type is the
-subtype of \leanline{I}, consisting of a value and a proof of
-\leanline{J ⊆ I}. In other words, this is $(x_j)_{j\in J}$.
+\leanlinecolor{isProjective} and
+\leanlinecolor{isProjectiveMeasureFamily}. In the latter, the first
+variable of \leanlinecolor{isProjective} is the family \leanline{P} of
+finite measures for all finite subsets of \leanline{ι}. The second
+variable is the functions which maps two sets \leanline{I J : Finset
+  ι} and a proof \leanline{hJI} of \leanline{J ⊆ I} together with
+\leanline{P I} (which is \leanline{μ} in the statement) to the image
+measure on \leanline{J}, which is \leanline{μ.map (fun x : ∀ i : I, α
+  i => fun j => x ⟨j, hJI j.2⟩)}.  The map defined by the
+\leanline{=>}-notation maps every $(x_i)_{i\in I}$ to a function of
+\leanline{j}, whose type is the subtype of \leanline{I}, consisting of
+a value and a proof of \leanline{J ⊆ I}. In other words, this is
+$(x_j)_{j\in J}$.
 
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 def projectiveLimitWithWeakestHypotheses
@@
 Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
-open balls around $x$ and $y$ overlap\todo{it would be nice to explain how the formalization led us to remove that hypothesis}.
+open balls around $x$ and $y$ overlap. {\color{blue}We note that this
+  generalization was only possible since underlying results in
+  \leanline{mathlib} were already provided on the same level of
+  generality: More precisely,
+  \leanline{isCompact_iff_totallyBounded_isComplete}, which shows that
+  a set $A \subseteq \alpha$ is compact iff it is complete and totally
+  bounded {\color{blue}(see note~\ref{note:tot})}, requires $\alpha$
+  to be a uniform space (recall that every metric space is uniform);
+  recall such spaces from Remark~\ref{rem:uniform}, and see
+  note~\ref{note:tot} for a definition of total boundedness. Since we
+  also require the underlying space(s) to be second countable (see
+  also in the proof of Lemma~\ref{L:relcoPol}), we have to make a
+  countabiliy assumption, which leads us to (extended) pseudo-metric
+  spaces by \leanline{UniformSpace.pseudoMetrizableSpace}; see also
+  Remark~\ref{rem:uniform} for some more details.}
+
+% \todo{it would be nice to explain how the formalization led us to
+% remove that hypothesis}.
 
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   structure SetSemiring (C : Set (Set α)) : Prop where
@@
 %\mathcal H$.
-Since we are dealing with product spaces, we will use cylinders\todo{R: rewrite this}, which
+The import ring of sets in our formalization is the ring of measurable cylinders on a product space, which
 are defined as follows:
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def cylinder (s : Finset ι) (S : Set (∀ i : s, α i)) :
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  def cylinder
+    (s : Finset ι) (S : Set (∀ i : s, α i)) :
     Set ((i : ι) → α i) :=
@@
 
+\noindent
 In this definition, \leanline{s} is a finite subset of the index set
@@
 \alpha_i \to \prod_{i \in s} \alpha_i$, and consider the preimage of
-$S$. (This is what the last line in previous definition gives.) From
-this, we can define the set of all measurable cylinders.
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def cylinders : Set (Set ((i : ι) → α i)) :=
-    ⋃ (s) (S) (_ : MeasurableSet S), {cylinder s S}
+$S$. (This is what the last line in previous definition gives.)  {\color{blue}Note,
+however, that \leanline{s} (comprising the potential coordinates where
+the cylinder deviates from the whole set) need not be unique. However,
+we have
+
+\begin{minted}[highlightlines={1}, framesep = 0pt, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+theorem mem_cylinders
+  (t : Set ((i : ι) → α i)) :
+  t ∈ cylinders α ↔
+  ∃ (s S : _) (_ : MeasurableSet S), t = cylinder s S
+\end{minted}
+By the exists-statement, from this result, we are able to choose a
+\leanline{Finset ι}, which can be used. This leads to the definition
+of \leanlinecolor{cylinders.finset}, which will be needed below.}
+
+From the definition of a \leanlinecolor{cylinder}, we can define the
+set of all measurable cylinders.
+
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  def cylinders : Set (Set ((i : ι) → α i)) := ⋃ (s) (S) (_ :
+  MeasurableSet S), {cylinder s S}
 \end{minted}
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1-3}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   theorem setField_cylinders : SetField (cylinders α) 
@@
 From the field/ring/semi-ring of cylinders, we have to define the
-generated $\sigma$-alebra. This uses the following\todo{one of those 2 is ours, the other is from mathlib. How do we show the difference?}:
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+generated $\sigma$-algebra. This uses the following:
+%\todo{one of those 2 is ours, the other is from mathlib. How do we show the  difference?}:
+\begin{minted}[highlightlines={1}, framesep = 0pt, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   theorem generateFrom_cylinders :
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   theorem exists_disjoint_finset_diff_eq
@@
 an element from an $\exists$-statement. In addition, we do not allow
-$\emptyset \in \mathcal K$ without loss of generality\todo{that exclusion was useful for something. Find out what that is}:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+$\emptyset \in \mathcal K$ without loss of generality\todo{that
+  exclusion was useful for something. Find out what that is}:
+
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 def diff₀ (hC : SetSemiring C)
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 def indexedDiff₀ (hC : SetSemiring C)
@@
   n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
-is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)}
-is the set consisting of the
+is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin
+  (card J)) : Finset (Set α)} is the set consisting of the
 \leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
-\leanline{finsetLT_subset'} gives a proof that \leanline{↑(finsetLT J n) ⊆ C}.
-So, in terms of Lemma~\ref{l2}, this gives, for
-$n=1,...,m$, a construction for $\mathcal K_n \subseteq_f \mathcal H$.
+\leanlinecolor{finsetLT_subset'} gives a proof that
+\leanline{↑(finsetLT J n) ⊆ C}.  So, in terms of Lemma~\ref{l2}, this
+gives, for $n=1,...,m$, a construction for $\mathcal K_n \subseteq_f
+\mathcal H$.
 
@@
 
-\begin{definition}
-  For some set $\alpha$, let $\mathcal H \subseteq 2^\alpha$ and $m :
-  \mathcal H \to [0,\infty]$.
+\begin{definition}\label{def:content}
+  For some set $\alpha$, let $\mathcal H \subseteq 2^\alpha$
+  {\color{blue}and call any $m : \mathcal H \to [0,\infty]$ a
+    content.}
   \begin{enumerate}
@@
   \item If $\mathcal H$ is a $\sigma$-algebra and $m$ is
-    $\sigma$-additive, we call $m$ a measure\todo{need 0 at the empty set? Otherwise could be infinite everywhere}.
+    $\sigma$-additive {\color{blue} with $m(\emptyset) = 0$}, we call
+    $m$ a measure.
   \item If $\mathcal H = 2^\alpha$ and $m$ is monotone and
@@
 \noindent
-For additive contents\todo{"content" not defined?}, we need two definitions. \leanline{AddContent}
-defines a set-function \leanline{tuFun} on \leanline{Set α}, whereas
-\leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+For additive contents, we need two
+definitions. \leanlinecolor{AddContent} defines a set-function
+\leanline{toFun} on \leanline{Set α}, whereas
+\leanlinecolor{extendContent} comes with \leanline{m : ∀ s : Set α, s
+  ∈ C → ℝ≥0∞}.
+
+\begin{minted}[highlightlines={1, 10}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 structure AddContent (C : Set (Set α)) where
@@
 \noindent
-In these formalizations, we introduced $m(\emptyset)=0$ as a separate
-hypothesis with \leanline{empty'}. Note that this also follows
+\xout{In these formalizations, we introduced $m(\emptyset)=0$ as a
+  separate hypothesis with empty'. Note that this also follows
   formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
-m(\emptyset)$\todo{not if m is infinite everywhere}. However, since \leanline{Finset}s are assumed to have
-distinct elements, we cannot have \leanline{∅} twice in \leanline{I}.
-
-For the concrete application we have in mind, we introduce a definition which uses a specific semiring, the measurable cylinders, based on a projective family of measures.
-That is the goal of the two definitions \leanline{kolmogorovFun} and \leanline{kolContent} below:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  m(\emptyset)$. However, since Finsets are assumed to have distinct
+  elements, we cannot have ∅ twice in I.}
+
+For the concrete application we have in mind, we introduce a
+definition which uses a specific semi-ring, the measurable cylinders,
+based on a projective family of measures.  That is the goal of the two
+definitions \leanlinecolor{kolmogorovFun} and
+\leanlinecolor{kolContent} below:
+
+\begin{minted}[highlightlines={1,7}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 def kolmogorovFun
@@
 
-\todo{cylinders.finset not defined?}
-
-Here, \leanline{kolmogorovFun_empty} and
-\leanline{kolmogorovFun_additive} give proofs for \leanline{m_empty}
-and \leanline{m_add} in \leanline{extendContent} as applied to a
-\leanline{kolmogorovFun}.
+Here, \leanlinecolor{kolmogorovFun_empty} and
+\leanlinecolor{kolmogorovFun_additive} give proofs for
+\leanline{m_empty} and \leanline{m_add} in
+\leanlinecolor{extendContent} as applied to a
+\leanlinecolor{kolmogorovFun}.
 
@@
 well as $\sigma$-additive $\Rightarrow$ $\sigma$-sub-additive in the
-proof of the Carathéodory extension, Theorem~\ref{T:masseind}\todo{I don't understand this sentence}. The
-formalization of the next result comes in several lemmas, which we do
-not show here.
+proof of the Carathéodory extension,
+Theorem~\ref{T:masseind}. {\color{blue}We do not show any details
+  about its formalization here.}
 
@@
 Carathéodory. The first leads to the definition of the measurable
-space \leanline{OuterMeasure.caratheodory}, covered in {\tt
-  mathlib}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}
+space \leanline{OuterMeasure.caratheodory}, covered in
+\leanline{mathlib}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}
 
@@
 \leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
-\subseteq \mathcal F$ in a special case\todo{which special case?}. In addition, there is
+\subseteq \mathcal F$ in {\color{blue} the construction of a product
+  measure for a finite index set}. In addition, there is
 \leanline{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-def ofAddSubaddCaratheodory
-    (hC : SetSemiring C)
-    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
-    (m_empty : m ∅ hC.empty_mem = 0)
-    (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
-      (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
-      (h_mem : ⋃₀ ↑I ∈ C),
-      m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
-    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
-      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
-      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
-    @Measure α (inducedOuterMeasure m
-      hC.empty_mem m_empty).caratheodory
-\end{minted}
-\todo{is this the right lemma? Do we have a similar one about contents?}
-
-\noindent
-Here, \leanline{MeasureTheory.OuterMeasure.caratheodory}\todo{used in dot notation... but we don't want to explain this} is already
-implemented and gives the assertion that $\mathcal F$ from
-Theorem~\ref{T:cara} is a $\sigma$-algebra. Note that
-\leanline{@Measure} is the same as \leanline{Measure}, but all
-variables have to be given explicitely. The second argument is a proof
-that measurable sets form a $\sigma$-algebra which cannot be inferred
-by the type class system of \leanline{Lean}.
-
-\noindent
-In the proof of Kolmogorov's extension theorem, we use a closely
-related definition\todo{should we present only this one and omit the lemma above?}:
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+%% \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+%% def ofAddSubaddCaratheodory
+%%     (hC : SetSemiring C)
+%%     (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
+%%     (m_empty : m ∅ hC.empty_mem = 0)
+%%     (m_add : ∀ (I : Finset (Set α)) (h_ss : ↑I ⊆ C)
+%%       (_h_dis : PairwiseDisjoint (I : Set (Set α)) id)
+%%       (h_mem : ⋃₀ ↑I ∈ C),
+%%       m (⋃₀ I) h_mem = ∑ u : I, m u (h_ss u.prop))
+%%     (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
+%%       (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
+%%       m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
+%%     @Measure α (inducedOuterMeasure m
+%%       hC.empty_mem m_empty).caratheodory
+%% \end{minted}
+
+%% \noindent
+%% Here, \leanline{MeasureTheory.OuterMeasure.caratheodory}\todo{used in
+%%   dot notation... but we don't want to explain this} is already
+%% implemented and gives the assertion that $\mathcal F$ from
+%% Theorem~\ref{T:cara} is a $\sigma$-algebra. Note that
+%% \leanline{@Measure} is the same as \leanline{Measure}, but all
+%% variables have to be given explicitely. The second argument is a proof
+%% that measurable sets form a $\sigma$-algebra which cannot be inferred
+%% by the type class system of \leanline{Lean}.
+
+%% \noindent
+%% In the proof of Kolmogorov's extension theorem, we use a closely
+%% related definition\todo{should we present only this one and omit the lemma above?}:
+
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 def Measure.ofAddContent
@@
     (m : AddContent C)
-    (m_sigma_subadd : ...) :
+    (m_sigma_subadd : ∀ ⦃f : ℕ → Set α⦄
+      (hf : ∀ i, f i ∈ C) (hf_Union : (⋃ i, f i) ∈ C),
+      m (⋃ i, f i) hf_Union ≤ ∑' i, m (f i) (hf i)) :
     Measure α
 \end{minted}
-with \leanline{m_sigma_subadd} as above.
 
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem countably_additive_of_todo (hC : SetRing C)
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  theorem countably_additive_of_todo
+    (hC : SetRing C)
     (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   def MeasureTheory.Measure.InnerRegular
@@
 \end{definition}
-\todo{no definition for this in the code for now}
+
+Here is the formalization:
+
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  def IsCompactFamily (p : Set α → Prop) : Prop :=
+    ∀ C : ℕ → Set α, (∀ i, p (C i)) → ⋂ i, C i = ∅ →
+    ∃ (s : Finset ℕ), ⋂ i ∈ s, C i = ∅
+\end{minted}
 
@@
 
-\todo[inline]{Work on this, if time permits!}
-
-In our formalization, we are less general and use a direct argument
-for a \leanline{kol_content}. More precisely, we show that a
-\leanline{kolContent}, which is based on a projective family which is
-inner regular with respect to compact and closed sets is
-$\sigma$-additive.
-
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+%% def closedCompactCylinders :
+%%   Set (Set ((i : ι) → α i)) :=
+%%   ⋃ (s) (S) (_ : IsClosed S) (_ : IsCompact S), {cylinder s S}
+
+%%     theorem isCompactFamily_cylinders :
+%% IsCompactFamily (fun t ↦ t ∈ closedCompactCylinders α)
+
+%% lemma AddContent.sigma_additive_of_regular [Nonempty α] (hR : SetRing R) (m : AddContent R)
+%%     (hm_ne_top : ∀ {s} (_ : s ∈ R), m s ≠ ∞)
+%%     (hC : IsCompactFamily C) (hCR : C ⊆ R)
+%%     (h_reg : ∀ A (_ : A ∈ R) (ε : ℝ≥0∞) (_ : 0 < ε), ∃ K ∈ C, K ⊆ A ∧ m (A \ K) ≤ ε)
+%%     ⦃f : ℕ → Set α⦄ (hf : ∀ i, f i ∈ R) (hUf : (⋃ i, f i) ∈ R) (h_disj : Pairwise (Disjoint on f)) :
+%%     m (⋃ i, f i) = ∑' i, m (f i)
+
+%In our formalization, we are less general and use a direct argument
+%for a \leanlinecolor{kolContent}. More precisely, we show that a
+%\leanlinecolor{kolContent}, which is based on a projective family
+%which is inner regular with respect to compact and closed sets is
+%$\sigma$-additive.
+
+In order to apply this result and show the extension theorem, we need
+to show that
+%$\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact}\}$ as
+%  well as
+$\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact and
+  closed}\}$ is a compact systems. Note that compact sets are closed
+in Hausdorff spaces, but we do not have this property since we are
+working with pseudo-metric spaces. Since Lemma~\ref{l:stetigcompact}
+gives the $\sigma$-additivity of a \leanlinecolor{kolContent}, which
+is defined through the projective family \leanline{P}, we have:
+
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 theorem kolContent_sigma_additive_of_innerRegular
@@
   kolContent hP (⋃ i, f i) = ∑' i, kolContent hP (f i)
-
-  theorem kolContent_countably_subadditive_of_innerRegular
-    (hP : IsProjectiveMeasureFamily P)
-    (hP_inner : ∀ J,
-      (P J).InnerRegular (fun s => IsCompact s ∧ IsClosed s)
-      MeasurableSet)
-    ⦃f : ℕ → Set (∀ i, α i)⦄
-    (hf : ∀ i, f i ∈ cylinders α)
-    (hf_Union : (⋃ i, f i) ∈ cylinders α) :
-    kolContent hP (⋃ i, f i) ≤ ∑' i, kolContent hP (f i)
-\end{minted}
-
-For using the more general version, we would need to show that
-$\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact}\}$ as
-well as $\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact
-  and closed}\}$ are compact systems. Note that compact sets are
-closed in Hausdorff spaces, but we do not have this property since we
-are working with pseudo-metric spaces. In addition, the assumption
-\leanline{hP_inner} from above translates directly to inner regularity
-of the \leanline{kolContent}, which is defined through the projective
-family \leanline{P}. Then, Lemma~\ref{l:stetigcompact} gives the same
-result as the formalization. Moreover, since the Carathéodory
-extension theorem requires $\sigma$-subadditivity (rather than
-$\sigma$-additivity), we can use
-\leanline{countably_subadditive_of_countably_additive} in order to
-show \leanline{kolContent_countably_subadditive_of_innerRegular},
-which shows that the assumptions of the last result also imply
-$\sigma$-subadditivity.
+\end{minted}
+
+Note that the assumption \leanline{hP_inner} from above translates
+directly to inner regularity of the \leanlinecolor{kolContent}, which
+is defined through the projective family \leanline{P}. Moreover, since
+the Carathéodory extension theorem requires $\sigma$-subadditivity
+(rather than $\sigma$-additivity), we can use
+\leanlinecolor{countably_subadditive_of_countably_additive} in order
+to show
+\leanlinecolor{kolContent_countably_subadditive_of_innerRegular}, and
+the assumptions of the last result also imply $\sigma$-subadditivity.
+
+%  theorem kolContent_countably_subadditive_of_innerRegular
+%    (hP : IsProjectiveMeasureFamily P)
+%    (hP_inner : ∀ J,
+%      (P J).InnerRegular (fun s => IsCompact s ∧ IsClosed s)
+%      MeasurableSet)
+%    ⦃f : ℕ → Set (∀ i, α i)⦄
+%    (hf : ∀ i, f i ∈ cylinders α)
+%    (hf_Union : (⋃ i, f i) ∈ cylinders α) :
+%    kolContent hP (⋃ i, f i) ≤ ∑' i, kolContent hP (f i)
 
@@
 every measurable set is inner regular with respect to closed sets.
-The claimed outer reularity is part of
+The claimed outer regularity is part of
 \leanline{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.
@@
 to extended pseudo-metric spaces, we use that any subset of such a
-space is compact iff it is complete and totally bounded (see
+space is compact iff it is complete and totally
+bounded\footnote{\label{note:tot}A subset of a pseudo-metric space is
+  totally bounded iff, for all $\varepsilon>0$, it can be covered by
+  finitely many balls of radius $\epsilon$.} (see
 \leanline{isCompact_iff_totallyBounded_isComplete} in {\tt
@@
   \]
-  which by construction is totally bounded\todo{totally bounded not defined?} (for $r>0$ take $n$ such
+  which by construction is totally bounded (for $r>0$ take $n$ such
   that $\varepsilon_n < r$, such that $B_r(x_1),...,B_r(x_{N_n})$
-  cover $A$), and therefore has a compact closure as noted before the
-  lemma\todo{is it noted before the lemma? where?}. In addition, by $\sigma$-sub-additivity of $\mu$,
+  cover $A$), and therefore has a compact closure as noted
+  {\color{blue} in the lines directly preceding the lemma}. In
+  addition, by $\sigma$-sub-additivity of $\mu$,
   \begin{align*}
@@
 
-\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
   theorem innerRegular_isCompact_isClosed_
@@
 \end{minted}
-Let us look closer at the formalized proof:
-\leanline{kolContent_countably_subadditive_of_innerRegular} has the
-same hypotheses as
-\leanline{kolContent_sigma_additive_of_innerRegular}, except
+\sloppy Let us look closer at the formalized proof:
+\leanlinecolor{kolContent_countably_subadditive_of_innerRegular} has
+the same hypotheses as
+\leanlinecolor{kolContent_sigma_additive_of_innerRegular}, except
 \leanline{h_disj}. So, the lemma in the last brackets shows
 $\sigma$-sub-additivity of the content $P$ from
-Theorem~\ref{T1}. Then, \leanline{generateFrom_cylinders} gives the \leanline{MeasurableSpace} on which we define the
-\leanline{addContent}.
+Theorem~\ref{T1}. Then, \leanlinecolor{generateFrom_cylinders} gives
+the \leanline{MeasurableSpace} on which we define the
+\leanlinecolor{addContent}.
 
 \section{Concluding thoughts}
-We will highlight some paths we imagine can be taken, leading to
-further extensions of \leanline{mathlib}, which build on our
-contribution.
+\label{ss:concl}
+{\color{blue} The usability of our contribution must be proved by an
+  example: Adding to our general theory, we have implemented the
+  example of an infinite product measure. (As an application of
+  \leanlinecolor{projectiveLimitWithWeakestHypotheses}, we have to use
+  complete and separable extended pseudo-metric spaces, but note that
+  an implementation of the Theorem by Ionescu and Tulcea would give
+  the same result without these prerequisites.) Here, we can build on
+  \leanline{Measure.pi}, which defines a finite product measure.
+  Using the same automatically inferred instances, we obtain:}
+
+\begin{minted}[highlightlines={1,8}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  def independentFamily
+    (P : ∀ i, Measure (α i))
+    [∀ i, IsProbabilityMeasure (P i)] : Measure (∀ i, α i)
+    := projectiveLimitWithWeakestHypotheses
+    (fun J : Finset ι => Measure.subset_pi P J)
+    (product_isProjective P)
+
+  def Measure.subset_pi
+    (P : ∀ i, Measure (α i)) (I : Finset ι) :
+    Measure ((i : I) → (α i)) :=
+    Measure.pi (fun (i : I) => (P (i : ι)))
+\end{minted}
+
+Moreover, we will highlight some paths we imagine can be taken,
+leading to further extensions of \leanline{mathlib}, which build on
+our contribution.
 

### 2023-09-13 Manuscript/proof_outline.tex
another small change

@@
 
-\todo[inline]{Rename the todo! And replace by the version about contents?}
+\todo[inline]{Rename the todo! \xout{And replace by the version about contents?}}
 
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem countably_additive_of_todo
+  theorem countably_additive_addContent_of_todo
   (hC : SetRing C)
-    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
-    (hm_ne_top : ∀ {s} (hs : s ∈ C), m s hs ≠ ∞)
-    (hm_add : ∀ {s t : Set α} (hs : s ∈ C) (ht : t ∈ C),
-      Disjoint s t →
-      m (s ∪ t) (hC.union_mem hs ht) = m s hs + m t ht)
-    (hm : ∀ ⦃s : ℕ → Set α⦄ (hs : ∀ n, s n ∈ C),
+  (m : AddContent C)
+  (hm_ne_top : ∀ {s} (_ : s ∈ C), m s ≠ ∞)
+  (hm : ∀ ⦃s : ℕ → Set α⦄ (_ : ∀ n, s n ∈ C),
     Antitone s → (⋂ n, s n) = ∅ →
-      Tendsto (fun n => m (s n) (hs n)) atTop (nhds 0))
-      ⦃f : ℕ → Set α⦄ (h : ∀ i, f i ∈ C) (hUf : (⋃ i, f i) ∈ C)
+    Tendsto (fun n => m (s n)) atTop (nhds 0))
+    ⦃f : ℕ → Set α⦄ (hf : ∀ i, f i ∈ C) (hUf : (⋃ i, f i) ∈ C)
     (h_disj : Pairwise (Disjoint on f)) :
-    m (⋃ i, f i) hUf = ∑' i, m (f i) (h i)
+  m (⋃ i, f i) = ∑' i, m (f i) 
 \end{minted}

### 2023-09-14 Manuscript/proof_outline.tex
rename todo in lemma name

@@
 
-%\todo[inline]{Clearly differentiate refs to mathlib vs refs to new lemmas?}
-
 \todo[inline]{Mention a precise commit of Mathlib}
@@
 \todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}
-
-%\todo[inline]{explain leanlinem}
 
@@
 So, our work complements \cite{ying2023formalization} in the sense that probability spaces $(\Omega, \mathcal F, \mathbf P)$ on which you can define a martingale exist.
-{\color{blue}As an application of our implementation, we construct a probability measure on an infinite product space where all coordinates are independent; see Section~\ref{ss:concl}.} % However, we do not give any specific example for the construction of a specific instance of such a space\todo{could we give one?}.
+{\color{blue}As an application of our implementation, we construct a probability measure on an infinite product space where all coordinates are independent; see Section~\ref{ss:concl}.}
 
@@
 \item {\color{blue}a definition of additive contents over semi-rings (see Definition~\ref{def:content});}
-%\item adding some main lemmas to measure theory, e.g.\ continuity of a measure at the emptyset\todo{for a measure, this is already in mathlib?} (see e.g.\ Lemma~\ref{l:halbextRing}, Proposition~\ref{P:stetigmass}, Lemma~\ref{l:stetigcompact} below);
 \item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below);
@@
 This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical.
-In addition, {\color{blue}Isabelle works with simple type theory, while Lean is an implementation of dependent theory.} %However, type theory as the basis of the formalization, as well as several mathematical concepts (e.g.\ projective families of measures or the continuity of a measure in the emptyset) are the same in both formalisations\todo{no, the type theory is not the same} in these two interactive theorem provers.
+In addition, {\color{blue}Isabelle works with simple type theory, while Lean is an implementation of dependent theory.}
 
@@
 
-% \todo{it would be nice to explain how the formalization led us to
-% remove that hypothesis}.
-
 \subsection{Extending a set function}
@@
 generated $\sigma$-algebra. This uses the following:
-%\todo{one of those 2 is ours, the other is from mathlib. How do we show the  difference?}:
 \begin{minted}[highlightlines={1}, framesep = 0pt, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
 an element from an $\exists$-statement. In addition, we do not allow
-$\emptyset \in \mathcal K$ without loss of generality\todo{that
-  exclusion was useful for something. Find out what that is}:
+$\emptyset \in \mathcal K$ without loss of generality:
 
@@
 
-\todo[inline]{Rename the todo! \xout{And replace by the version about contents?}}
-
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  theorem countably_additive_addContent_of_todo
+  theorem sigma_additive_addContent_of_tendsto_zero
   (hC : SetRing C)

### 2023-09-14 Manuscript/proof_outline.tex
change countably to sigma in lemma names

@@
 (rather than $\sigma$-additivity), we can use
-\leanlinecolor{countably_subadditive_of_countably_additive} in order
+\leanlinecolor{sigma_subadditive_of_sigma_additive} in order
 to show
-\leanlinecolor{kolContent_countably_subadditive_of_innerRegular}, and
+\leanlinecolor{kolContent_sigma_subadditive_of_innerRegular}, and
 the assumptions of the last result also imply $\sigma$-subadditivity.
 
-%  theorem kolContent_countably_subadditive_of_innerRegular
+%  theorem kolContent_sigma_subadditive_of_innerRegular
 %    (hP : IsProjectiveMeasureFamily P)
@@
     generateFrom_cylinders (kolContent hP)
-    (kolContent_countably_subadditive_of_innerRegular hP
+    (kolContent_sigma_subadditive_of_innerRegular hP
       fun J => innerRegular_isCompact_isClosed_
@@
 \sloppy Let us look closer at the formalized proof:
-\leanlinecolor{kolContent_countably_subadditive_of_innerRegular} has
+\leanlinecolor{kolContent_sigma_subadditive_of_innerRegular} has
 the same hypotheses as

### 2023-09-14 Manuscript/proof_outline.tex
rename compactfamily to compactsystem

@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def IsCompactFamily (p : Set α → Prop) : Prop :=
+  def IsCompactSystem (p : Set α → Prop) : Prop :=
     ∀ C : ℕ → Set α, (∀ i, p (C i)) → ⋂ i, C i = ∅ →
@@
 
-%%     theorem isCompactFamily_cylinders :
-%% IsCompactFamily (fun t ↦ t ∈ closedCompactCylinders α)
+%%     theorem isCompactSystem_cylinders :
+%% IsCompactSystem (fun t ↦ t ∈ closedCompactCylinders α)
 
@@
 %%     (hm_ne_top : ∀ {s} (_ : s ∈ R), m s ≠ ∞)
-%%     (hC : IsCompactFamily C) (hCR : C ⊆ R)
+%%     (hC : IsCompactSystem C) (hCR : C ⊆ R)
 %%     (h_reg : ∀ A (_ : A ∈ R) (ε : ℝ≥0∞) (_ : 0 < ε), ∃ K ∈ C, K ⊆ A ∧ m (A \ K) ≤ ε)

### 2023-09-14 Manuscript/proof_outline.tex
remove some blue/todos

@@
 Extending Lemma~\ref{l1}, we would like to write a finite union of
-members of a semi-ring as a finite union of disjoint sets.
-
-\todo[inline]{Is this section going into too much detail?}
+members of a semi-ring as a finite union of disjoint sets\todo{Is this section going into too much detail?}.
 
@@
   For some set $\alpha$, let $\mathcal H \subseteq 2^\alpha$
-  {\color{blue}and call any $m : \mathcal H \to [0,\infty]$ a
-    content.}
+  and call any $m : \mathcal H \to [0,\infty]$ a
+    content.
   \begin{enumerate}
@@
   \item If $\mathcal H$ is a $\sigma$-algebra and $m$ is
-    $\sigma$-additive {\color{blue} with $m(\emptyset) = 0$}, we call
+    $\sigma$-additive with $m(\emptyset) = 0$, we call
     $m$ a measure.
@@
 
-\noindent
-\xout{In these formalizations, we introduced $m(\emptyset)=0$ as a
-  separate hypothesis with empty'. Note that this also follows
-  formally from $m(\emptyset) = m(\emptyset \cup \emptyset) = 2
-  m(\emptyset)$. However, since Finsets are assumed to have distinct
-  elements, we cannot have ∅ twice in I.}
-
 For the concrete application we have in mind, we introduce a
@@
 proof of the Carathéodory extension,
-Theorem~\ref{T:masseind}. {\color{blue}We do not show any details
-  about its formalization here.}
+Theorem~\ref{T:masseind}. We do not show any details
+  about its formalization here.
 

### 2023-09-14 Manuscript/proof_outline.tex
reorder explanation of cylinders

@@
 We do not show the formalization of rings here.
-%In our formalization, we also use the notion of a field $\mathcal H$,
-%which is a ring with $\alpha \in
-%\mathcal H$.
-The import ring of sets in our formalization is the ring of measurable cylinders on a product space, which
-are defined as follows:
+The important ring of sets in our formalization is the ring of measurable cylinders on a product space.
+A cylinder is defined as follows:
 
@@
 \alpha_i \to \prod_{i \in s} \alpha_i$, and consider the preimage of
-$S$. (This is what the last line in previous definition gives.)  {\color{blue}Note,
-however, that \leanline{s} (comprising the potential coordinates where
-the cylinder deviates from the whole set) need not be unique. However,
-we have
-
+$S$ (This is what the last line in previous definition gives.).
+We can define \leanlinecolor{cylinders α}, the set of all measurable cylinders.
+
+\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+  def cylinders : Set (Set ((i : ι) → α i)) := ⋃ (s) (S) (_ :
+  MeasurableSet S), {cylinder s S}
+\end{minted}
+
+Membership of that set is characterized as follows:
 \begin{minted}[highlightlines={1}, framesep = 0pt, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
@@
 \end{minted}
-By the exists-statement, from this result, we are able to choose a
-\leanline{Finset ι}, which can be used. This leads to the definition
-of \leanlinecolor{cylinders.finset}, which will be needed below.}
-
-From the definition of a \leanlinecolor{cylinder}, we can define the
-set of all measurable cylinders.
-
-\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def cylinders : Set (Set ((i : ι) → α i)) := ⋃ (s) (S) (_ :
-  MeasurableSet S), {cylinder s S}
-\end{minted}
-
-The set-system of cylinders is in fact a field, hence a ring and hence
+\noindent
+The finite set \leanline{s} of the potential coordinates where
+the cylinder deviates from the whole set needs not be unique.
+From the exists-statement of \leanline{mem_cylinders}, we however choose an
+arbitrary such \leanline{Finset ι}, called \leanlinecolor{cylinders.finset}.
+
+The set-system of cylinders is in fact a field, hence a ring and
 a semi-ring. This means we can prove the following:

### 2023-09-14 Manuscript/proof_outline.tex
review spacing of code in the paper
remove some blue

@@
   Definition~\ref{def:semi});
-\item {\color{blue}a definition of additive contents over semi-rings (see Definition~\ref{def:content});}
+\item a definition of additive contents over semi-rings (see Definition~\ref{def:content});
 \item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below);
@@
 We now come to our contribution. 
-%which we will PR to mathlib :
-{\color{blue}The extension theorem is a statement about extending a
+The extension theorem is a statement about extending a
 set-function on a product space to a (finite) measure, where the
-  product space can come with an arbitraty index set.}  The next
+product space can come with an arbitraty index set.  The next
 definition covers the important concept of a projective family of
@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def IsProjective
-    [Preorder ι]
+def IsProjective [Preorder ι]
     (P : ∀ j : ι, α j) (π : ∀ {i j : ι}, j ≤ i → α i → α j) : Prop :=
@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-  def cylinders : Set (Set ((i : ι) → α i)) := ⋃ (s) (S) (_ :
-  MeasurableSet S), {cylinder s S}
+def cylinders : Set (Set ((i : ι) → α i)) :=
+  ⋃ (s) (S) (_ : MeasurableSet S), {cylinder s S}
 \end{minted}
@@
   cover $A$), and therefore has a compact closure as noted
-  {\color{blue} in the lines directly preceding the lemma}. In
+  in the lines directly preceding the lemma. In
   addition, by $\sigma$-sub-additivity of $\mu$,
@@
     (P : ∀ i, Measure (α i))
-    [∀ i, IsProbabilityMeasure (P i)] : Measure (∀ i, α i)
-    := projectiveLimitWithWeakestHypotheses
+    [∀ i, IsProbabilityMeasure (P i)] : Measure (∀ i, α i) :=
+  projectiveLimitWithWeakestHypotheses
     (fun J : Finset ι => Measure.subset_pi P J)

### 2023-09-14 Manuscript/proof_outline.tex
fix: Ionescu-Tulcea is 1 person, not 2

@@
   complete and separable extended pseudo-metric spaces, but note that
-  an implementation of the Theorem by Ionescu and Tulcea would give
+  an implementation of the Ionescu-Tulcea Theorem would give
   the same result without these prerequisites.) Here, we can build on

### 2023-09-14 Manuscript/proof_outline.tex
expand todo
rewrite the topology remark. Also fix some spaces in leanline

@@
 
-\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}
+\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ? Explain somewhere that this is the mathlib notation for the product space}
 
@@
   \begin{enumerate}
-    \item \sloppy In \leanline{Lean}, we have the \leanline{class
-      PseudoEMetricSpace (α : Type u) extends EDist}, where
-      \leanline{EDist} comes with an extended distance
+    \item \sloppy \leanline{mathlib} provides the \leanline{class PseudoEMetricSpace (α : Type u)},
+      which comes with an extended pseudo-distance
       \leanline{edist: α → α → ℝ≥0∞} (the notation \leanline{ℝ≥0∞}
-      stands for $[0, \infty]$), and the extension to
-      \leanline{PseudoEMetricSpace} consists of the components,
-      \leanline{edist_self}, \leanline{edist_comm},
-      \leanline{edist_triangle} providing the properties of the
-      pseudo-metric, and \leanline{toUniformSpace},
-      \leanline{uniformity_edist} defining a uniform space from the
-      extended pseudo-metric. Such a space does not come with a
+      stands for $[0, \infty]$), and the properties
+      \leanline{edist_self} (distance to self is zero),
+      \leanline{edist_comm} (commutativity) and
+      \leanline{edist_triangle} (triangle inequality).
+      That class also comes with \leanline{toUniformSpace} and
+      \leanline{uniformity_edist}, defining a uniform space from the
+      extended pseudo-metric.
+    \item A uniform space (\leanline{class UniformSpace (α : Type u)}) does not come with a
       metric, but with a filter on $\alpha \times \alpha$, which
@@
       \cite{melikhov2011metrizable}.
-    \item \sloppy For topological spaces, we have \leanline{class
-      TopologicalSpace(α : Type u) : Type u}, which comes with
-      \leanline{IsOpen : Prop}, \leanline{isOpen_univ : IsOpen
-        Set.univ}, \leanline{isOpen_inter} and
+    \item \sloppy A topological spaces is defined in \leanline{mathlib} as \leanline{class TopologicalSpace(α : Type u)},
+      which comes with a predicate \leanline{IsOpen}, and properties \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
       \leanline{isOpen_sUnion}, which are exactly the properties of a
       topological space described above.
-    \item Any (extended pseudo-)metric on $\alpha$ defines a topology,
-      namely the topology generated by $\mathcal H := \{\{y:
-      r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
-      (0,\infty)\}$. Actually, in \leanline{mathlib}, as we have seen
-      above, every extended pseudo-metric defines a uniform space, and
-      a uniform space is a \leanline{class UniformSpace(α : Type u)
-        extends TopologicalSpace , UniformSpace.Core} (where the
-      uniformity is defined in \leanline{UniformSpace.Core}), which
-      connects the \leanline{uniformity} with the topology using
-      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
-      uniform space defines a topological space, which can be used by
-      the typeclass system.
+      An uniform spaces defines a topology, given by
+      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}.
+%    \item Any (extended pseudo-)metric on $\alpha$ defines a topology,
+%      namely the topology generated by $\mathcal H := \{\{y:
+%      r(x,y)<\varepsilon\}: x \in \alpha, \varepsilon \in
+%      (0,\infty)\}$. Actually, in \leanline{mathlib}, as we have seen
+%      above, every extended pseudo-metric defines a uniform space, and
+%      a uniform space is a \leanline{class UniformSpace(α : Type u) extends TopologicalSpace, UniformSpace.Core} (where the
+%      uniformity is defined in \leanline{UniformSpace.Core}), which
+%      connects the \leanline{uniformity} with the topology using
+%      \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}. In particular, a
+%      uniform space defines a topological space, which can be used by
+%      the typeclass system.
   \end{enumerate}
@@
   n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
-is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin
-  (card J)) : Finset (Set α)} is the set consisting of the
+is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)} is the set consisting of the
 \leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
@@
 \leanline{toFun} on \leanline{Set α}, whereas
-\leanlinecolor{extendContent} comes with \leanline{m : ∀ s : Set α, s
-  ∈ C → ℝ≥0∞}.
+\leanlinecolor{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.
 
@@
 
-%The next result talks about an \leanline{induced_outer_measure} in
-%terms of {\tt mathlib}.
 The next result extends a set function on a semi-ring to an outer measure.

### 2023-09-14 Manuscript/proof_outline.tex
typos

@@
 \end{definition}
-\todo{reorder that next remark? I'll make a proposal}
+
 \begin{remark}[Metric and topological spaces in {\tt mathlib}] \label{rem:uniform}\mbox{}
@@
       \cite{melikhov2011metrizable}.
-    \item \sloppy A topological spaces is defined in \leanline{mathlib} as \leanline{class TopologicalSpace(α : Type u)},
+    \item \sloppy A topological space is defined in \leanline{mathlib} as \leanline{class TopologicalSpace(α : Type u)},
       which comes with a predicate \leanline{IsOpen}, and properties \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
@@
       topological space described above.
-      An uniform spaces defines a topology, given by
+      A uniform space defines a topology, given by
       \leanline{∀ (s : Set α), IsOpen s ↔ ∀ (x : α), x ∈ s → {p | p.fst = x → p.snd ∈ s} ∈ uniformity}.

### 2023-09-16 Manuscript/proof_outline.tex
mention a precise mathlib commit, with date

@@
 
-\todo[inline]{Mention a precise commit of Mathlib}
-
-\todo[inline]{write about sets or types?}
-
 \todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ? Explain somewhere that this is the mathlib notation for the product space}
@@
 The goal of our contribution is a formalization of the proof of that theorem in {\tt Lean}.
-Here, we are going to use \leanline{Lean4}, since \leanline{mathlib} \cite{mathlib2020lean}, the mathematical library of \leanline{Lean} has recently been ported to \leanline{Lean4}.
-%Subsequently, we aim to make a pull request for the publication of our code in \leanline{mathlib}.
+We using \leanline{Lean4} and its mathematical library \leanline{mathlib} \cite{mathlib2020lean}\footnote{All references to results in \leanline{mathlib} are accurate for commit {\tt 537e7f030d6df737d728c7c002fb1928f4615afa}, August 22, 2023.}.
 The Kolmogorov extension theorem is on the interface between measure theory and probability theory.

### 2023-09-16 Manuscript/proof_outline.tex
typo

@@
 The goal of our contribution is a formalization of the proof of that theorem in {\tt Lean}.
-We using \leanline{Lean4} and its mathematical library \leanline{mathlib} \cite{mathlib2020lean}\footnote{All references to results in \leanline{mathlib} are accurate for commit {\tt 537e7f030d6df737d728c7c002fb1928f4615afa}, August 22, 2023.}.
+We are using \leanline{Lean4} and its mathematical library \leanline{mathlib} \cite{mathlib2020lean}\footnote{All references to results in \leanline{mathlib} are accurate for commit {\tt 537e7f030d6df737d728c7c002fb1928f4615afa}, August 22, 2023.}.
 The Kolmogorov extension theorem is on the interface between measure theory and probability theory.

### 2023-09-16 Manuscript/proof_outline.tex
explain product space notation

@@
 
-{\color{blue}In contrast to statements already part of
+In contrast to statements already part of
 \leanline{mathlib}, in the sequel we \leanlinecolor{highlight}
-  definitions and statements which are part of our own code.} In the
-formalization of the above, we use \leanline{variable {ι : Type _} {α
-    : ι → Type _} }, which fixes the index set $\iota$ and all spaces
-$\alpha_t, t\in\iota$ as global variables. In addition, note that the
-projective property we use here works as long as we have a preorder
+definitions and statements which are part of our own code. In the
+formalization of the above, we use \leanline{variable {ι : Type _} {α: ι → Type _}},
+which fixes the index set $\iota$ and all spaces
+$\alpha_t, t\in\iota$ as global variables.
+In Lean, the product space is then \leanline{∀ j : ι, α j} and the product over a finite set \leanline{J} is \leanline{∀ j : J, α j}.
+
+We formalize the projective property as follows. This definition is not specific to measures but works as long as we have a preorder
 (which is the subset relation on \leanline{Finset ι} below).
@@
 \noindent
-With this, we can define the projetive family as follows. The typeclass inference system of
+With this, we can define the projective family as follows. The typeclass inference system of
 \leanline{Lean} automatically uses the subset relation to generate

### 2023-09-16 Manuscript/proof_outline.tex
use same product notation everywhere

@@
     (s : Finset ι) (S : Set (∀ i : s, α i)) :
-    Set ((i : ι) → α i) :=
-  (fun f : (i : ι) → α i => fun i : s => f i) ⁻¹' S
+    Set (∀ i : ι, α i) :=
+  (fun (f : ∀ j : ι, α j) => fun i : s => f i) ⁻¹' S
 \end{minted}
@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-def cylinders : Set (Set ((i : ι) → α i)) :=
+def cylinders : Set (Set (∀ i : ι, α i)) :=
   ⋃ (s) (S) (_ : MeasurableSet S), {cylinder s S}
@@
 theorem mem_cylinders
-    (t : Set ((i : ι) → α i)) :
+    (t : Set (∀ i : ι, α i)) :
     t ∈ cylinders α ↔
@@
     (P : ∀ i, Measure (α i)) (I : Finset ι) :
-    Measure ((i : I) → (α i)) :=
+    Measure (∀ i : I, α i) :=
   Measure.pi (fun (i : I) => (P (i : ι)))

### 2023-09-16 Manuscript/proof_outline.tex
remove todo
remove remark about type theory in Isabelle vs Lean
rewrite remark about extension to pseudo-metric spaces

@@
 
-\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ? Explain somewhere that this is the mathlib notation for the product space}
-
 \section{Introduction}
@@
 This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical.
-In addition, {\color{blue}Isabelle works with simple type theory, while Lean is an implementation of dependent theory.}
+%In addition, Isabelle/HOL works with simple type theory, while Lean is an implementation of dependent type theory.
 
@@
 Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
-open balls around $x$ and $y$ overlap. {\color{blue}We note that this
-  generalization was only possible since underlying results in
+open balls around $x$ and $y$ overlap.
+  This generalization was only possible since underlying results in
   \leanline{mathlib} were already provided on the same level of
-  generality: More precisely,
-  \leanline{isCompact_iff_totallyBounded_isComplete}, which shows that
+  generality.
+  More precisely, \leanline{isCompact_iff_totallyBounded_isComplete}, which shows that
   a set $A \subseteq \alpha$ is compact iff it is complete and totally
-  bounded {\color{blue}(see note~\ref{note:tot})}, requires $\alpha$
-  to be a uniform space (recall that every metric space is uniform);
-  recall such spaces from Remark~\ref{rem:uniform}, and see
-  note~\ref{note:tot} for a definition of total boundedness. Since we
-  also require the underlying space(s) to be second countable (see
-  also in the proof of Lemma~\ref{L:relcoPol}), we have to make a
-  countabiliy assumption, which leads us to (extended) pseudo-metric
-  spaces by \leanline{UniformSpace.pseudoMetrizableSpace}; see also
-  Remark~\ref{rem:uniform} for some more details.}
+  bounded (see note~\ref{note:tot}), requires $\alpha$
+  to be a uniform space (Remark~\ref{rem:uniform}, recall that every metric space is uniform).
+  We also require the underlying space(s) to be second-countable (used in the proof of Lemma~\ref{L:relcoPol}).
+  A second-countable uniform space can be made into an (extended) pseudo-metric space
+  (\leanline{UniformSpace.pseudoMetrizableSpace}); see also
+  Remark~\ref{rem:uniform} for some more details.
 

### 2023-09-16 Manuscript/proof_outline.tex
remove class/structure from leanline code

@@
   \begin{enumerate}
-    \item \sloppy \leanline{mathlib} provides the \leanline{class PseudoEMetricSpace (α : Type u)},
+    \item \sloppy \leanline{mathlib} provides the class \leanline{PseudoEMetricSpace α},
       which comes with an extended pseudo-distance
@@
       extended pseudo-metric.
-    \item A uniform space (\leanline{class UniformSpace (α : Type u)}) does not come with a
+    \item A uniform space (\leanline{UniformSpace α}) does not come with a
       metric, but with a filter on $\alpha \times \alpha$, which
@@
       \cite{melikhov2011metrizable}.
-    \item \sloppy A topological space is defined in \leanline{mathlib} as \leanline{class TopologicalSpace(α : Type u)},
+    \item \sloppy A topological space is defined in \leanline{mathlib} as \leanline{TopologicalSpace α},
       which comes with a predicate \leanline{IsOpen}, and properties \leanline{isOpen_univ : IsOpen Set.univ}, \leanline{isOpen_inter} and
@@
   \begin{enumerate}
-  \item \sloppy The \leanline{class MeasurableSpace α} is very similar
+  \item \sloppy The class \leanline{MeasurableSpace α} is very similar
     to a topological space in \leanline{mathlib}, since it comes with
@@
   \item We are using outer measures in our construction. In
-    \leanline{mathlib}, this is \leanline{structure MeasureTheory.OuterMeasure α} which is a function
+    \leanline{mathlib}, this is \leanline{MeasureTheory.OuterMeasure α} which is a structure with a function
     \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf ∅ = 0} (i.e.\ the empty set has measure~$0$),

### 2023-09-16 Manuscript/proof_outline.tex
remove blue

@@
 i.e.\ every ring is a $\pi$-system.
-Using \leanline{variable {α : Type _} {s t : Set α}}, we define a semi-ring as follows:
+We define a semi-ring on a type \leanline{α} as follows:
 
@@
 \leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
-\subseteq \mathcal F$ in {\color{blue} the construction of a product
-  measure for a finite index set}. In addition, there is
+\subseteq \mathcal F$ in the construction of a product
+  measure for a finite index set. In addition, there is
 \leanline{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal

### 2023-09-16 Manuscript/proof_outline.tex
rewrite explanation of addContent and extendContent

@@
 \noindent
-For additive contents, we need two
-definitions. \leanlinecolor{AddContent} defines a set-function
-\leanline{toFun} on \leanline{Set α}, whereas
-\leanlinecolor{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.
+For additive contents, we need two definitions.
+The structure \leanlinecolor{AddContent} defines
+an additive set-function \leanline{toFun} on \leanline{Set α}.
+The definition \leanlinecolor{extendContent} takes an additive function defined only on a semi-ring of sets, and extends it to an additive content on all sets by choosing the value infinity outside of the semi-ring.
+We use this because total functions are easier to use in subsequent code.
 

### 2023-09-16 Manuscript/proof_outline.tex
minor
manuscript semi-final

@@
 
-\documentclass[sigplan,screen]{acmart}
+% Anonymous paper:
+% \documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+
+% Paper with authors:
+\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+
 \usepackage[utf8x]{inputenc}
 \usepackage[T1]{fontenc}
-\usepackage{ulem}
+%\usepackage{ulem}
 %\usepackage{amsmath}
@@
 \definecolor{LightCyan}{rgb}{0.8784,1,1}
-\newcommand{\leanline}[1]{\mintinline[breaklines, breakafter=., breaklines]{Lean}{#1}}%\textsuperscript{\textcolor{gray}{\tiny(m)}}}
-\newcommand{\leanlinecolor}{\mintinline[breaklines, breakafter=., breaklines, bgcolor=LightCyan]{Lean}}
+\newcommand{\leanline}[1]{\mintinline[breaklines, breakafter=_., breaklines]{Lean}{#1}}%\textsuperscript{\textcolor{gray}{\tiny(m)}}}
+\newcommand{\leanlinecolor}{\mintinline[breaklines, breakafter=_., breaklines, bgcolor=LightCyan]{Lean}}
 %\usepackage{refcheck}
@@
 \begin{abstract}
-  We present a formalization in Lean of the Kolmogorov extension theorem,
-  which is a main building block in the construction of the (distribution of)
-  stochastic processes with arbitrary index sets. Our
-  approach is based on mathlib, the mathematical library for Lean. On
-  our way, we provide a formalization of the Caratheodory extension
-  theorem, which allows to extend a set function on a semi-ring to a
-  proper measure on the $\sigma$-algebra generated by the semi-ring. A
-  slight generalization of the classical Kolmogorov extension theorem
-  allows us to construct the (distribution of) a stochastic process on
-  a complete and separable pseudo-metric space rather than a metric
-  space.
+  We present a formalization of the Kolmogorov extension theorem in
+  Lean. This result is a main building block in the construction of
+  the (distribution of) stochastic processes with arbitrary index
+  sets. Our approach is based on mathlib, the mathematical library for
+  Lean. On our way, we provide a formalization of the Caratheodory
+  extension theorem, which allows to extend a set function on a
+  semi-ring to a proper measure on the $\sigma$-algebra generated by
+  the semi-ring. A slight generalization of the classical Kolmogorov
+  extension theorem allows us to construct the (distribution of) a
+  stochastic process on a complete and separable pseudo-metric space
+  rather than a metric space. {\color{blue}As an example, we define a
+    product probability measure with arbitrary index set.}
 \end{abstract}
@@
 So, our work complements \cite{ying2023formalization} in the sense that probability spaces $(\Omega, \mathcal F, \mathbf P)$ on which you can define a martingale exist.
-{\color{blue}As an application of our implementation, we construct a probability measure on an infinite product space where all coordinates are independent; see Section~\ref{ss:concl}.}
+As an application of our implementation, we construct a probability measure on an infinite product space where all coordinates are independent; see Section~\ref{ss:concl}.
+{\color{blue} Our code is available at xxx.}
+
+~
 
@@
 A formalization would complement our work.
-In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} without requirements on $(\alpha_t)_{t\in\iota}$.
+In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} {\color{blue} with fewer requirements on $(\alpha_t)_{t\in\iota}$ than in the example given in Section~\ref{ss:concl}}.
 All of these results can be used in order to construct countably many independent random variables, which are frequently used in concrete constructions in probability theory.
@@
 formalized with \leanline{Finset ι}. % Such a \leanline{Finset} is
-defined as a \leanline{Multiset} (a
-%quotient type of the \leanline{List}-type with respect to permutations
-%of the list) together with a proof that it contains no duplicates\todo{remove this implementation detail?}.
+defined as a \leanline{Multiset}.
+%(a quotient type of the \leanline{List}-type with respect to
+%permutations of the list) together with a proof that it contains no
+%duplicates\todo{remove this implementation detail?}.
 
@@
 
-In contrast to statements already part of
-\leanline{mathlib}, in the sequel we \leanlinecolor{highlight}
-definitions and statements which are part of our own code. In the
-formalization of the above, we use \leanline{variable {ι : Type _} {α: ι → Type _}},
-which fixes the index set $\iota$ and all spaces
-$\alpha_t, t\in\iota$ as global variables.
-In Lean, the product space is then \leanline{∀ j : ι, α j} and the product over a finite set \leanline{J} is \leanline{∀ j : J, α j}.
+In contrast to statements already part of \leanline{mathlib}, in the
+sequel we \leanlinecolor{highlight} definitions and statements which
+are part of our own code. In the formalization of the above, we use
+\leanline{variable {ι : Type _} {α: ι → Type _}}, which fixes the
+index set $\iota$ and all spaces $\alpha_t, t\in\iota$ as global
+variables.  In Lean, the product space is then \leanline{∀ j : ι, α j}
+and the product over a finite set \leanline{J} is \leanline{∀ j : J, α
+  j}. {\color{blue} Note that \leanline{J} is a subtype of
+  \leanline{ι} in this formulation.}
 
@@
 \leanline{=>}-notation maps every $(x_i)_{i\in I}$ to a function of
-\leanline{j}, whose type is the subtype of \leanline{I}, consisting of
-a value and a proof of \leanline{J ⊆ I}. In other words, this is
-$(x_j)_{j\in J}$.
+\leanline{j}, whose type is the subtype {\color{blue}\leanline{j} of}
+\leanline{I}, consisting of a value and a proof of \leanline{J ⊆
+  I}. In other words, this is $(x_j)_{j\in J}$.
 
@@
 
-We note that we extend the standard assumption that all $\alpha_t$ are
+\noindent
+Note that we extend the standard assumption that all $\alpha_t$ are
 separable, complete metric spaces (or Polish, i.e.\ separable and
@@
 Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
-open balls around $x$ and $y$ overlap.
-  This generalization was only possible since underlying results in
-  \leanline{mathlib} were already provided on the same level of
-  generality.
-  More precisely, \leanline{isCompact_iff_totallyBounded_isComplete}, which shows that
-  a set $A \subseteq \alpha$ is compact iff it is complete and totally
-  bounded (see note~\ref{note:tot}), requires $\alpha$
-  to be a uniform space (Remark~\ref{rem:uniform}, recall that every metric space is uniform).
-  We also require the underlying space(s) to be second-countable (used in the proof of Lemma~\ref{L:relcoPol}).
-  A second-countable uniform space can be made into an (extended) pseudo-metric space
-  (\leanline{UniformSpace.pseudoMetrizableSpace}); see also
-  Remark~\ref{rem:uniform} for some more details.
+open balls around $x$ and $y$ overlap.  This generalization was only
+possible since underlying results in \leanline{mathlib} were already
+provided on the same level of generality.  More precisely,
+\leanline{isCompact_iff_totallyBounded_isComplete}, which shows that a
+set $A \subseteq \alpha$ is compact iff it is complete and totally
+bounded (see note~\ref{note:tot}), requires $\alpha$ to be a uniform
+space (Remark~\ref{rem:uniform}, recall that every metric space is
+uniform).  We also require the underlying space(s) to be
+second-countable (used in the proof of Lemma~\ref{L:relcoPol}).  A
+second-countable uniform space can be made into an (extended)
+pseudo-metric space (\leanline{UniformSpace.pseudoMetrizableSpace});
+see also Remark~\ref{rem:uniform} for some more details.
 
@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-structure SetSemiring (C : Set (Set α)) : Prop where
+structure SetSemiring (C : Set (Set α))
+    : Prop where
   empty_mem : ∅ ∈ C
@@
 
+\noindent
 Membership of that set is characterized as follows:
@@
 The finite set \leanline{s} of the potential coordinates where
-the cylinder deviates from the whole set needs not be unique.
-From the exists-statement of \leanline{mem_cylinders}, we however choose an
+the cylinder deviates from the whole set need not be unique.
+From the $\exists$-statement of \leanline{mem_cylinders}, we however choose an
 arbitrary such \leanline{Finset ι}, called \leanlinecolor{cylinders.finset}.
@@
 
-\begin{minted}[highlightlines={1-3}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+\begin{minted}[highlightlines={1-4}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 theorem setField_cylinders : SetField (cylinders α) 
 theorem setRing_cylinders :  SetRing (cylinders α)
-theorem setSemiringCylinders : SetSemiring (cylinders α)   
+theorem setSemiringCylinders
+  : SetSemiring (cylinders α)   
 \end{minted}
@@
 
+\noindent
 Extending Lemma~\ref{l1}, we would like to write a finite union of
-members of a semi-ring as a finite union of disjoint sets\todo{Is this section going into too much detail?}.
+members of a semi-ring as a finite union of disjoint sets.
 
@@
 
-\noindent
-The formalized version again gives rise to a definition:
-
-\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-def indexedDiff₀ (hC : SetSemiring C)
-    (J : Finset (Set α)) (hJ : ↑J ⊆ C)
-    (n : Fin J.card) : Finset (Set α) :=
-  hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
-    (finsetLT_subset' J hJ n)
-\end{minted}
-
-Here, recall that \leanline{Fin n} is the subtype of \leanline{Nat}
-consisting of all numbers \leanline{<n}. In addition,
-\leanline{ordered_mem' (hJ : ↑J ⊆ C) (n : Fin (card J)) : ↑(ordered J)
-  n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
-is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)} is the set consisting of the
-\leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
-\leanlinecolor{finsetLT_subset'} gives a proof that
-\leanline{↑(finsetLT J n) ⊆ C}.  So, in terms of Lemma~\ref{l2}, this
-gives, for $n=1,...,m$, a construction for $\mathcal K_n \subseteq_f
-\mathcal H$.
+%% \noindent
+%% The formalized version again gives rise to a definition:
+
+%% \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+%% def indexedDiff₀ (hC : SetSemiring C)
+%%     (J : Finset (Set α)) (hJ : ↑J ⊆ C)
+%%     (n : Fin J.card) : Finset (Set α) :=
+%%   hC.diff₀ (ordered_mem' hJ n) (finsetLT J n)
+%%     (finsetLT_subset' J hJ n)
+%% \end{minted}
+
+%% Here, recall that \leanline{Fin n} is the subtype of \leanline{Nat}
+%% consisting of all numbers \leanline{<n}. In addition,
+%% \leanline{ordered_mem' (hJ : ↑J ⊆ C) (n : Fin (card J)) : ↑(ordered J)
+%%   n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
+%% is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)} is the set consisting of the
+%% \leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
+%% \leanlinecolor{finsetLT_subset'} gives a proof that
+%% \leanline{↑(finsetLT J n) ⊆ C}.  So, in terms of Lemma~\ref{l2}, this
+%% gives, for $n=1,...,m$, a construction for $\mathcal K_n \subseteq_f
+%% \mathcal H$.
 
@@
 \subsection{Carathéodory's extension theorem}
-Measures are set functions defined on a $\sigma$-algebra (i.e.\ an
-algebra stable under countable unions), satisfying some properties
-which we recall below. Mostly, defining such a measure on the full
-$\sigma$-algebra is not possible, but defining a set function on a
-semi-ring is possible. As an example, consider Lebesgue-measure on
-$\mathbb R$, with the Borel $\sigma$-algebra on $\mathbb R$, which is
+Measures are set functions defined on a $\sigma$-algebra, satisfying
+some properties which we recall below. {\color{blue}Frequently,
+  constructively} defining such a measure on the full $\sigma$-algebra
+is not possible, but defining a set function on a semi-ring is
+possible. As an example, consider Lebesgue-measure on $\mathbb R$,
+with the Borel $\sigma$-algebra on $\mathbb R$, which is
 $\sigma(\mathcal O)$, where $\mathcal O$ is the set of open sets, and
@@
 (outer measure) $\mu$ on $2^\alpha$, and then show $\mu$ extends $m$
-and defines a measure on $\sigma(\mathcal H)$. (Note that the set of
-semi-open intervals generates the Borel $\sigma$-algebra on $\mathbb
-R$.) We will follow this abstract construction, and start by stating
-some basic concepts.
+and defines a measure on {\color{blue}the $\sigma$-algebra of
+  measurable sets (see Theorem~\ref{T:cara} below), which contains}
+$\sigma(\mathcal H)$.
+%(Note that the set of semi-open intervals
+%generates the Borel $\sigma$-algebra on $\mathbb R$.)
+We will follow this abstract construction, and start by stating some
+basic concepts.
 
@@
 The structure \leanlinecolor{AddContent} defines
-an additive set-function \leanline{toFun} on \leanline{Set α}.
+an additive set-function on \leanline{Set α}.
 The definition \leanlinecolor{extendContent} takes an additive function defined only on a semi-ring of sets, and extends it to an additive content on all sets by choosing the value infinity outside of the semi-ring.
@@
 Here, \leanlinecolor{kolmogorovFun_empty} and
-\leanlinecolor{kolmogorovFun_additive} give proofs for
+\leanlinecolor{kolmogorovFun_⌋} \leanlinecolor{additive} give proofs for
 \leanline{m_empty} and \leanline{m_add} in
@@
   i.e.\ the set of sets $A$ satisfying
-  \begin{align}
-    \label{eq:mumess}
+  \begin{align*}
     \mu(B) = \mu(B\cap A) + \mu(B\cap A^c), \qquad B \subseteq E.
-  \end{align}
+  \end{align*}
   Then, $\mathcal F$ is a $\sigma$-Algebra and $\mu|_{\mathcal F}$ is
@@
 
+\noindent
 Here is a formalization of the measure resulting from
@@
 
+\noindent
 The above definition closely resembles its formalization in
@@
 
+\noindent
 Here is the formalization:
@@
 \end{minted}
+
+\noindent
+{\color{blue} Such compact systems are important since they allow a
+  proof of $\sigma$-additivity of a content on a ring, which is the
+  missing piece for applying the Carathéodory Theorem in the proof of
+  the Kolmogorov extension theorem.}
 
@@
 (rather than $\sigma$-additivity), we can use
-\leanlinecolor{sigma_subadditive_of_sigma_additive} in order
-to show
-\leanlinecolor{kolContent_sigma_subadditive_of_innerRegular}, and
-the assumptions of the last result also imply $\sigma$-subadditivity.
+\leanlinecolor{sigma_subadditive_of_sigma_additive} {\color{blue}(see
+  Lemma~\ref{P:stetigmass}, 1.$\Rightarrow$2.)} in order to show that
+the assumptions of the last result also imply $\sigma$-subadditivity;
+this gives
+\leanlinecolor{kolContent_sigma_subadditive_of_innerRegular}.
 
@@
 closed) sets. For this, we need some assumption on the underlying
-spaces. The following lemma provides the generalizations to extended
-pseudo-metric spaces:
+spaces.
 
@@
 the formalization of the next lemma only in conjunction with
-Lemma~\ref{L:relcoPol}.
+Lemma~\ref{L:relcoPol}. {\color{blue} We omit the proof, since the
+  only interesting part is an application of inner regularity from
+  Lemma~\ref{l:pseude1}.}
 
@@
 
-\todo{too much detail? omit the proof?}
-
-\begin{proof}
-  2. $\Rightarrow$ 1.: This is clear since $\alpha \in \mathcal B$.
-  1. $\Rightarrow$ 2.: Fix $\varepsilon>0$ and $A \in \mathcal B$. Let
-  $K$ be closed and compact with $\mu(K^c) < \varepsilon/2$. By
-  Lemma~\ref{l:pseude1}, there is some closed $F \subseteq A$
-  with $\mu(A) < \mu(F) + \varepsilon/2$. Now, we have that $F \cap K
-  \subseteq A$ is closed and compact, and with $A \setminus (F\cap K)
-  = (A \setminus F) \cup (A \setminus K)$,
-  $$\mu(A \setminus (F\cap K)) \leq \mu(A \setminus F) + \mu(A
-  \setminus K) < \varepsilon/2 + \mu(K^c) < \varepsilon. $$
-\end{proof}
+%% \todo{too much detail? omit the proof?}
+
+%% \begin{proof}
+%%   2. $\Rightarrow$ 1.: This is clear since $\alpha \in \mathcal B$.
+%%   1. $\Rightarrow$ 2.: Fix $\varepsilon>0$ and $A \in \mathcal B$. Let
+%%   $K$ be closed and compact with $\mu(K^c) < \varepsilon/2$. By
+%%   Lemma~\ref{l:pseude1}, there is some closed $F \subseteq A$
+%%   with $\mu(A) < \mu(F) + \varepsilon/2$. Now, we have that $F \cap K
+%%   \subseteq A$ is closed and compact, and with $A \setminus (F\cap K)
+%%   = (A \setminus F) \cup (A \setminus K)$,
+%%   $$\mu(A \setminus (F\cap K)) \leq \mu(A \setminus F) + \mu(A
+%%   \setminus K) < \varepsilon/2 + \mu(K^c) < \varepsilon. $$
+%% \end{proof}
 
@@
 
-Let us combine the last two results in one combined formalized result.
-This the gives the desired result that a measure on a second
-countable, complete extended pseudo-metric space is inner regular with
-respect to closed compact sets.
-
-\begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+Let us combine the last two results in one formalized result.  This
+the gives the desired statement that a measure on a second countable,
+complete extended pseudo-metric space is inner regular with respect to
+closed compact sets.
+
+\begin{minted}[highlightlines={1-2}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 theorem innerRegular_isCompact_isClosed_
@@
     P_J(C)$, so we need to show that $P_J$ is inner regular with
-    respect to compact and closed sets, $J\subseteq_f \iota$;
+    respect to compact closed sets, $J\subseteq_f \iota$;
   \item use Lemma~\ref{L:relcoPol} in combination with
@@
 \sloppy Let us look closer at the formalized proof:
-\leanlinecolor{kolContent_sigma_subadditive_of_innerRegular} has
-the same hypotheses as
-\leanlinecolor{kolContent_sigma_additive_of_innerRegular}, except
-\leanline{h_disj}. So, the lemma in the last brackets shows
+\leanlinecolor{kolContent_sigma_subadditive_of_innerRegular} has the
+same hypotheses as
+\leanlinecolor{kolContent_sigma_additive_of_innerRegular},
+except \leanline{h_disj}. So, the lemma in the last brackets shows
 $\sigma$-sub-additivity of the content $P$ from
@@
 \label{ss:concl}
-{\color{blue} The usability of our contribution must be proved by an
-  example: Adding to our general theory, we have implemented the
-  example of an infinite product measure. (As an application of
-  \leanlinecolor{projectiveLimitWithWeakestHypotheses}, we have to use
-  complete and separable extended pseudo-metric spaces, but note that
-  an implementation of the Ionescu-Tulcea Theorem would give
-  the same result without these prerequisites.) Here, we can build on
-  \leanline{Measure.pi}, which defines a finite product measure.
-  Using the same automatically inferred instances, we obtain:}
-
-\begin{minted}[highlightlines={1,8}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
+The usability of our contribution must be proved by an example: Adding
+to our general theory, we have implemented the example of an infinite
+product measure.
+%(As an application of
+%\leanlinecolor{projectiveLimitWithWeakestHypotheses}, we have to
+%use complete and separable extended pseudo-metric spaces, but note
+%that an implementation of the Ionescu-Tulcea Theorem would give the
+%same result without these prerequisites.)
+Here, we can build on \leanline{Measure.pi}, which defines a finite
+product measure.  Using the same automatically inferred instances,
+we obtain:
+
+\begin{minted}[highlightlines={1,8,13}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
 def independentFamily
-    (P : ∀ i, Measure (α i))
-    [∀ i, IsProbabilityMeasure (P i)] : Measure (∀ i, α i) :=
+    (P : ∀ i, Measure (α i)) [∀ i,
+    IsProbabilityMeasure (P i)] : Measure (∀ i, α i) :=
   projectiveLimitWithWeakestHypotheses
-    (fun J : Finset ι => Measure.subset_pi P J)
-    (product_isProjective P)
+  (fun J : Finset ι =>
+  Measure.subset_pi P J) (product_isProjective P)
 
@@
   Measure.pi (fun (i : I) => (P (i : ι)))
+
+theorem product_isProjective
+  (P : ∀ i, Measure (α i)) [∀ i, IsProbabilityMeasure (P i)] :
+  IsProjectiveMeasureFamily
+  (fun J : Finset ι => Measure.subset_pi P J)
+
 \end{minted}
@@
 \noindent
-{\em Prohorovs theorem:} The proof that single finite measures on a
+{\em Prohorovs theorem:} The proof that a single finite measures on a
 Polish space is tight (i.e.\ inner regular with respect to compact
@@
 
-~~
-
 \section*{Acknowledgements}
+{\color{blue}PP ist partly funded by the Vector foundation, grant {\em
+    LeanAI (P2022-0241)}}.
 

### 2023-09-17 Manuscript/proof_outline.tex
accept most of the blue, rewrite one or two small passages

@@
   stochastic process on a complete and separable pseudo-metric space
-  rather than a metric space. {\color{blue}As an example, we define a
-    product probability measure with arbitrary index set.}
+  rather than a metric space. As an example, we define a
+  product probability measure with arbitrary index set.
 \end{abstract}
@@
 A formalization would complement our work.
-In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} {\color{blue} with fewer requirements on $(\alpha_t)_{t\in\iota}$ than in the example given in Section~\ref{ss:concl}}.
+In addition, using the Theorem of Ionescu-Tulcea one can define infinitely many independent random variables (arbitrary $\iota$) on a joint probability space \cite[Theorem 8.24]{Kallenberg2020} with fewer requirements on $(\alpha_t)_{t\in\iota}$ than in the example given in Section~\ref{ss:concl}.
 All of these results can be used in order to construct countably many independent random variables, which are frequently used in concrete constructions in probability theory.
@@
 variables.  In Lean, the product space is then \leanline{∀ j : ι, α j}
-and the product over a finite set \leanline{J} is \leanline{∀ j : J, α
-  j}. {\color{blue} Note that \leanline{J} is a subtype of
-  \leanline{ι} in this formulation.}
+and the product over a finite subset \leanline{J} of \leanline{ι} is \leanline{∀ j : J, α
+  j}.
 
@@
 measure on \leanline{J}, which is \leanline{μ.map (fun x : ∀ i : I, α
-  i => fun j => x ⟨j, hJI j.2⟩)}.  The map defined by the
-\leanline{=>}-notation maps every $(x_i)_{i\in I}$ to a function of
-\leanline{j}, whose type is the subtype {\color{blue}\leanline{j} of}
-\leanline{I}, consisting of a value and a proof of \leanline{J ⊆
-  I}. In other words, this is $(x_j)_{j\in J}$.
+  i => fun j => x ⟨j, hJI j.2⟩)}.  The function used in that image
+measure maps every $(x_i)_{i\in I}$ to $(x_i)_{i\in J}$.
+%map defined by the
+%\leanline{=>}-notation maps every $(x_i)_{i\in I}$ to a function of
+%\leanline{j}, whose type is the subtype {\color{blue}\leanline{j} of}
+%\leanline{I}, consisting of a value and a proof of \leanline{J ⊆
+%  I}. In other words, this is $(x_j)_{j\in J}$.
 
@@
 Measures are set functions defined on a $\sigma$-algebra, satisfying
-some properties which we recall below. {\color{blue}Frequently,
-  constructively} defining such a measure on the full $\sigma$-algebra
+some properties which we recall below. Frequently,
+constructively defining such a measure on the full $\sigma$-algebra
 is not possible, but defining a set function on a semi-ring is
@@
 (outer measure) $\mu$ on $2^\alpha$, and then show $\mu$ extends $m$
-and defines a measure on {\color{blue}the $\sigma$-algebra of
-  measurable sets (see Theorem~\ref{T:cara} below), which contains}
+and defines a measure on the $\sigma$-algebra of
+  measurable sets (see Theorem~\ref{T:cara} below), which contains
 $\sigma(\mathcal H)$.
@@
 \noindent
-{\color{blue} Such compact systems are important since they allow a
+Such compact systems are important since they allow a
 proof of $\sigma$-additivity of a content on a ring, which is the
 missing piece for applying the Carathéodory Theorem in the proof of
-  the Kolmogorov extension theorem.}
+the Kolmogorov extension theorem.
 
@@
 (rather than $\sigma$-additivity), we can use
-\leanlinecolor{sigma_subadditive_of_sigma_additive} {\color{blue}(see
-  Lemma~\ref{P:stetigmass}, 1.$\Rightarrow$2.)} in order to show that
+\leanlinecolor{sigma_subadditive_of_sigma_additive} (see
+  Lemma~\ref{P:stetigmass}, 1.$\Rightarrow$2.) in order to show that
 the assumptions of the last result also imply $\sigma$-subadditivity;
@@
 the formalization of the next lemma only in conjunction with
-Lemma~\ref{L:relcoPol}. {\color{blue} We omit the proof, since the
+Lemma~\ref{L:relcoPol}. We omit the proof, since the
   only interesting part is an application of inner regularity from
-  Lemma~\ref{l:pseude1}.}
+  Lemma~\ref{l:pseude1}.
 
@@
   cover $A$), and therefore has a compact closure as noted
-  in the lines directly preceding the lemma. In
-  addition, by $\sigma$-sub-additivity of $\mu$,
+  in the lines directly preceding the lemma. Finally, by $\sigma$-sub-additivity of $\mu$,
   \begin{align*}
@@
      \sum_{n=1}^\infty \mu\Big( \Big(\bigcup_{k=1}^{N_n}
-     B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon,
+     B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon
+     \: .
   \end{align*}
- and we are done.
 \end{proof}
 
-Let us combine the last two results in one formalized result.  This
-the gives the desired statement that a measure on a second countable,
+Combining the last two results, we obtain the desired statement that a measure on a second countable,
 complete extended pseudo-metric space is inner regular with respect to
@@
 \section*{Acknowledgements}
-{\color{blue}PP ist partly funded by the Vector foundation, grant {\em
-    LeanAI (P2022-0241)}}.
+PP ist partly funded by the Vector foundation, grant
+{\em LeanAI (P2022-0241)}.
+RD acknowledges the support of the French Ministry of Higher Education and Research and the Hauts-de-France region.
 

### 2023-09-17 Manuscript/proof_outline.tex
switch to anonymous, comment out acks

@@
 % Anonymous paper:
-% \documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
 % Paper with authors:
-\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
@@
 
-\section*{Acknowledgements}
-PP ist partly funded by the Vector foundation, grant
-{\em LeanAI (P2022-0241)}.
-RD acknowledges the support of the French Ministry of Higher Education and Research and the Hauts-de-France region.
+%\section*{Acknowledgements}
+%PP ist partly funded by the Vector foundation, grant
+%{\em LeanAI (P2022-0241)}.
+%RD acknowledges the support of the French Ministry of Higher Education and Research and the Hauts-de-France region.
 

### 2023-09-18 Manuscript/proof_outline.tex
add link to anonymous repository
corrected typos in the manuscript

@@
 As an application of our implementation, we construct a probability measure on an infinite product space where all coordinates are independent; see Section~\ref{ss:concl}.
-{\color{blue} Our code is available at xxx.}
+Our code is available at \url{https://anonymous.4open.science/r/kolmogorov_extension4-5F4B}.
 
@@
 Other Gaussian processes -- indexed by $[0, \infty)$ or $\mathbb R$ -- might as well be constructed the same way.
-In addition, fields like the Gaussien free field indexed by $\mathbb R^d$ \cite{sheffield2007gaussian} can also be given.
+In addition, fields like the Gaussian free field indexed by $\mathbb R^d$ \cite{sheffield2007gaussian} can also be given.
 For these tasks, we would have to define the finite-dimensional distributions (Poisson and normally distributed, respectively), and apply the extension theorem.
@@
 Examples are (one direction of) the Borel-Cantelli-Lemma \cite[Theorem 4.28]{Kallenberg2020}, Kolmogorov's 0-1-law \cite[Theorem 4.13]{Kallenberg2020}, random walks \cite[Chapter 12]{Kallenberg2020}, branching processes  \cite[Chapter 13]{Kallenberg2020}, and percolation  \cite[Chapter 2.4]{Klenke2013}, to name just a few.
-While the first two are already formalized in \leanline{mathlib}, it is only clear due to the Kolmorov extension theorem (or the Theorem of Ionesu-Tulcea) that infinitely many independent random variables can indeed be defined on a joint probability space.
-In this sense, our contribution is also important for giving more sense to already formalised parts of \leanline{mathlib}.
+While the first two are already formalized in \leanline{mathlib}, it is only clear due to the Kolmogorov extension theorem (or the Theorem of Ionesu-Tulcea) that infinitely many independent random variables can indeed be defined on a joint probability space.
+In this sense, our contribution is also important for giving more sense to already formalized parts of \leanline{mathlib}.
 
@@
       for all $m,n > N$. The (extended pseudo-) metric is called
-      complete if every Cauchy-sequency has a limit in
+      complete if every Cauchy-sequence has a limit in
       $\alpha$.
@@
 set-function on a product space to a (finite) measure, where the
-product space can come with an arbitraty index set.  The next
+product space can come with an arbitrary index set.  The next
 definition covers the important concept of a projective family of
@@
 metrizable through a complete metric) to cover the case of extended
-pseude-metric spaces. Such spaces do not satisfy the frequently used
+pseudo-metric spaces. Such spaces do not satisfy the frequently used
 Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
@@
   \item $m$ is called additive if for $\mathcal K \subseteq_f \mathcal
-    H$ pariwise disjoint and $\bigcup_{K \in \mathcal K} K \in
+    H$ pairwise disjoint and $\bigcup_{K \in \mathcal K} K \in
     \mathcal H$, we have $m \Big(\bigcup_{K \in \mathcal K} K \Big) =
@@
       write $A \subseteq_c B$ if $A$ is a countable subset of
-      $B$.}$\mathcal K \subseteq_c \mathcal H$ pariwise disjoint, we
+      $B$.}$\mathcal K \subseteq_c \mathcal H$ pairwise disjoint, we
     say that $m$ is $\sigma$-additive.
@@
 \noindent
-{\em Prohorovs theorem:} The proof that a single finite measures on a
+{\em Prohorov's theorem:} The proof that a single finite measures on a
 Polish space is tight (i.e.\ inner regular with respect to compact
@@
 
-%\section*{Acknowledgements}
-%PP ist partly funded by the Vector foundation, grant
-%{\em LeanAI (P2022-0241)}.
-%RD acknowledges the support of the French Ministry of Higher Education and Research and the Hauts-de-France region.
+%\section*{Acknowledgements} We thank Johan Commelin, who helped PP to
+% get started with formalizing mathematics in Lean. PP ist partly
+% funded by the Vector foundation, grant {\em LeanAI (P2022-0241)}.
+% RD acknowledges the support of the French Ministry of Higher
+% Education and Research and the Hauts-de-France region.
 

### 2023-09-18 Manuscript/proof_outline.tex
add lean ref

@@
 The goal of our contribution is a formalization of the proof of that theorem in {\tt Lean}.
-We are using \leanline{Lean4} and its mathematical library \leanline{mathlib} \cite{mathlib2020lean}\footnote{All references to results in \leanline{mathlib} are accurate for commit {\tt 537e7f030d6df737d728c7c002fb1928f4615afa}, August 22, 2023.}.
+We are using \leanline{Lean4}\cite{moura2021lean} and its mathematical library \leanline{mathlib} \cite{mathlib2020lean}\footnote{All references to results in \leanline{mathlib} are accurate for commit {\tt 537e7f030d6df737d728c7c002fb1928f4615afa}, August 22, 2023.}.
 The Kolmogorov extension theorem is on the interface between measure theory and probability theory.

### 2024-07-05 Manuscript/proof_outline.tex

@@
 We start off from metric spaces, but quickly introduce projective families of measures.
-In all definitions, $\alpha$ will be some set (or type, since {\tt Lean} is a dependently typed language, \leanline{α : Type _}).
+In all definitions, $\alpha$ will be some set (or type, since {\tt Lean} is a dependently typed language, \leanline{α : Type*}).
 
@@
 are part of our own code. In the formalization of the above, we use
-\leanline{variable {ι : Type _} {α: ι → Type _}}, which fixes the
+\leanline{variable {ι : Type*} {α: ι → Type*}}, which fixes the
 index set $\iota$ and all spaces $\alpha_t, t\in\iota$ as global

### 2024-07-22 Manuscript/proof_outline.tex
use IsSetSemiring from mathlib, lint

@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-structure SetSemiring (C : Set (Set α))
+structure IsSetSemiring (C : Set (Set α))
     : Prop where
@@
 \begin{minted}[highlightlines={1-4}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-theorem setField_cylinders : SetField (cylinders α) 
-theorem setRing_cylinders :  SetRing (cylinders α)
-theorem setSemiringCylinders
-  : SetSemiring (cylinders α)   
+theoremIsSetField_cylinders :IsSetField (cylinders α) 
+theorem setRing_cylinders :  IsSetRing (cylinders α)
+theorem IsSetSemiringCylinders
+  : IsSetSemiring (cylinders α)   
 \end{minted}
@@
 theorem exists_disjoint_finset_diff_eq
-    (hC : SetSemiring C) (hs : s ∈ C)
+    (hC : IsSetSemiring C) (hs : s ∈ C)
     (I : Finset (Set α)) (hI : ↑I ⊆ C) :
@@
 \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-def diff₀ (hC : SetSemiring C)
+def diff₀ (hC : IsSetSemiring C)
     (hs : s ∈ C) (I : Finset (Set α)) (hI : ↑I ⊆ C)
@@
 %% \begin{minted}[highlightlines={1}, mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
-%% def indexedDiff₀ (hC : SetSemiring C)
+%% def indexedDiff₀ (hC : IsSetSemiring C)
 %%     (J : Finset (Set α)) (hJ : ↑J ⊆ C)
@@
 def extendContent
-    (hC : SetSemiring C)
+    (hC : IsSetSemiring C)
     (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
@@
     AddContent (cylinders α) :=
-  extendContent setSemiringCylinders
+  extendContent IsSetSemiringCylinders
     (kolmogorovFun P) (kolmogorovFun_empty hP)
@@
 %% def ofAddSubaddCaratheodory
-%%     (hC : SetSemiring C)
+%%     (hC : IsSetSemiring C)
 %%     (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
@@
 def Measure.ofAddContent
-    [mα : MeasurableSpace α] (hC : SetSemiring C)
+    [mα : MeasurableSpace α] (hC : IsSetSemiring C)
     (hC_gen : MeasurableSpace.generateFrom C = mα)
@@
 theorem sigma_additive_addContent_of_tendsto_zero
-    (hC : SetRing C)
+    (hC : IsSetRing C)
     (m : AddContent C)
@@
 
-%% lemma AddContent.sigma_additive_of_regular [Nonempty α] (hR : SetRing R) (m : AddContent R)
+%% lemma AddContent.sigma_additive_of_regular [Nonempty α] (hR : IsSetRing R) (m : AddContent R)
 %%     (hm_ne_top : ∀ {s} (_ : s ∈ R), m s ≠ ∞)
@@
     Measure (∀ i, α i) :=
-  Measure.ofAddContent setSemiringCylinders
+  Measure.ofAddContent IsSetSemiringCylinders
     generateFrom_cylinders (kolContent hP)

### 2024-07-22 Manuscript/proof_outline.tex
better name

@@
 theoremIsSetField_cylinders :IsSetField (cylinders α) 
-theorem setRing_cylinders :  IsSetRing (cylinders α)
+theorem isSetRing_cylinders :  IsSetRing (cylinders α)
 theorem IsSetSemiringCylinders

### 2024-09-12 article/article.tex
begin article

\documentclass{article}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[a4paper, total={\textwidth, 23cm}]{geometry}
\usepackage{setspace}
\usepackage{bbold}
\usepackage{chngcntr}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	citecolor=red,
	pdftitle={Rapport de Stage},
	pdfauthor={Etienne MARION},
	pdfpagemode=FullScreen,
}

\newcommand{\lean}[1]{\lstinline[language=lean]{#1}}

\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{tacticcolor}{rgb}{0.0, 0.1, 0.3}    % dark blue
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{stringcolor}{rgb}{0.5, 0.3, 0.2}    % brown
\definecolor{symbolcolor}{rgb}{0.1, 0.2, 0.7}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
\definecolor{attributecolor}{rgb}{0.7, 0.1, 0.1} % red
\definecolor{errorcolor}{rgb}{1, 0, 0}           % bright red

\usepackage{listings}
\def\lstlanguagefiles{lstlean.tex}
\lstloadlanguages{lean}
\lstset{language=lean}
\usepackage{scalefnt}

\newcommand{\sorry}[0]{\lean{sorry}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\imp}{\Rightarrow}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\inv}{^{-1}}
\newcommand{\eps}{\epsilon}
\newcommand{\type}{\texttt{Type}}
\newcommand{\prop}{\texttt{Prop}}
\newcommand{\prun}{\texttt{pr}_1}
\newcommand{\prde}{\texttt{pr}_2}
\newcommand{\inl}{\texttt{inl}}
\newcommand{\inr}{\texttt{inr}}
\newcommand{\skl}{\vspace{\baselineskip}}
\newcommand{\gui}[1]{\og{}#1\fg{}}
\newcommand{\Xge}[1]{X^{\ge#1}}
\newcommand{\Xle}[1]{X^{\le#1}}
\newcommand{\Xgt}[1]{X^{>#1}}
\newcommand{\Age}[1]{\mathcal{A}^{\ge#1}}
\newcommand{\Agt}[1]{\mathcal{A}^{>#1}}
\newcommand{\Ale}[1]{\mathcal{A}^{\le#1}}
\newcommand{\Xgtn}{X^{>n}}
\newcommand{\Xgtm}{X^{>m}}
\newcommand{\Xgtk}{X^{>k}}
\newcommand{\Xlen}{X^{\le n}}
\newcommand{\Alen}{\mathcal{A}^{\le n}}
\newcommand{\dx}{\mathrm{d}x}
\newcommand{\dy}{\mathrm{d}y}
\newcommand{\prth}[1]{\left(#1\right)}
\renewcommand{\empty}{\varnothing}
\newcommand{\tendl}[1]{\longrightarrow_{#1}}
\newcommand{\dbrack}[1]{\llbracket #1 \rrbracket}
\newcommand{\priv}{\,\backslash\,}
\newcommand{\restr}[1]{\vert_{#1}}
\newcommand{\mub}{\overline{\mu}}
\newcommand{\ox}{\otimes}
\renewcommand{\phi}{\varphi}
\newcommand{\piun}{{\pi_{\dbrack{1,n}}}}
\newcommand{\pizn}{{\pi_{\dbrack{0,n}}}}
\newcommand{\Xint}[1]{X^{\dbrack{#1}}}
\newcommand{\tost}{\to_*}
\newcommand{\ind}[1]{\mathbb{1}_{#1}}
\newcommand{\dmu}{\mathrm{d}\mu}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\theoremstyle{definition}
\newtheorem{defi}[thm]{Definition}
\newtheorem{eg}[thm]{Example}
\theoremstyle{remark}
\newtheorem*{nota}{Notation}

\begin{document}

	\begin{center}
		{\Large\bf A Formalization of the Ionescu-Tulcea Theorem in mathlib} \\
		\vspace{1cm}
		Auteur : Etienne Marion \\
		\vspace{1cm}
		ENS de Lyon, 46, Allée d’Italie, 69007 Lyon, France \\
		\texttt{etienne.marion@ens-lyon.fr}
	\end{center}

	\vspace{0.5cm}

	\section*{Abstract}

	\section{Introduction}
	Being able to talk about the joint distribution of an infinite family of random variables is crucial in probability theory. For example, one often requires a family of independent random variables. The existence of such a family relies on the existence of an infinite product measure. Indeed, given $(\Omega_i, \F_i, \mu_i)_{i\in\iota}$ a family of probability spaces, the existence of the product measure $\bigotimes_{i\in\iota}\mu_i$ yields a new probability space $(\prod_{i\in\iota}\Omega_i, \bigotimes_{i\in\iota}\F_i, \bigotimes_{i\in\iota}\mu_i)$, and the projections $X_i : \prod_{j\in\iota}\Omega_j \to \Omega_i$ give the desired family. For another example, consider discrete-time Markov chains. Given a measurable space $(E, \A)$ and a Markov kernel $\kappa : E \to E$, one might want to build a sequence $(X_n)_{n\in\N}$ of random variables with values in $E$ such that the conditional distribution of $X_{n+1}$ given $X_0, ..., X_n$ is $\kappa(X_n,\cdot)$. Such objects are fundamental in probability theory: families of independent variables allow to build more complicated objects, such as Brownian motion, while discrete-time Markov chains form a huge class of stochastic processes which contains random walks for instance. It so happens that those objects always exist without any restrictions on the spaces we consider. This is a direct consequence of the Ionescu-Tulcea theorem.

\end{document}

### 2024-09-12 article/article.tex
WIP article

@@
 \newcommand{\dmu}{\mathrm{d}\mu}
+\newcommand{\mathlib}{\texttt{mathlib}}
 
 \newtheorem{thm}{Theorem}[section]
+\newtheorem{cor}{Corollary}[thm]
 \newtheorem{lem}[thm]{Lemma}
@@
 
+	The goal of this contribution is to provide a formalization of the proof of the aforementioned theorem using the proof assistant Lean and the associated library \mathlib.
+It heavily relies on previous work done in [cite] which formalizes the Kolmogorov extension theorem.
+While this specific theorem is not used directly in the formalization presented here, we use the notion of measurable cylinders, projective families or additive contents which were introduced to prove the Kolmogorov extension theorem.
+
+	\section{Statement of the theorem}
+	The Ionescu-Tulcea theorem is about the existence of a certain Markov kernel. We therefore start with a reminder about those.
+
+	\begin{defi}\label{def:markov-kernel}
+		Let $(X,\A)$ and $(Y,\B)$ be two measurable spaces. A \emph{Markov kernel} from $X$ to $Y$ is a map $\kappa : X \times \B \to [0,1]$ such that:
+		\begin{itemize}
+			\item for any $x \in X$, $\kappa(x,\cdot)$ is a probability measure;
+			\item for any $B \in \B$, $\kappa(\cdot,B)$ is measurable.
+		\end{itemize}
+		One can therefore consider a Markov kernel as a measurable map which to any point in $X$ associates a probability distribution over the space $Y$.
+	\end{defi}
+
+	To ease the writing, let us give some notations.
+
+	\begin{nota}
+		If $I$ is a set of indices and $(X_i, \A_i)_{i \in I}$ a family of measurable spaces, then:
+		\begin{itemize}
+			\setstretch{1.5}
+			\item we write $X$ for $\prod_{i \in I} X_i$ and $\A$ for $\bigotimes_{i \in I} \A_i$;
+			\item if $J \subseteq I$, we write $X^J$ for $\prod_{i \in J} X_i$ and $\A^J$ for $\bigotimes_{i \in J} \A_i$;
+			\item if $I = \N$ and $n \in \N$ we write $\Xge{n}$ for $X^{\set{k \in \N \mid k \ge n}}$ and $\Age{n}$ for $\A^{\set{k \in \N \mid k \ge n}}$, with similar notations for $\le, >, <$;
+			\item if $J \subseteq I$, we write $\pi_J$ for the canonical map from $X$ to $X^J$;
+			\item if $ K \subseteq J \subseteq I$, we write $\pi_{J\to K}$ for the canonical map from $X^J$ to $X^K$.
+		\end{itemize}
+	\end{nota}
+
+	Let us now state the theorem.
+
+	\begin{thm}[Ionescu-Tulcea]\label{thm:it}
+		Let $(X_n,\A_n)_{n\in\N}$ be a family of measurable spaces.
+Let $(\kappa_n)_{n\in\N}$ be a family of Markov kernels such that for any $n$, $\kappa_n$ is a kernel from $X_0 \times ...
+\times X_n$ to $X_{n+1}$.
+Then there exists a unique Markov kernel $\eta : X_0 \to \prod_{n\ge1}X_n$ such that for any $n\ge1$,
+		$${\pi_n}_*\eta = \kappa_0 \otimes ... \otimes \kappa_{n-1}.$$
+	\end{thm}
+
+	Let's start by explaining why this theorem implies the existence of product measures and discrete-time Markov chains.
+
+	\begin{cor}\label{cor:product-measure}
+		Let $(X_i, \A_i, \mu_i)_{i\in\iota}$
+	\end{cor}
+
 \end{document}

### 2025-01-18 article/article.tex
article
wip

@@
 \documentclass{article}
-\usepackage[french]{babel}
 \usepackage[T1]{fontenc}
@@
 		\vspace{1cm}
-		Auteur : Etienne Marion \\
+		Author : Etienne Marion \\
 		\vspace{1cm}
@@
 
-	Let's start by explaining why this theorem implies the existence of product measures and discrete-time Markov chains.
+	By taking each kernel to be constant we easily get the product of a countable family of probability measures, which can then be used to build the product of an arbitrary family of measures.
+Likewise if we take $X_n = X$ for all $n$ and $\kappa_n = \kappa$ a kernel from $X \to X$ we get a Markov kernel from $X$ to $X^{\N_+}$ which, if we compose it with a measure $\mu$ at the beginning, gives the measure associated to a Markov chain with initial distribution $\mu$ and transition kernel $\kappa$.
 
-	\begin{cor}\label{cor:product-measure}
-		Let $(X_i, \A_i, \mu_i)_{i\in\iota}$
-	\end{cor}
+	We will now give a sketch of the proof of the theorem to highlight the elements which are required for the formalization.
+The idea is to use the Carathéodory extension theorem which allows to extend a measure once we have defined it on a restricted family of measurable sets for which it is to give the expression of the measure.
+In what follows we consider $(X_n, \A_n)_{n\in\N}$ a family of measurable spaces and we set $\pi_n := \piun$.
+we also consider of Markov kernels $(\kappa_n : \Xlen \tost X_{n+1})_{n\in\N}$ and we set $\eta_{k,n} := \kappa_k \ox ...
+\ox \kappa_{n-1} : \Xle{k} \tost \Xint{k+1,n}$ and $\eta_n := \eta_{0,n}$.
 
+	\begin{defi}
+		\label{def:cylindres}
+		Sets of the form $A \times \Xgtn$, where $n \ge 1$ and $A \in \Alen$ are called cylinders.
+	\end{defi}
+
+	By writing $A \times \Xgtn = \pi_n\inv(A)$ it is easy to check that the set of cylinders is closed by union and complement, and thus it is an algebra of sets which generates the product $\sigma$-algebra.
+We now have to define a measure over cylinders which we will then extend to the whole $\sigma$-algebra.
+We also do this for cylinders in $\Xge{n}$ for $n\ge1$ as we will need it for the extension afterwards.
+
+	\begin{defi}
+		\label{def:content}
+		Let $x_0 \in X_0$. If $A := A' \times \Xgtn \subseteq \Xge{1}$ is a cylinder we set
+		$$P_0(x_0, A) := \eta_n(x_0, A').$$
+		This map is well defined because if $n \le m$ then
+		\begin{align*}
+			\eta_m\prth{x_0, A' \times \Xint{n+1,m}} &= \eta_n \ox \eta_{n+1,m} \prth{x_0, A' \times \Xint{n+1,m}} \\
+			&= \int_{A'} \eta_{n+1, m}\prth{x_0, ..., x_n, \Xint{n+1,m}} \eta_n(x_0, \dx_1...\dx_n) \\
+			&= \int \ind{A'} \eta_n(x_0, \dx_1...\dx_n) \\
+			&= \eta_n(x_0, A').
+		\end{align*}
+
+		For $k \ge 1$ and $A := A' \times \Xgtn \subseteq \Xgt k$,we define likewise
+		$$P_k (x_0, ..., x_k, A) := \eta_{k,n} (x_0,...,x_k,A').$$
+	\end{defi}
+
+	\begin{lem}
+		\label{lem:add-cont}
+		For any $k \in \N$ and any $(x_0,...,x_k) \in \Xle{k}$, the map $P_k(x_0,...,x_k, \cdot)$ is an additive content over the cylinders.
+Namely if $A, B$ are two disjoint cylinders then $P_0(x_0,...,x_k,A \cup B) = P_0(x_0,...,x_k,A) + P_0(x_0,...,x_k,B)$.
+	\end{lem}
+
+	\begin{proof}
+		We only give the proof in the case $k=0$ as the other cases are identical.
+Let $A := A' \times \Xgtm$ and $B := B' \times \Xgtn$ be two disjoint cylinders and assume without loss of generality that $m \le n$.
+We then set $A'' := A' \times \Xint{m+1,n}$.
+As $A \cap B = \empty$ this implies $A'' \cap B' = \empty$.
+Moreover, $A \cup B = (A'' \cup B') \times \Xgtn$.
+Consequently :
+		$$P_0(x_0,A\cup B) = \eta_n(x_0,A'' \cup B') = \eta_n(x_0,A'') + \eta_n(x_0,B') = P_0(x_0,A) + P_0(x_0,B).\eqno\mbox{\qedhere}$$
+	\end{proof}
+
+	To apply Carathéodory theorem it remains to prove that $P_0(x_0,\cdot)$ is $\sigma$-additive over cylinders.
+Because it is an additive content and cylinders are an algebra of sets it is enough to show the following result:
+
+	\begin{lem}
+		\label{lem:tendsto_zero}
+		Let $x_0 \in X_0$. If $(A_n)_{n\in\N}$ is a non-increasing sequence of cylinders with emty intersection then $P_0(x_0, A_n) \tendl{n\to\infty} 0$.
+	\end{lem}
+
+	\begin{proof}
+		For any $n$ we write $A_n := B_n \times \Xgt{a_n}$.
+In this proof if $A \subseteq \Xge{1}$ et $(x_1, ..., x_n) \in \Xint{1,n}$ we will denote by $A(x_1,...,x_n) := \set{y \in \Xgtn : (x_1,...,x_n,y) \in A}$ the associated section of $A$.
+
+		We use contraposition.
+Assume there exists $\eps>0$ such that for any $n \in \N$, $P_0(x_0,A_n) \ge \eps$.
+We wish to prove that the intersection of $A_n$ is nonempty.
+We define the following sequence of functions:
+		\begin{align*}
+			f_n : X_1 &\to [0, 1] \\
+			x_1 &\mapsto P_1(x_0,x_1, A_n(x_1)).
+		\end{align*}
+		It allows us to write
+		\begin{align*}
+			P_0(x_0,A_n) &= \eta_{a_n}(x_0,B_n) \\
+			&=\kappa_0 \ox \eta_{1, a_n}(B_n) \\
+			&= \int_{X_1} \eta_{1, a_n}(x_0, x_1, B_n(x_1)) \kappa_0(x_0, \dx_1) \\
+			&= \int_{X_1} P_1(x_0,x_1,A_n(x_1)) \kappa_0(x_0, \dx_1) \\
+			&= \int_{X_1} f_n(x_1) \kappa_0(x_0, \dx_1).
+		\end{align*}
+		Because $(A_n)$ is non-increasing it is also the case of $(f_n)$.
+We set $g := \lim_{n\to\infty} f_n$.
+By dominated convergence theorem we get $\eps \le P_0(x_0,A_n) \tendl{n\to\infty} \int_{X_1} g(x_1) \kappa_0(x_0,\dx_1)$.
+As $\kappa_0(x_0, \cdot)$ is a probability measure there exists $x_1 \in X_1$ such that $g(x_1) \ge \eps$, and so for any $n\in\N$, $f_n(x_1) \ge \eps$ which implies that $P_1(x_0,x_1, A_n(x_1)) \ge \eps$.
+
+		We can thus repeat the same step replacing $P_0(x_0,\cdot)$ by $P_1(x_0,x_1,\cdot)$ and $A_n$ by $A_n(x_1)$.
+We get a sequence $x := (x_k)_{k\ge1}$ such that for any $n\in\N$ and $k\ge1$ we have $P_k(x_0,...,x_k,A_n(x_1,...,x_k)) \ge \eps$.
+It remains to show that $x \in \bigcap_{n\in\N}A_n$.
+
+		Let $n\in\N$.
+As $A_n = B_n \times \Xgt{a_n}$ we in particular have that $A_n(x_1,...,x_{a_n})$is either $\Xgt{a_n}$ or empty according to wether $(x_1,...,x_{a_n})$ belongs to $B_n$ or not.
+But as $P_{a_n}(x_0,...,x_{a_n},A_n(x_1,...,x_{a_n})) \ge \eps > 0$ we have $A_n(x_1,...,x_{a_n}) = \Xgt{a_n}$, which proves that $x \in A_n$.
+	\end{proof}
+
+	This concludes the main argument of the proof.
+The measurability is easily provided by applying the $\pi$-$\lambda$ theorem.
+We will now focus on those aspects of the proof which proved difficult to formalize and how they were overcome.
+
+	\section{Dealing with composition of kernels}
+
+	One of the key tools for the Ionescu-Tulcea theorem is to be able to speak about the composition of Markov kernels.
+Indeed that is necessary even to be able to state the theorem.
+However formalizing this in a way which allows us to smoothly write products proved to be quite challenging.
+
+	One common rule in formalization is that if we use "..." in a proof it will likely be hard to formalize.
+In our case the statement of Theorem~\ref{thm:it} itself makes use of these when we write the composition of kernels, and they hide a real problem.
+
+	Transition kernels and their composition are already defined in Lean.
+However we need to take the composition of a finite family of kernels which does not exist in Mathlib, and here is why.
+By hand, if we have three Markov kernels $\kappa : X \tost Y$, $\eta : X \times Y \tost Z$ and $\xi : X \times Y \times Z \tost T$ we can define $\kappa \ox \eta \ox \xi := (\kappa \ox \eta) \ox \xi$.
+Then we easily get $\kappa \ox \eta \ox \xi = \kappa \ox (\eta \ox \xi)$.
+But here we forget a detail: $X \times Y \times Z$ does not mean anything a priori.
+We can write $(X \times Y) \times Z$ or $X \times (Y \times Z)$ but those are two different types.
+In paper maths they are canonically isomorphic and we often act as though they were the same but it Lean this isomorphism has to be made explicit.
+
+	Coming back to kernels I can write that $\xi$ is a kernel from $X \times (Y \times Z)$ to $T$.
+Then $(\kappa \ox \eta) \ox \xi$ is a kernel from $X$ to $(Y \times Z) \times T$.
+But in that case $\kappa \ox (\eta \ox \xi)$ is not even well defined because $\xi$ should start from $(X \times Y) \times Z$.
+If we were to add an isomorphism to define this object anyway the equality $(\kappa \ox \eta) \ox \xi = \kappa \ox (\eta \ox \xi)$ would still make no sense because on the left we have a kernel from $X$ to $(Y \times Z) \times T$ while on the right it is a kernel from $X$ to $Y \times (Z \times T)$, so they do not have the same type.
+
+	A similar issue arises when formalizing Theorem~\ref{thm:it}: if we have two kernels $\kappa : X_0 \tost \Xint{1,n}$ and $\eta : \Xlen \tost \Xint{n+1,m}$, their composition does not make sense, we need isomorphisms.
+First to see $\eta$ as a kernel from $X_0 \times \Xint{1,n}$ to $\Xint{n+1,m}$ to get $\kappa \ox \eta : X_0 \tost \Xint{1,n} \times \Xint{n+1,m}$, then to see $\kappa \ox \eta$ as a kernel from $X_0$ to $\Xint{1,m}$.
+To solve this problem we need to make these isomorphisms explicit and then hide them inside a black box to be able to use it without worrying about the implementation.
+
+	This work was initiated by Degenne. He introduced a compositon of kernels indexed by integer intervals which makes the equivalence invisible.
+	\begin{lstlisting}
+	def compProdNat {i j k : ℕ} (κ : Kernel ((l : Iic i) → X l)
+		((l : Ioc i j) → X l))
+		(η : Kernel ((l : Iic j) → X l) ((l : Ioc j k) → X l)) :
+		Kernel ((l : Iic i) → X l) ((l : Ioc i k) → X l)
+	\end{lstlisting}
+	This takes two kernels $\kappa : \Xle{i} \tost X^{\dbrack{i+1,j}}$ and $\eta : \Xle{j} \tost X^{\dbrack{j+1,k}}$ and returns a kernel from $\Xle{i}$ to $X^{\dbrack{i+1,k}}$.
+Degenne then proves that this composition is associative, and uses it to define
+	\begin{lstlisting}
+	def kerNat (κ : (k : ℕ) → Kernel ((l : Iic k) → X l) (X (k + 1))) (i j : ℕ) :
+		Kernel ((l : Iic i) → X l) ((l : Ioc i j) → X l).
+	\end{lstlisting}
+	This takes as an argument a family of kernels $(\kappa_k)_{k\in\N}$ with $\kappa_k : \Xle{k}\to X_{k+1}$ just as in Theorem~\ref{thm:it}.
+Then given two integers $i$ and $j$ it returns a kernel \texttt{kerNat $\kappa$ i j} which is the composition
+	$$\kappa_i \ox \kappa_{i+1} \ox \cdots \ox \kappa_{j-1}.$$
+	We are therefore now able to speak about those compositions which are necessary to state the theorem.
+However \texttt{kerNat} proved to be not enough for later purposes.
+Indeed after proving the theorem and trying to write some API we thought that it would be best to to have a kernel which does not necesarrily starts at time 0 but at any instant.
+This  would prove nicer for example when it comes to write formulas for the conditional expectation against the Ionescu-Tulcea kernel (see Section).
+However we bumped into an issue.
+Having different kernels starting at different times was proving hard to manipulate even with the compositions that were implemented.
+That is why I chose to add another layer on top of \texttt{kerNat} which proved to work out smoothly.
+	\begin{lstlisting}
+		content
+	\end{lstlisting}
 \end{document}

### 2025-01-21 article/article.tex
WIP

@@
 \usepackage{subcaption}
-\usepackage[a4paper, total={\textwidth, 23cm}]{geometry}
+\usepackage[a4paper, total={16cm, 23cm}]{geometry}
 \usepackage{setspace}
@@
 	\begin{lstlisting}
-		content
+	def partialKernel (a b : ℕ) :
+		Kernel ((i : Iic a) → X i) ((i : Iic b) → X i)
+	\end{lstlisting}
+	Given integer $a$ and $b$ this function will output a kernel from $\Xle{a}$ to $\Xle{b}$.
+The idea is that the input of this kernel is the full trajectory up to time $a$ and the output is the distribution of the full trajectory up to time $b$.
+Of course this distribution is deterministic up to time $a$ and then we compose with \texttt{kerNat $\kappa$ a b}.
+The advantage of this construction is that in the end we always manipulate kernels with domains and codomains of the form $\Xlen$ for some integer $n$ which makes all the composition much easier.
+For example if we take $a$, $b$ and $c$ three integers satisfying $a \land c \le b$ we get
+	\begin{lstlisting}
+	(partialKernel κ b c) ∘ₖ (partialKernel κ a b) = partialKernel κ a c
+	\end{lstlisting}
+	where $\circ_{\rm k}$ is the composition of kernels.
+This nice property has the advantage to be expressed with a regular operation on kernels rather than one which would be made up for the occasion.
+With this tool we get a satisfactory way of expressing the composition of multiple kernels.
+
+	\section{Sections of sets}
+
+	We see that the proof of Lemma~\ref{lem:tendsto_zero} heavily relies on the use of sections of sets.
+This however is not well suited for formalization purposes because of the following.
+The set $\prod_{n\in\N} X_n$ is represented in lean as the type of functions $f$ with domain $\N$ and such that for any $n \in \N$ $f(n)$ has type $X_n$.
+We encounter the same issue as before: $\prod_{n\in\N} X_n$ is different from $X_0 \times \Xge{1}$ so we have to use isomorphisms.
+Moreover, if $A \subseteq \prod_{n\in\N} X_n$ and $x \in X_0$then $A_x \subseteq \Xge{1}$.
+	So $A_x$ is a set of sequences indexed by $\N_+$ which is not a subset of $\N$ but rather a subtype in Lean.
+This makes things hard to manipulate because we want to deal with an element of $\N_+$ we have to provide a proof that it is positive.
+This would have to be done not only for non-zero integers but also for integers greater than $2$, $3$ and so on with the same troubles each time.
+Therefore Sébastien Gouëzel suggested another approach.
+
+	The idea relies on the notion of marginals exposed in Section~4 of [Van2024].
+I generalized their idea as follows: given $a$ and $b$ integers, $f : X \to [0,+\infty]$ and $x \in X$ I defined \texttt{lamrginalPartialKernel $\kappa$ a b f x} as
+	$$\int_{\Xle{b}} f(y \leadsto_b x) \mathrm{d}\eta_{a,b}(x_0,...,x_a).$$
+
+	This is great because it allows to write $$\texttt{lamrginalPartialKernel $\kappa$ b c (lamrginalPartialKernel $\kappa$ a b f x) x},$$
+	in other words we can first integrate between $a$ and $b$ and then between $b$ and $c$. This allows us to peel off the measure of a cylinder as the integral of its indicator:
+	$$P_0(x_0, A_n) = \int_{X_1} P_1(x_0, x_1, A_n(x_1)) \kappa_0(x_0,\dx_1).$$
+
+%	In this paper they explain how we can't directly write
+%	$$\int_{\R^{n+m}}f(x)\dx=\int_{\R^n}\int_{\R^m}f(x,y)\dx\dy$$
+%	because $\R^{n+m}$ and $\R^n\times\R^m$ are not the same type in lean.
+Therefore given $(X_i, \A_i, \mu_i)_{i\in\iota}$ a family of measure spaces, $I$ a finite subset of $\iota$ and $f : \prod_{i\in\iota} X_i \to [0,+\infty]$,	they define the following function
+%	\begin{align*}
+%		\int\cdots\int_{i\in I}f\dmu_i : \prod_{i\in\iota}X_i &\to [0,+\infty] \\
+%		x &\mapsto \int_{\prod_{i \in I}X_i}f(y \leadsto_I x) \mathrm d \bigotimes_{i\in A}\mu_i(dy),
+%	\end{align*}
+%	where $y \leadsto_J x$ is defined as
+%	$$(y \leadsto_J x)_i :=
+%	\begin{cases}
+%		y_i, &\text{si $i \in J$} \\
+%		x_i, &\text{sinon}
+%	\end{cases}.
+%	$$
+%	What is happening here is that we are computing
+
+	\section{Dependence in coordinates}
+
+	This trick of considering some function as defined over $X$ while it really only depends on the coordinates $\ge n$ brings the necessity of introducing a simple way to indicate which variables a given function actually depends on.
+We therefore introduced the following predicate.
+	\begin{lstlisting}
+	def DependsOn (f : ((i : ι) → α i) → β) (s : Set ι) : Prop :=
+	∀ ⦃x y⦄, (∀ i ∈ s, x i = y i) → f x = f y
+	\end{lstlisting}
+	The idea is that if $f : \prod_{i\in\iota} \alpha_i \to \beta$ and $s \subset \iota$, \texttt{DependsOn f s} means that to check $f(x) = f(y)$ it is enough to show that for any $i \in s$, $x_i = y_i$.
+In other words $f$ only depends on coordinates indexed by $s$.
+For example, if $f$ only depends on coordinates $\le b$ and we integrate over its coordinates between $a$ and $b$ then the output only depends on coordinates $\le a$.
+This is formalized as follows.
+	\begin{lstlisting}
+	theorem dependsOn_lmarginalPartialKernel (a : ℕ) {b : ℕ}
+		{f : ((n : ℕ) → X n) → ℝ≥0∞}
+		(hf : DependsOn f (Iic b)) (mf : Measurable f) :
+		DependsOn (lmarginalPartialKernel κ a b f) (Iic a)
 	\end{lstlisting}

### 2025-01-21 article/article.tex
WIP

@@
 \newcommand{\dmu}{\mathrm{d}\mu}
+\newcommand{\dlamb}{\mathrm{d}\lambda}
 \newcommand{\mathlib}{\texttt{mathlib}}
@@
 	This takes as an argument a family of kernels $(\kappa_k)_{k\in\N}$ with $\kappa_k : \Xle{k}\to X_{k+1}$ just as in Theorem~\ref{thm:it}.
-Then given two integers $i$ and $j$ it returns a kernel \texttt{kerNat $\kappa$ i j} which is the composition
+Then given two integers $i$ and $j$ it returns a kernel \lean{kerNat κ i j} which is the composition
 	$$\kappa_i \ox \kappa_{i+1} \ox \cdots \ox \kappa_{j-1}.$$
 	We are therefore now able to speak about those compositions which are necessary to state the theorem.
-However \texttt{kerNat} proved to be not enough for later purposes.
+However \lean{kerNat} proved to be not enough for later purposes.
 Indeed after proving the theorem and trying to write some API we thought that it would be best to to have a kernel which does not necesarrily starts at time 0 but at any instant.
@@
 Having different kernels starting at different times was proving hard to manipulate even with the compositions that were implemented.
-That is why I chose to add another layer on top of \texttt{kerNat} which proved to work out smoothly.
+That is why I chose to add another layer on top of \lean{kerNat} which proved to work out smoothly.
 	\begin{lstlisting}
@@
 The idea is that the input of this kernel is the full trajectory up to time $a$ and the output is the distribution of the full trajectory up to time $b$.
-Of course this distribution is deterministic up to time $a$ and then we compose with \texttt{kerNat $\kappa$ a b}.
+Of course this distribution is deterministic up to time $a$ and then we compose with \lean{kerNat κ a b}.
 The advantage of this construction is that in the end we always manipulate kernels with domains and codomains of the form $\Xlen$ for some integer $n$ which makes all the composition much easier.
@@
 	The idea relies on the notion of marginals exposed in Section~4 of [Van2024].
-I generalized their idea as follows: given $a$ and $b$ integers, $f : X \to [0,+\infty]$ and $x \in X$ I defined \texttt{lamrginalPartialKernel $\kappa$ a b f x} as
+I generalized their idea as follows: given $a$ and $b$ integers, $f : X \to [0,+\infty]$ and $x \in X$ I defined \lean{lamrginalPartialKernel κ a b f x} as
 	$$\int_{\Xle{b}} f(y \leadsto_b x) \mathrm{d}\eta_{a,b}(x_0,...,x_a).$$
@@
 	\end{lstlisting}
-	The idea is that if $f : \prod_{i\in\iota} \alpha_i \to \beta$ and $s \subset \iota$, \texttt{DependsOn f s} means that to check $f(x) = f(y)$ it is enough to show that for any $i \in s$, $x_i = y_i$.
+	The idea is that if $f : \prod_{i\in\iota} \alpha_i \to \beta$ and $s \subset \iota$, \lean{DependsOn f s} means that to check $f(x) = f(y)$ it is enough to show that for any $i \in s$, $x_i = y_i$.
 In other words $f$ only depends on coordinates indexed by $s$.
@@
 	\end{lstlisting}
+
+	\section{Writing an API}
+
+	When formalizing a new mathematical object one usually does not stop at writing the definition.
+Indeed the objected is intended to be usable by all the users without having to know how the object was implemented.
+That is why it is important to accompany the object with many results which allow other users to use the object without knowing the implementation details.
+
+	We now present some of these results.
+The API is still very basic and will be developed as other users pick up the object to formalize more objects.
+The best way to find new lemmas to formalize is to choose a target statement and try and formalize it.
+In our case we aim to prove the following result.
+
+	\begin{nota}
+		For $n \in \N$ we set $\F_n := \pi_{\dbrack{0,n}}\inv\prth{\Alen}$.
+	\end{nota}
+
+	\begin{thm}
+		\label{thm:condexp}
+		Let $a, b \in \N$ with $a \le b$.
+Let $u \in \Xle{a}$.
+Let $f : X \to \R$ be an integrable function with respect to $\lambda_a(u)$.
+Then for $\lambda_a(u)$-almost every $x := (x_n)_{n\in\N}$, $f$ is integrable with respect to $\lambda_b(x_0,...,x_b)$ and
+		$$\E_{\lambda_a(u)}[f \mid \F_n] (x) = \int_X f(y) \lambda_b(x_0,...,x_b, \dy).$$
+	\end{thm}
+
+	Let us give the proof to highlight the necessary intermediate results. Take $A \in \Ale{b}$ and set $\pi := \pi_{\dbrack{0,b}}$. Our goal is to prove that
+	$$\int_{\pi\inv(A)} \int_X f(y) \lambda_b(\pi(x), \dy)\lambda_a(u,\dx) = \int_{\pi\inv(A)} f \dlamb_a.$$
+	We start by rewriting the outer integral using the push-forward measure:
+	$$\int_{\pi\inv(A)} \int_X f(y) \lambda_b(\pi(x), \dy)\lambda_a(u,\dx) = \int_A \int_X f(y) \lambda_b(\pi(x), \dy)(\pi_*\lambda_a)(u,\dx).$$
+	The first we will need is that $\pi_*\lambda_a = \eta_{a,b}$.
+This is quite obvious: as $\lambda_a$ gives the whole trajectory, keeping only the cooordinates up to time $b$ yields the trajectory up to time $b$.
+This is formalized as below.
+	\begin{lstlisting}
+theorem ionescuTulceaKernel_proj (a b : ℕ) :
+(ionescuTulceaKernel κ a).map (frestrictLe b) = partialKernel κ a b
+	\end{lstlisting}
+	So we get $$\int_{\pi\inv(A)} \int_X f(y) \lambda_b(\pi(x), \dy)\lambda_a(u,\dx) = \int_A \int_X f(y) \lambda_b(\pi(x), \dy)\eta_{a,b}(u,\dx).$$
 \end{document}
